### 2024-04-02 paper/paper.qmd
further clean data and update on graphs

---
title: "Stream of Conciousness Genre and Mental Landscapes"
subtitle: "Word Frequency and Mental Health Themes in Stream of Consciousness Works by James Joyce, Virginia Woolf, Marcel Proust, and Fyodor Dostoevsky from Project Gutenberg"
author: Quang Mai
thanks: "Code and data are available at: https://github.com/ponolite/stream-consciousness-language.git"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(tidytext)
library(rmarkdown)
library(gutenbergr)
library(tm)
library(dplyr)
library(ggplot2)
library(scales)

all_books <- read_csv(here::here("data/raw_data/all_books.csv"))
combined_books <- read_csv(here::here("data/raw_data/combined_books.csv"))
portrait_words <- read_csv(here::here("data/analysis_data/portrait_words.csv"))
swann_words <- read_csv(here::here("data/analysis_data/swann_words.csv"))
dalloway_words <- read_csv(here::here("data/analysis_data/dalloway_words.csv"))
notes_words <- read_csv(here::here("data/analysis_data/notes_words.csv"))

```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

frequency <- bind_rows(mutate(portrait_words, author = "James Joyce"),
                       mutate(swann_words, author = "Marcel Proust"),
                       mutate(dalloway_words, author = "Virginia Woolf"),
                       mutate(notes_words, author = "Fyodor Dostoevsky")) |>
  mutate(word = str_extract(word, "[a-z']+")) |>
  count(author, word) |>
  group_by(author) |>
  mutate(proportion = n / sum(n)) |>
  select(-n) |>
  pivot_wider(names_from = author, values_from = proportion)

```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

ggplot(frequency, aes(x = proportion, y = `James Joyce`, 
                            color = abs(`James Joyce` - proportion))) +
  geom_abline(color = "gray50", lty = 2) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.3, height = 0.3) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  scale_color_gradient(limits = c(0, 0.001), 
                       low = "lightyellow2", high = "gray80") +
  facet_wrap(~author, ncol = 4) +
  theme(legend.position="none") +
  labs(y = "James Joyce", x = NULL)

# Visualize word frequencies (e.g., bar plot or word cloud)
# Example using ggplot2 for bar plot
top_words <- head(word_freq, 20)
ggplot(top_words, aes(x = reorder(word, -n), y = n)) +
  geom_bar(stat = "identity") +
  labs(x = "Word", y = "Frequency", title = "Top 20 Most Frequent Words")

```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-04-02 scripts/02-data_cleaning.R

@@ -1,11 +1,10 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Clean the self-generated datasets 
+# Author: Quang Mai
+# Date: 30 March 2024 
+# Contact: quangpono@gmail.com
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Read four raw datasets containing the four novels
 
 #### Workspace setup ####
 library(tidyverse)
@@ -15,6 +14,8 @@ library(gutenbergr)
 library(tm)
 library(dplyr)
 library(ggplot2)
+library(scales)
+
 
 #### Clean data ####
 portrait <- read_csv("data/raw_data/portrait.csv")
@@ -22,7 +23,7 @@ swann <- read_csv("data/raw_data/swann_way.csv")
 dalloway <- read_csv("data/raw_data/mrs_dalloway.csv")
 notes <- read_csv("data/raw_data/notes_underground.csv")
 
-# Tokenize the words from the text
+# Tokenize the words from the text, breaking the text down into individual words to insert them into individual datasets
 # Codes sourced from https://medium.com/the-data-nudge/nlp-basics-exploring-word-frequency-with-the-tidytext-r-package-ac35a6d805f4
 
 portrait_tokenized <- portrait |>
@@ -69,59 +70,9 @@ dalloway_words <- dalloway_clean |>
 notes_words <- notes_clean |>
   count(word, sort = TRUE)
 
-# Convert to Corpus
-corpus <- Corpus(VectorSource(text$text))
-
-# Perform text preprocessing
-corpus_clean <- tm_map(corpus, content_transformer(tolower))
-corpus_clean <- tm_map(corpus_clean, removePunctuation)
-corpus_clean <- tm_map(corpus_clean, removeNumbers)
-corpus_clean <- tm_map(corpus_clean, removeWords, stopwords("english"))
-corpus_clean <- tm_map(corpus_clean, stripWhitespace)
-
-# Convert the corpus to a tidy format
-tidy_corpus <- tidy(corpus_clean)
-
-# Calculate word frequencies
-word_freq <- tidy_corpus %>%
-  count(word, sort = TRUE)
-
[…]

### 2024-04-03 paper/paper.qmd
construct graphs

@@
 ---
-title: "Stream of Conciousness Genre and Mental Landscapes"
-subtitle: "Word Frequency and Mental Health Themes in Stream of Consciousness Works by James Joyce, Virginia Woolf, Marcel Proust, and Fyodor Dostoevsky from Project Gutenberg"
+title: "The Linguistic Landscape of Stream-of-Consciousness Literature"
+subtitle: "Exploring Word Frequency and Mental Health Themes in the Works of Joyce, Woolf, Proust, Mansfield and Eliot from Project Gutenberg"
 author: Quang Mai
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This project focuses on understanding the language used by renowned, transnational stream of consciousness authors James Joyce, Virginia Woolf, Marcel Proust, Katherine Mansfield and T.S Eliot.
+By analyzing the word frequency of their famous works, I explore mental health themes like anxiety, depression, trauma, and existential angst.
+Through word frequency analysis of nine novels, I aim to uncover shared linguistic patterns and gain insights into the authors' mental states, offering a glimpse into themes of self-identity and existential contemplation"
 format: pdf
@@
 library(scales)
+library(knitr)
+library(wordcloud)
 
+portrait <- read_csv(here::here("data/raw_data/portrait.csv"))
+portrait_clean <- read_csv(here::here("data/analysis_data/portrait_clean.csv"))
 all_books <- read_csv(here::here("data/raw_data/all_books.csv"))
@@
 dalloway_words <- read_csv(here::here("data/analysis_data/dalloway_words.csv"))
-notes_words <- read_csv(here::here("data/analysis_data/notes_words.csv"))
+bliss_words <- read_csv(here::here("data/analysis_data/bliss_words.csv"))
+prufrock_words <- read_csv(here::here("data/analysis_data/prufrock_words.csv"))
 
@@
 
-The remainder of this paper is structured as follows. @sec-data....
+The American family landscape is undergoing a significant transformation, marked by a growing trend of childlessness.
+Traditionally viewed as a social norm, having children now becomes a choice [@voluntaryChild].
+We examine factors such as age, gender, education, socioeconomic class, health conditions and marital status, and how they affect childbearing rates from 1972 to 2022.
+Existing research suggests a complex interplay of factors that contribute to childlessness, including economic considerations, rising educational attainment, shifting social norms related to career and family life, and individual preferences [@voluntaryChild].
+
+Texts from a total of nine novels were retrieved from the volunteer archive to examine the mental health themes of famous stream of consciousness authors from the Modernst Era of literature from late 19th century to the mid-20th century
+
+A Portrait of the Artist as a Young Man, Dubliners, Mrs Dalloway, Jacob's Room, Swann Way, Bliss, The Garden Story, The Waste Land and The Love Song of J. Alfred Prufrock
+
+By analyzing the datasets, I seek to pose and answer crucial questions: What are some important factors contributing to this trend?
+How does childlessness manifest differently across demographic groups?
+Finally, what are the potential social, economic, and individual consequences of this new family structure in the United States?
+Understanding these dynamics is crucial in having an informed policy discussion and navigating the implications of childlessness in a rapidly changing world.
+
+Thus, the estimand is the correlation between stream of consciousness authors and words related to mental health themes appearing in their famous works.
+This is considered in terms of those who would have a third birth only if the first two children were of the same sex.
+Through our analysis, we found that the childless rate is consistently on the rise from 1972 to 2022, with a correlation coefficient of 0.73 between the childless percentage and time (or a 6.99% increase over the last 50 years).
+Factors that directly contribute to this rise are higher educational attainment, higher socioeconomic class, lower gender disparity, better health status and dissolution of marriages through divorces, separation and being widowed.
+In addition, global events like pandemics and recessions also heighten the rate of childlessness from 1972-2022, including the AIDs pandemic, 9/11, 2008 Recession and the COVID-19 pandemic.
+During the 2008 Recession, from 2008 to 2014, the rate of childlessness increases the most, at 3%.
+
+To further understand the correlation between stream of consciousness novels and mental health themes, in [Introduction], our paper briefly discusses the nature of stream of consciousness literature, relevant authors and the works that we've chosen to analyze.
+Subsequently, in [Data] and [Result], we talk about the nature of the data obtained and analyze the results garnered from the data with suitable tables and charts.
+Next, [Discussion] provides further insights and future areas of study.
+Finally, [Conclusion] summarizes our main findings.
+To complete the paper, [Appendix] clarifies how each variable within each dataset is generated and tables to accordingly demonstrate this.
+
+The novel texts used for analysis were sourced from Project Gutenberg under the library `gutenbergr` [@rguten].
+Data was generated, extracted and cleaned using the open-source statistical programming language R [@r], leveraging functions from `tidyverse` [@rTidyverse], `gutenbergr` [@rGuten], `tidytext` [@rTidytext], `rmarkdown` [@rMark], `dplyr` [@rDplyr], `ggplot2` [@rGgplot2], `scales` [@rScales], `here` [@rHere], `wordcloud` [@rCloud], `tm` [@rTm] and `knitr` [@rKnitr].
 
@@
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
 #| echo: false
+#| message: false
+#| label: tbl-genderclass
+#| tbl-cap: Table of Number of Classes Students Considered for Regrade Requests by Students’ Gender
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+classes_dta |>
+  filter(consider_regrade != "No") |>
+  rename("Count" = n, "Number of Classes" = num_class, "Consider Regrade" = consider_regrade) |>
+  head(12) |>
+  kable(
+    col.names = c("Number of Classes", "Gender", "Consider Regrade", "Count", "Percentage"),
+    booktabs = TRUE,
+    align = c("c", "c", "c", "c", "c")
+  )
 ```
@@
 
-analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
+# Visualize word frequencies (e.g., bar plot or word cloud)
+# Example using ggplot2 for bar plot
+top_words <- head(word_freq, 20)
+ggplot(top_words, aes(x = reorder(word, -n), y = n)) +
+  geom_bar(stat = "identity") +
+  labs(x = "Word", y = "Frequency", title = "Top 20 Most Frequent Words")
 
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
 ```
@@
 #| eval: true
+#| label: tbl-frequency
+#| tbl-cap: "Word Frequency of Stream of Consciousness Novels, A Comparison Between Five Authors "
 #| warning: false
-#| message: false
 
@@
                        mutate(dalloway_words, author = "Virginia Woolf"),
-                       mutate(notes_words, author = "Fyodor Dostoevsky")) |>
+                       mutate(bliss_words, author = "Katherine Mansfield"),
+                       mutate(prufrock_words, author = "T.S. Eliot")) |>
   mutate(word = str_extract(word, "[a-z']+")) |>
@@
 
+frequency |>
+  rename("Word" = word) |>
+  head(12) |>
+  kable(
+    col.names = c("Word", "James Joyce", "Katherine Mansfield", "Marcel Proust", "T.S. Eliot", "Virignia Woolf"),
+    booktabs = TRUE,
+    align = c("c", "c", "c", "c")
+  )
+
 ```
@@
 ```{r}
+#| eval: true
 #| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+#| message: false
 #| warning: false
+#| label: fig-frequency
+#| layout-ncol: 2
+#| fig-cap: "Examining how the model fits, and is affected by, the data"
+#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
 
-ggplot(frequency, aes(x = proportion, y = `James Joyce`, 
-                            color = abs(`James Joyce` - proportion))) +
+# Calculate the word frequency over all unique words and bind all frequency rate of authors together to conduct a comparative comparison
+# Code sourced from https://medium.com/the-data-nudge/nlp-basics-exploring-word-frequency-with-the-tidytext-r-package-ac35a6d805f4
+
+frequency_female <- bind_rows(mutate(dalloway_words, author = "Virginia Woolf"),
+                              mutate(portrait_words, author = "Katherine Mansfield")) |>
+  mutate(word = str_extract(word, "[a-z']+")) |>
+  count(author, word) |>
+  group_by(author) |>
+  mutate(proportion = n / sum(n)) |>
+  select(-n) |>
+  pivot_wider(names_from = author, values_from = proportion) |>
+  pivot_longer(`Virginia Woolf`,
+               names_to = "author", values_to = "proportion")
+
+ggplot(frequency_female, aes(x = proportion, y = `Katherine Mansfield`, color = abs(`Katherine Mansfield` - proportion))) +
   geom_abline(color = "gray50", lty = 2) +
@@
   scale_color_gradient(limits = c(0, 0.001), 
-                       low = "lightyellow2", high = "gray80") +
-  facet_wrap(~author, ncol = 4) +
+                       low = "darkslateblue", high = "gray30") +
+  facet_wrap(~author, ncol = 1) +
   theme(legend.position="none") +
-  labs(y = "James Joyce", x = NULL)
+  labs(y = "Katherine Mansfield", x = NULL)
 
-# Visualize word frequencies (e.g., bar plot or word cloud)
-# Example using ggplot2 for bar plot
-top_words <- head(word_freq, 20)
-ggplot(top_words, aes(x = reorder(word, -n), y = n)) +
-  geom_bar(stat = "identity") +
-  labs(x = "Word", y = "Frequency", title = "Top 20 Most Frequent Words")
+frequency_male <- bind_rows(
+                       mutate(swann_words, author = "Marcel Proust"),
+                       mutate(prufrock_words, author = "T.S. Eliot"),
+                       mutate(portrait_words, author = "James Joyce")) |>
+  mutate(word = str_extract(word, "[a-z']+")) |>
+  count(author, word) |>
+  group_by(author) |>
+  mutate(proportion = n / sum(n)) |>
+  select(-n) |>
+  pivot_wider(names_from = author, values_from = proportion) |>
+  pivot_longer(`Marcel Proust`:`James Joyce`,
+               names_to = "author", values_to = "proportion")
+
+ggplot(frequency_male, aes(x = proportion, y = `T.S. Eliot`, color = abs(`T.S. Eliot` - proportion))) +
+  geom_abline(color = "gray50", lty = 2) +
+  geom_jitter(alpha = 0.1, size = 2.5, width = 0.3, height = 0.3) +
+  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
+  scale_x_log10(labels = percent_format()) +
+  scale_y_log10(labels = percent_format()) +
+  scale_color_gradient(limits = c(0, 0.001), 
+                       low = "darkslateblue", high = "gray30") +
+  facet_wrap(~author, ncol = 2) +
+  theme(legend.position="none") +
+  labs(y = "T.S. Eliot", x = NULL)
+
+frequency_britain <- bind_rows(
+                       mutate(dalloway_words, author = "Virginia Woolf"),
+                       mutate(prufrock_words, author = "T.S. Eliot"),
+                       mutate(portrait_words, author = "James Joyce")) |>
+  mutate(word = str_extract(word, "[a-z']+")) |>
+  count(author, word) |>
+  group_by(author) |>
+  mutate(proportion = n / sum(n)) |>
+  select(-n) |>
+  pivot_wider(names_from = author, values_from = proportion) |>
+  pivot_longer(`T.S. Eliot`:`James Joyce`,
+               names_to = "author", values_to = "proportion")
+
+ggplot(frequency_britain, aes(x = proportion, y = `Virginia Woolf`, color = abs(`Virginia Woolf` - proportion))) +
+  geom_abline(color = "gray50", lty = 2) +
+  geom_jitter(alpha = 0.1, size = 2.5, width = 0.3, height = 0.3) +
+  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
+  scale_x_log10(labels = percent_format()) +
+  scale_y_log10(labels = percent_format()) +
+  scale_color_gradient(limits = c(0, 0.001), 
+                       low = "darkslateblue", high = "gray30") +
+  facet_wrap(~author, ncol = 2) +
+  theme(legend.position="none") +
+  labs(y = "Virginia Woolf", x = NULL)
 
@@
 
+# Appendix
+
+### Strip Search and Items Found Correlation
+
+@tbl-reasons-strip-search connotes the summary statistics how many reasons the police force attributes to each person prior to conducting strip searches and the number of items found during each search.
+This table leverages `mutate` to create three new columns on the data's percentages and counts, enabling the data to be proprotionate to one another.
+"Total Reasons" conveys the total number of reasons a strip search has, while "Items Found Percentage" conveys the percentage of items found for each total of reasons.
+"Total Searches by Total Reasons" conveys the total number of searches for each total of reasons, while "Searches Percentage" conveys the percentage of each total of reasons out of all strip searches.
+
+```{r fig.pos="H"}
+#| echo: false
+#| message: false
+#| label: tbl-reasons-strip-search
+#| tbl-cap: Summary Statistics of Reasons for Police Strip Searches
+
+cleaned_search_reasons_data |>
+  group_by(total_reasons, items_found) |>
+  count() |>
+  rename(count = n) |>
+  group_by(total_reasons) |>
+  mutate(percentage = round(count/sum(count)*100),
+         total_searches = sum(count),
+         percentage_total = round(total_searches/7801*100)) |>
+  kable(
+    col.names = c("Total Reasons", "Items Found", "Count", "Items Found Percentage", "Searches by Total Reasons", "Searches Percentage"),
+    booktabs = TRUE,
+    align = c("c", "c", "c", "c", "c", "c")
+  ) |> 
+  kable_styling(latex_options="scale_down")
+
+```
+
+### Strip Search and Items Found Correlation
+
+@tbl-items-strip-search conveys the total number of strip searches happening in Toronto from 2020 to 2021, and whether or not items were found during each search.
+
+```{r fig.pos="H"}
+#| echo: false
+#| message: false
+#| label: tbl-items-strip-search
+#| tbl-cap: Summary Statistics of Items Found During Police Strip Searches
+
+cleaned_race_gender_data |>
+  group_by(strip_search, items_found) |>
+  count() |>
+  rename(count = n) |>
+  filter(strip_search == "Yes") |>
+  select(items_found, count) |>
+  group_by(strip_search) |>
+  mutate(total_searches = sum(count),
+         percentage_of_items_found = round(count/total_searches*100)) |>
+  kable(col.names = c("Strip Search", "Items Found", "Count", "Total Searches", "Percentage of Items Found"), booktabs = TRUE, align = c("c", "c", "c", "c", "c")
+        )
+```
+
 # References

### 2024-04-03 scripts/01-download_data.R

@@ -1,12 +1,10 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Downloads and saves the novel texts from Project Gutenberg (https://www.gutenberg.org_
+# Author: Quang Mai
+# Date: 29 March 2024 
+# Contact: quangpono@gmail.com
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Pre-requisites: Install and load packages tidytext and gutenbergr
 
 #### Workspace setup ####
 library(tidyverse)
@@ -29,38 +27,48 @@ joyce_books <- all_books %>%
 woolf_books <- all_books %>%
   filter(str_detect(author, 'Woolf, Virginia'))
 
-dostoevsky_books <- all_books %>%
-  filter(str_detect(author, 'Dostoyevsky, Fyodor'))
+mansfield_books <- all_books %>%
+  filter(str_detect(author, 'Mansfield, Katherine'))
 
 proust_books <- all_books %>%
   filter(str_detect(author, 'Proust, Marcel'))
 
-combined_books <- rbind(joyce_books, woolf_books, dostoevsky_books, proust_books)
+ts_books <- all_books %>%
+  filter(str_detect(author, 'Eliot, T. S.'))
 
 
-# Download the text of the most stream-of-consciousness works of each author:
 
-portrait <- gutenberg_download(c('4217')) |>
-  mutate(book = ifelse(gutenberg_id == 4217, "A Portrait of the Artist as a Young Man", "NA")) |>
+combined_books <- rbind(joyce_books, woolf_books, dostoevsky_books, proust_books, ts_books, mansf)
+
+
+# Download the more famous stream-of-consciousness work of each author:
+
+portrait <- gutenberg_download(c('4217', '2817')) |>
+  mutate(book = ifelse(gutenberg_id == 4217, "A Portrait of the Artist as a Young Man", "Chamber Music")) |>
   slice(-(1:20))
 
 swann_way <- gutenberg_download(c('7178')) |>
[…]

### 2024-04-03 scripts/02-data_cleaning.R

@@ -4,7 +4,7 @@
 # Date: 30 March 2024 
 # Contact: quangpono@gmail.com
 # License: MIT
-# Pre-requisites: Read four raw datasets containing the four novels
+# Pre-requisites: Read four raw datasets containing the novel texts from five authors
 
 #### Workspace setup ####
 library(tidyverse)
@@ -21,7 +21,9 @@ library(scales)
 portrait <- read_csv("data/raw_data/portrait.csv")
 swann <- read_csv("data/raw_data/swann_way.csv")
 dalloway <- read_csv("data/raw_data/mrs_dalloway.csv")
-notes <- read_csv("data/raw_data/notes_underground.csv")
+bliss <- read_csv("data/raw_data/bliss.csv")
+prufrock <- read_csv("data/raw_data/prufrock.csv")
+
 
 # Tokenize the words from the text, breaking the text down into individual words to insert them into individual datasets
 # Codes sourced from https://medium.com/the-data-nudge/nlp-basics-exploring-word-frequency-with-the-tidytext-r-package-ac35a6d805f4
@@ -38,24 +40,36 @@ dalloway_tokenized <- dalloway |>
   na.omit(dalloway) |>
   unnest_tokens(word, text)
 
-notes_tokenized <- notes |>
-  na.omit(notes) |>
+prufrock_tokenized <- prufrock |>
+  na.omit(prufrock) |>
+  unnest_tokens(word, text)
+
+bliss_tokenized <- bliss |>
+  na.omit(bliss) |>
   unnest_tokens(word, text)
 
 # Exclude stop word using the tidytext stop_words dataset, also, use anti_join() to specify the name of the dataset with words that need to be excluded
 data(stop_words)
 
 portrait_clean <- portrait_tokenized |>
-  anti_join(stop_words)
+  anti_join(stop_words) |>
+  filter(nchar(word) > 1)
 
 swann_clean <- swann_tokenized |>
-  anti_join(stop_words)
+  anti_join(stop_words) |>
+  filter(nchar(word) > 1)
 
 dalloway_clean <- dalloway_tokenized |>
-  anti_join(stop_words)
+  anti_join(stop_words) |>
+  filter(nchar(word) > 1)
 
-notes_clean <- notes_tokenized |>
-  anti_join(stop_words)
+bliss_clean <- bliss_tokenized |>
+  anti_join(stop_words) |>
+  filter(nchar(word) > 1)
+
[…]

### 2024-04-03 paper/paper.qmd
structure paper & refine writing and data cleaning

@@
 By analyzing the word frequency of their famous works, I explore mental health themes like anxiety, depression, trauma, and existential angst.
-Through word frequency analysis of nine novels, I aim to uncover shared linguistic patterns and gain insights into the authors' mental states, offering a glimpse into themes of self-identity and existential contemplation"
+Through word frequency analysis of nine novels, I aim to uncover shared linguistic patterns and gain insights into the authors' mental states, offering a glimpse into themes of self-identity and existential contemplation."
 format: pdf
@@
 prufrock_words <- read_csv(here::here("data/analysis_data/prufrock_words.csv"))
+combined_trigram <- read_csv(here::here("data/analysis_data/combined_trigram.csv"))
+combined_bigram <- read_csv(here::here("data/analysis_data/combined_bigram.csv"))
+combined_text <- read_csv(here::here("data/analysis_data/combined_text.csv"))
 
@@
 
-You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.
+We use [@flow] [@pathology] [@resampling] and @rohan.
 
@@
 
-A Portrait of the Artist as a Young Man, Dubliners, Mrs Dalloway, Jacob's Room, Swann Way, Bliss, The Garden Story, The Waste Land and The Love Song of J. Alfred Prufrock
+By analyzing the datasets of these novel texsts, I seek to pose and answer crucial questions: What are some important factors contributing to this relationship between mental health, disassociation and stream of consciousness?
+Moreover, how does this relationship vary differently across differnet demographics of authors, for instance, authors with different geographical locations and genders?
+Understanding these dynamics is crucial in having an informed understanding of the West's late 19th to mid-20th century socio-political landscape, especially in regards to how authors and creative writers navigate and deal with then-taboo topics such as existential angst, mental health issues and disabilities.
 
-By analyzing the datasets, I seek to pose and answer crucial questions: What are some important factors contributing to this trend?
-How does childlessness manifest differently across demographic groups?
-Finally, what are the potential social, economic, and individual consequences of this new family structure in the United States?
-Understanding these dynamics is crucial in having an informed policy discussion and navigating the implications of childlessness in a rapidly changing world.
-
-Thus, the estimand is the correlation between stream of consciousness authors and words related to mental health themes appearing in their famous works.
-This is considered in terms of those who would have a third birth only if the first two children were of the same sex.
-Through our analysis, we found that the childless rate is consistently on the rise from 1972 to 2022, with a correlation coefficient of 0.73 between the childless percentage and time (or a 6.99% increase over the last 50 years).
-Factors that directly contribute to this rise are higher educational attainment, higher socioeconomic class, lower gender disparity, better health status and dissolution of marriages through divorces, separation and being widowed.
-In addition, global events like pandemics and recessions also heighten the rate of childlessness from 1972-2022, including the AIDs pandemic, 9/11, 2008 Recession and the COVID-19 pandemic.
-During the 2008 Recession, from 2008 to 2014, the rate of childlessness increases the most, at 3%.
+Thus, my estimand is the correlation between famous stream of consciousness literature by Joyce, Woolf, Proust, Mansfield and Eliot, words related to mental health themes and their frequency.
+This is considered in terms of the nine selected novels only, namely Joyce's A Portrait of the Artist as a Young Man and Chamber Music; Woolf's Mrs Dalloway and Jacob's Room; Proust's Swann Way; Mansfield's Bliss and The Garden Party; and Eliot's The Waste Land and The Love Song of J.
+Alfred Prufrock.
+Through our analysis, we found that (percentage, number and data here, main results)...
 
@@
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+## Measurement
+
+## Source Data
+
+## Data Cleaning and Word Tokenization
 
@@
 
-classes_dta |>
-  filter(consider_regrade != "No") |>
-  rename("Count" = n, "Number of Classes" = num_class, "Consider Regrade" = consider_regrade) |>
-  head(12) |>
-  kable(
-    col.names = c("Number of Classes", "Gender", "Consider Regrade", "Count", "Percentage"),
-    booktabs = TRUE,
-    align = c("c", "c", "c", "c", "c")
-  )
 ```
-
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
 
@@
 
-# Visualize word frequencies (e.g., bar plot or word cloud)
-# Example using ggplot2 for bar plot
-top_words <- head(word_freq, 20)
-ggplot(top_words, aes(x = reorder(word, -n), y = n)) +
-  geom_bar(stat = "identity") +
-  labs(x = "Word", y = "Frequency", title = "Top 20 Most Frequent Words")
+```
+
+### Word Count
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-joyce
+#| tbl-cap: "Examplary Table of Unprocessed Stream of Consciousness Novels (James Joyce)"
+#| warning: false
+
+portrait |>
+  rename("Book ID" = gutenberg_id,
+         "Text" = text,
+         "Book" = book) |>
+  tail(10) |>
+  kable(
+    col.names = c("Book ID", "Text", "Book"),
+    booktabs = TRUE
+  )
 
@@
 
-Talk way more about it.
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-joyce-token
+#| tbl-cap: "Examplary Table of Tokenized Stream of Consciousness Novels (James Joyce)"
+#| warning: false
 
-# Model
+portrait_clean |>
+  rename("Book ID" = gutenberg_id,
+         "Book" = book,
+         "Word" = word) |>
+  tail(10) |>
+  kable(
+    col.names = c("Book ID", "Book", "Word"),
+    booktabs = TRUE
+  )
 
-The goal of our modelling strategy is twofold. Firstly,...
+```
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-joyce-words
+#| tbl-cap: "Examplary Table of Word Count of Stream of Consciousness Novels (James Joyce)"
+#| warning: false
 
-## Model set-up
+portrait_words |>
+  rename("Word" = word) |>
+  head(10) |>
+  kable(
+    col.names = c("Word", "Count"),
+    booktabs = TRUE
+  )
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+```
 
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
-# Results
-
-Our results are summarized in @tbl-modelresults.
+### Comparative Word Frequency
 
@@
     col.names = c("Word", "James Joyce", "Katherine Mansfield", "Marcel Proust", "T.S. Eliot", "Virignia Woolf"),
-    booktabs = TRUE,
-    align = c("c", "c", "c", "c")
+    booktabs = TRUE
   )
@@
 ```
+
+# Model
+
+## Model set-up
+
+### Model justification
+
+# Results
+
+## The Dominant Vocabulary of Stream of Consciousness
 
@@
 #| warning: false
-#| label: fig-frequency
+#| label: fig-count
 #| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
+#| fig-cap: "Comparative Analysis of Top 20 Word Frequencies by Stream of Consciousness Authors' Novels"
+#| fig-subcap: ["James Joyce", "Virginia Woolf", "Marcel Proust", "T.S. Eliot", "Katherine Mansfield"]
+
+portrait_words |> head(20) |>
+ggplot(aes(x = n, y = reorder(word, n))) +
+  geom_bar(stat = "identity", fill = "darkgrey") +
+  labs(x = "Frequency", y = "Word")
+
+dalloway_words |> head(20) |>
+ggplot(aes(x = n, y = reorder(word, n))) +
+  geom_bar(stat = "identity", fill = "darkgrey") +
+  labs(x = "Frequency", y = "Word")
+
+swann_words |> head(20) |>
+ggplot(aes(x = n, y = reorder(word, n))) +
+  geom_bar(stat = "identity", fill = "darkgrey") +
+  labs(x = "Frequency", y = "Word")
+
+prufrock_words |> head(20) |>
+ggplot(aes(x = n, y = reorder(word, n))) +
+  geom_bar(stat = "identity", fill = "darkgrey") +
+  labs(x = "Frequency", y = "Word")
+
+bliss_words |> head(20) |>
+ggplot(aes(x = n, y = reorder(word, n))) +
+  geom_bar(stat = "identity", fill = "darkgrey") +
+  labs(x = "Frequency", y = "Word")
+
+```
+
+## Great Britain as the Center of Stream of Consciousness Literature
+
+```{r}
+#| eval: true
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-frequency-brit
+#| fig-cap: "Comparative Analysis of Word Frequency in British Stream of Consciousness Authors"
+
+frequency_britain <- bind_rows(
+                       mutate(dalloway_words, author = "Virginia Woolf"),
+                       mutate(prufrock_words, author = "T.S. Eliot"),
+                       mutate(portrait_words, author = "James Joyce")) |>
+  mutate(word = str_extract(word, "[a-z']+")) |>
+  count(author, word) |>
+  group_by(author) |>
+  mutate(proportion = n / sum(n)) |>
+  select(-n) |>
+  pivot_wider(names_from = author, values_from = proportion) |>
+  pivot_longer(`T.S. Eliot`:`James Joyce`,
+               names_to = "author", values_to = "proportion")
+
+ggplot(frequency_britain, aes(x = proportion, y = `Virginia Woolf`, color = abs(`Virginia Woolf` - proportion))) +
+  geom_abline(color = "gray50", lty = 2) +
+  geom_jitter(alpha = 0.1, size = 2.5, width = 0.3, height = 0.3) +
+  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
+  scale_x_log10(labels = percent_format()) +
+  scale_y_log10(labels = percent_format()) +
+  scale_color_gradient(limits = c(0, 0.001), 
+                       low = "darkslateblue", high = "gray1") +
+  facet_wrap(~author, ncol = 2) +
+  theme(legend.position="none") +
+  labs(y = "Virginia Woolf", x = NULL)
+
+```
+
+## Gendered Mental Landscape of Stream of Consciousness Novels
+
+### Comparing Female Stream of Consciousness Authors
+
+```{r}
+#| eval: true
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-frequency-female
+#| fig-cap: "Comparative Analysis of Word Frequency in Female Stream of Consciousness Authors"
 
@@
 frequency_female <- bind_rows(mutate(dalloway_words, author = "Virginia Woolf"),
-                              mutate(portrait_words, author = "Katherine Mansfield")) |>
+                              mutate(bliss_words, author = "Katherine Mansfield")) |>
   mutate(word = str_extract(word, "[a-z']+")) |>
@@
   scale_color_gradient(limits = c(0, 0.001), 
-                       low = "darkslateblue", high = "gray30") +
+                       low = "darkslateblue", high = "gray50") +
   facet_wrap(~author, ncol = 1) +
@@
   labs(y = "Katherine Mansfield", x = NULL)
+
+```
+
+### Comparing Male Stream of Consciousness Authors
+
+```{r}
+#| eval: true
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-frequency-male
+#| fig-cap: "Comparative Analysis of Word Frequency in Male Stream of Consciousness Authors"
 
@@
   scale_color_gradient(limits = c(0, 0.001), 
-                       low = "darkslateblue", high = "gray30") +
+                       low = "darkslateblue", high = "gray100") +
   facet_wrap(~author, ncol = 2) +
@@
 
-frequency_britain <- bind_rows(
-                       mutate(dalloway_words, author = "Virginia Woolf"),
-                       mutate(prufrock_words, author = "T.S. Eliot"),
+```
+
+## Transnational Stream of Consciousness Novels and Mental Health Themes
+
+```{r}
+#| eval: true
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-frequency-int
+#| fig-cap: "Comparative Analysis of Word Frequency in Transnational Stream of Consciousness Authors"
+
+frequency_international <- bind_rows(
+                       mutate(swann_words, author = "Marcel Proust"),
+                       mutate(bliss_words, author = "Katherine Mansfield"),
                        mutate(portrait_words, author = "James Joyce")) |>
@@
   pivot_wider(names_from = author, values_from = proportion) |>
-  pivot_longer(`T.S. Eliot`:`James Joyce`,
+  pivot_longer(`Katherine Mansfield`:`James Joyce`,
                names_to = "author", values_to = "proportion")
 
-ggplot(frequency_britain, aes(x = proportion, y = `Virginia Woolf`, color = abs(`Virginia Woolf` - proportion))) +
+ggplot(frequency_international, aes(x = proportion, y = `Marcel Proust`, color = abs(`Marcel Proust` - proportion))) +
   geom_abline(color = "gray50", lty = 2) +
@@
   scale_color_gradient(limits = c(0, 0.001), 
-                       low = "darkslateblue", high = "gray30") +
-  facet_wrap(~author, ncol = 2) +
+                       low = "darkslateblue", high = "gray100") +
+  facet_wrap(~author, ncol = 1) +
   theme(legend.position="none") +
-  labs(y = "Virginia Woolf", x = NULL)
+  labs(y = "Marcel Proust", x = NULL)
+
+```
+
+## Combined Texts: Trends, Bigram and Trigram Analsis
+
+```{r}
+#| eval: true
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-cloud
+#| layout-ncol: 2
+#| fig-cap: "Visualizations of Word Frequency when Combining Stream of Consciousness Authors' Novels"
+#| fig-subcap: ["One Word"] #Add "Two Words" and "Three Words" subsequently when uncovering the trends for the previous one
+
+filtered_text <- combined_text[combined_text$n >= 2, ] #only select words with a frequency larger or equal to 2
+
+wordcloud(words = filtered_text$word, freq = filtered_text$n, scale=c(3,0.5), min.freq = 2, random.order = FALSE, colors=brewer.pal(8, "Dark2"))
+
+#filtered_bigram <- combined_bigram[combined_bigram$n >= 2, ]
+#wordcloud(words = filtered_bigram$word, freq = filtered_bigram$n, scale=c(3,0.5), min.freq = 2, random.order = #FALSE, colors=brewer.pal(8, "Dark2"))
+
+#filtered_trigram <- combined_trigram[combined_trigram$n >= 2, ]
+#wordcloud(words = filtered_trigram$word, freq = filtered_trigram$n, scale=c(3,0.5), min.freq = 2, random.order #= FALSE, colors=brewer.pal(8, "Dark2"))
 
@@
 
-## First discussion point {#sec-first-point}
+## Mental Health Vocabulary: Patterns and Trends
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+Discuss vocabulary patterns and word trends
 
-## Second discussion point
+## Insights into Socio-Political Landscape of the West's Modernist Era
 
-## Third discussion point
+The novels' linguistic patterns reflect the socio-political landscape of the Western hemisphere, from  late 19th century to the mid-20th century.
 
-## Weaknesses and next steps
+## Schizophrenic and Disassociative Tendencies in Female Stream of Consciousness
 
-Weaknesses and next steps should also be included.
+## Weaknesses
+
+### Lack of Thorough Word Cleaning
+
+### Limiting Literature Works and Publication Editions
+
+### Project Gutenberg's Focus on the Canon
+
+## Moving Forward and Next Steps
 
@@
 
-\appendix
+# Appendix
 
-# Appendix {-}
+## Additional Data Details
 
-# Additional data details
+```{r}
+#| eval: true
+#| echo: false
+#| message: false
+#| warning: false
+combined_books <- rbind(joyce_books, woolf_books, dostoevsky_books, proust_books, ts_books, mansfield_books)
+```
 
-# Model details {#sec-model-details}
+### Data Gathering
+
+```{r fig.pos="H"}
+#| echo: false
+#| message: false
+#| label: tbl-reasons-strip-search
+#| tbl-cap: Summary Statistics of Reasons for Police Strip Searches
+
+```
+
+### Data Cleaning
+
+```{r fig.pos="H"}
+#| echo: false
+#| message: false
+#| label: tbl-items-strip-search
+#| tbl-cap: Summary Statistics of Items Found During Police Strip Searches
+
+```
+
+## Model Details {#sec-model-details}
 
 ## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
 
@@
 
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
 ```
@@
 ## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
 
@@
 
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
 ```
@@
 
-# Appendix
-
-### Strip Search and Items Found Correlation
-
-@tbl-reasons-strip-search connotes the summary statistics how many reasons the police force attributes to each person prior to conducting strip searches and the number of items found during each search.
-This table leverages `mutate` to create three new columns on the data's percentages and counts, enabling the data to be proprotionate to one another.
-"Total Reasons" conveys the total number of reasons a strip search has, while "Items Found Percentage" conveys the percentage of items found for each total of reasons.
-"Total Searches by Total Reasons" conveys the total number of searches for each total of reasons, while "Searches Percentage" conveys the percentage of each total of reasons out of all strip searches.
-
-```{r fig.pos="H"}
-#| echo: false
-#| message: false
-#| label: tbl-reasons-strip-search
-#| tbl-cap: Summary Statistics of Reasons for Police Strip Searches
-
-cleaned_search_reasons_data |>
-  group_by(total_reasons, items_found) |>
-  count() |>
-  rename(count = n) |>
-  group_by(total_reasons) |>
-  mutate(percentage = round(count/sum(count)*100),
-         total_searches = sum(count),
-         percentage_total = round(total_searches/7801*100)) |>
-  kable(
-    col.names = c("Total Reasons", "Items Found", "Count", "Items Found Percentage", "Searches by Total Reasons", "Searches Percentage"),
-    booktabs = TRUE,
-    align = c("c", "c", "c", "c", "c", "c")
-  ) |> 
-  kable_styling(latex_options="scale_down")
-
-```
-
-### Strip Search and Items Found Correlation
-
-@tbl-items-strip-search conveys the total number of strip searches happening in Toronto from 2020 to 2021, and whether or not items were found during each search.
-
-```{r fig.pos="H"}
-#| echo: false
-#| message: false
-#| label: tbl-items-strip-search
-#| tbl-cap: Summary Statistics of Items Found During Police Strip Searches
-
-cleaned_race_gender_data |>
-  group_by(strip_search, items_found) |>
-  count() |>
-  rename(count = n) |>
-  filter(strip_search == "Yes") |>
-  select(items_found, count) |>
-  group_by(strip_search) |>
-  mutate(total_searches = sum(count),
-         percentage_of_items_found = round(count/total_searches*100)) |>
-  kable(col.names = c("Strip Search", "Items Found", "Count", "Total Searches", "Percentage of Items Found"), booktabs = TRUE, align = c("c", "c", "c", "c", "c")
-        )
-```
-
 # References

### 2024-04-03 scripts/01-download_data.R

@@ -38,7 +38,7 @@ ts_books <- all_books %>%
 
 
 
-combined_books <- rbind(joyce_books, woolf_books, dostoevsky_books, proust_books, ts_books, mansf)
+combined_books <- rbind(joyce_books, woolf_books, dostoevsky_books, proust_books, ts_books, mansfield_books)
 
 
 # Download the more famous stream-of-consciousness work of each author:

### 2024-04-03 scripts/02-data_cleaning.R

@@ -87,6 +87,31 @@ bliss_words <- bliss_clean |>
 prufrock_words <- prufrock_clean |>
   count(word, sort = TRUE)
 
+# Combine all texts to find the more general trend of word frequency
+
+## Bigram analysis or analysis by prepositioned words, codes sourced from https://www.youtube.com/watch?v=9T-hr3jinTw
+combined_trigram <-
+  rbind(portrait, swann, dalloway, prufrock, bliss)|>
+  na.omit(portrait, swann, dalloway, prufrock, bliss) |>
+  unnest_tokens(bigram, text, token = "ngrams", n = 3) |>
+  separate(bigram, c("word1", "word2", "word3"), sep = " ") #|>
+  #count(word1, word2, word3, sort = TRUE)
+
+combined_bigram <-
+  rbind(portrait, swann, dalloway, prufrock, bliss)|>
+  na.omit(portrait, swann, dalloway, prufrock, bliss) |>
+  unnest_tokens(bigram, text, token = "ngrams", n = 2) |>
+  separate(bigram, c("word1", "word2"), sep = " ") #|>
+  #count(word1, word2, sort = TRUE)
+
+combined_text <- prufrock |>
+  rbind(portrait, swann, dalloway, prufrock, bliss)|>
+  na.omit(portrait, swann, dalloway, prufrock, bliss) |>
+  unnest_tokens(word, text) |>
+  anti_join(stop_words) |>
+  filter(nchar(word) > 1) |>
+  count(word, sort = TRUE)
+
 #### Save data ####
 write_csv(portrait_clean, "data/analysis_data/portrait_clean.csv")
 write_csv(portrait_words, "data/analysis_data/portrait_words.csv")
@@ -94,4 +119,11 @@ write_csv(swann_words, "data/analysis_data/swann_words.csv")
 write_csv(dalloway_words, "data/analysis_data/dalloway_words.csv")
 write_csv(bliss_words, "data/analysis_data/bliss_words.csv")
 write_csv(prufrock_words, "data/analysis_data/prufrock_words.csv")
+write_csv(combined_text, "data/analysis_data/combined_text.csv")
+write_csv(combined_bigram, "data/analysis_data/combined_bigram.csv")
+write_csv(combined_trigram, "data/analysis_data/combined_trigram.csv")
+
+
+
+
 

### 2024-04-03 paper/paper.qmd
render paper

@@
 
-We use [@flow] [@pathology] [@resampling] and @rohan.
+Stream of consciousness is a narrative technique that aims to capture the continuous flow of thoughts, feelings, and sensations experienced by a character without conventional organization or punctuation  [@resampling].
+It mirrors the unpredictable and interconnected nature of human thought processes, often revealing the inner workings of the character's mind in an intimate and unfiltered manner [@flow].
+In literature, most scholars agree that stream of consciousness reveals the complexities of mental-scapes, shedding light on the nuances of characters' emotional well-ebing and psychological struggles [@pathology].
+Its relationship to mental health lies in its ability to offer insight into the depths of consciousness, providing a platform to explore themes such as anxiety, trauma, and existential questioning.
+As such, this paper has mined the texts of a total of nine novels from the volunteer archive, Project Gutenberg, to examine the mental health themes of famous stream of consciousness authors, namely by Joyce, Woolf, Proust, Mansfield and Eliot, from the modernist era of literature, spanning from late 19th century to the mid-20th century [@gutenberg].
+(more stats and data mentioned here)
 
-The American family landscape is undergoing a significant transformation, marked by a growing trend of childlessness.
-Traditionally viewed as a social norm, having children now becomes a choice [@voluntaryChild].
-We examine factors such as age, gender, education, socioeconomic class, health conditions and marital status, and how they affect childbearing rates from 1972 to 2022.
-Existing research suggests a complex interplay of factors that contribute to childlessness, including economic considerations, rising educational attainment, shifting social norms related to career and family life, and individual preferences [@voluntaryChild].
-
-Texts from a total of nine novels were retrieved from the volunteer archive to examine the mental health themes of famous stream of consciousness authors from the Modernst Era of literature from late 19th century to the mid-20th century
-
-By analyzing the datasets of these novel texsts, I seek to pose and answer crucial questions: What are some important factors contributing to this relationship between mental health, disassociation and stream of consciousness?
+By analyzing these textual datasets, I seek to pose and answer crucial questions: What are some important factors contributing to this relationship between mental health, disassociation and stream of consciousness?
 Moreover, how does this relationship vary differently across differnet demographics of authors, for instance, authors with different geographical locations and genders?
@@
 
-Thus, my estimand is the correlation between famous stream of consciousness literature by Joyce, Woolf, Proust, Mansfield and Eliot, words related to mental health themes and their frequency.
+Thus, my estimand is the correlation between famous stream of consciousness novels, words related to mental health themes and their frequency.
 This is considered in terms of the nine selected novels only, namely Joyce's A Portrait of the Artist as a Young Man and Chamber Music; Woolf's Mrs Dalloway and Jacob's Room; Proust's Swann Way; Mansfield's Bliss and The Garden Party; and Eliot's The Waste Land and The Love Song of J.
@@
 
-The novel texts used for analysis were sourced from Project Gutenberg under the library `gutenbergr` [@rguten].
-Data was generated, extracted and cleaned using the open-source statistical programming language R [@r], leveraging functions from `tidyverse` [@rTidyverse], `gutenbergr` [@rGuten], `tidytext` [@rTidytext], `rmarkdown` [@rMark], `dplyr` [@rDplyr], `ggplot2` [@rGgplot2], `scales` [@rScales], `here` [@rHere], `wordcloud` [@rCloud], `tm` [@rTm] and `knitr` [@rKnitr].
+The novel texts used for analysis were sourced from Project Gutenberg under the library `gutenbergr` [@rGuten] [@rgutenberg].
+Data was generated, extracted and cleaned using the open-source statistical programming language R [@r], leveraging functions from `tidyverse` [@rTidyverse], `tidytext` [@rTidytext], `rmarkdown` [@rMark], `dplyr` [@rDplyr], `ggplot2` [@rGgplot2], `scales` [@rScales], `here` [@rHere], `wordcloud` [@rCloud], `tm` [@rTm] and `knitr` [@rKnitr].
 
@@
   geom_abline(color = "gray50", lty = 2) +
-  geom_jitter(alpha = 0.1, size = 2.5, width = 0.3, height = 0.3) +
+  geom_jitter(alpha = 0.05, size = 1.5, width = 0.7, height = 0.7) +  
   geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
@@
   geom_abline(color = "gray50", lty = 2) +
-  geom_jitter(alpha = 0.1, size = 2.5, width = 0.3, height = 0.3) +
+  geom_jitter(alpha = 0.1, size = 1, width = 0.5, height = 0.5) +  
   geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
-  scale_x_log10(labels = percent_format()) +
-  scale_y_log10(labels = percent_format()) +
+  scale_x_log10(labels = scales::percent_format()) +
+  scale_y_log10(labels = scales::percent_format()) +
   scale_color_gradient(limits = c(0, 0.001), 
@@
   geom_abline(color = "gray50", lty = 2) +
-  geom_jitter(alpha = 0.1, size = 2.5, width = 0.3, height = 0.3) +
+  geom_jitter(alpha = 0.1, size = 1.5, width = 0.8, height = 0.8) +  
   geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
@@
   geom_abline(color = "gray50", lty = 2) +
-  geom_jitter(alpha = 0.1, size = 2.5, width = 0.3, height = 0.3) +
+  geom_jitter(alpha = 0.05, size = 1, width = 0.8, height = 0.8) +  
   geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
@@
 #| layout-ncol: 2
-#| fig-cap: "Visualizations of Word Frequency when Combining Stream of Consciousness Authors' Novels"
+#| fig-cap: "Visualizations of Word Frequency when Combining All Stream of Consciousness Novels"
 #| fig-subcap: ["One Word"] #Add "Two Words" and "Three Words" subsequently when uncovering the trends for the previous one
 
-filtered_text <- combined_text[combined_text$n >= 2, ] #only select words with a frequency larger or equal to 2
+filtered_text <- combined_text[combined_text$n >= 300, ] #only select words with a frequency larger or equal to 300
 
-wordcloud(words = filtered_text$word, freq = filtered_text$n, scale=c(3,0.5), min.freq = 2, random.order = FALSE, colors=brewer.pal(8, "Dark2"))
+wordcloud(words = filtered_text$word, freq = filtered_text$n, scale=c(3,0.5), min.freq = 2, random.order = FALSE, colors= colorRampPalette(c("lightgray", "black"))(9))
 
@@
 #| warning: false
-combined_books <- rbind(joyce_books, woolf_books, dostoevsky_books, proust_books, ts_books, mansfield_books)
+#combined_books 
 ```

### 2024-04-03 paper/paper.qmd
paper update

@@
 It mirrors the unpredictable and interconnected nature of human thought processes, often revealing the inner workings of the character's mind in an intimate and unfiltered manner [@flow].
-In literature, most scholars agree that stream of consciousness reveals the complexities of mental-scapes, shedding light on the nuances of characters' emotional well-ebing and psychological struggles [@pathology].
+In literature, most scholars agree that stream of consciousness reveals the complexities of mental-scapes, shedding light on the nuances of characters' emotional well-being and psychological struggles [@pathology].
 Its relationship to mental health lies in its ability to offer insight into the depths of consciousness, providing a platform to explore themes such as anxiety, trauma, and existential questioning.

### 2024-04-03 paper/paper.qmd
update graphs

@@
 library(knitr)
-library(wordcloud)
+library(igraph)
+library(widyr)
+library(ggraph)
 
@@
 prufrock_words <- read_csv(here::here("data/analysis_data/prufrock_words.csv"))
-combined_trigram <- read_csv(here::here("data/analysis_data/combined_trigram.csv"))
-combined_bigram <- read_csv(here::here("data/analysis_data/combined_bigram.csv"))
 combined_text <- read_csv(here::here("data/analysis_data/combined_text.csv"))
@@
 
-The novel texts used for analysis were sourced from Project Gutenberg under the library `gutenbergr` [@rGuten] [@rgutenberg].
-Data was generated, extracted and cleaned using the open-source statistical programming language R [@r], leveraging functions from `tidyverse` [@rTidyverse], `tidytext` [@rTidytext], `rmarkdown` [@rMark], `dplyr` [@rDplyr], `ggplot2` [@rGgplot2], `scales` [@rScales], `here` [@rHere], `wordcloud` [@rCloud], `tm` [@rTm] and `knitr` [@rKnitr].
+The novel texts used for analysis were sourced from Project Gutenberg under the library `gutenbergr` [@rGuten] [@gutenberg].
+Data was generated, extracted and cleaned using the open-source statistical programming language R [@r], leveraging functions from `tidyverse` [@rTidyverse], `tidytext` [@rTidytext], `rmarkdown` [@rMark], `dplyr` [@rDplyr], `ggplot2` [@rGgplot2], `scales` [@rScales], `here` [@rHere], `igraph` [@rIgraph], `widyr` [@rWidy], `ggraph` [@rGg] `tm` [@rTm] and `knitr` [@rKnitr].
 
@@
          "Book" = book) |>
-  tail(10) |>
+  head(10) |>
   kable(
@@
 ```
+
+### Generating Word Networks
 
@@
 
-### Comparing Female Stream of Consciousness Authors
-
 ```{r}
@@
 #| warning: false
-#| label: fig-frequency-female
-#| fig-cap: "Comparative Analysis of Word Frequency in Female Stream of Consciousness Authors"
+#| label: fig-word-gender
+#| layout-ncol: 2
+#| fig-cap: "Comparative Analysis of Word Frequency in Female and Male Stream of Consciousness Authors"
+#| fig-subcap: ["Female Authors", "Male Authors"]
 
@@
 
-```
-
-### Comparing Male Stream of Consciousness Authors
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-frequency-male
-#| fig-cap: "Comparative Analysis of Word Frequency in Male Stream of Consciousness Authors"
-
 frequency_male <- bind_rows(
@@
 ```
+
+### Comparing Female Stream of Consciousness Authors
+
+### Comparing Male Stream of Consciousness Authors
 
@@
 
-## Combined Texts: Trends, Bigram and Trigram Analsis
+## Combined Texts: Trends, Word Networks, Bigram and Trigram Analsis
 
@@
 #| warning: false
-#| label: fig-cloud
-#| layout-ncol: 2
-#| fig-cap: "Visualizations of Word Frequency when Combining All Stream of Consciousness Novels"
-#| fig-subcap: ["One Word"] #Add "Two Words" and "Three Words" subsequently when uncovering the trends for the previous one
 
-filtered_text <- combined_text[combined_text$n >= 300, ] #only select words with a frequency larger or equal to 300
+## Code sourced from https://www.youtube.com/watch?v=ae_XVhjHd_o
 
-wordcloud(words = filtered_text$word, freq = filtered_text$n, scale=c(3,0.5), min.freq = 2, random.order = FALSE, colors= colorRampPalette(c("lightgray", "black"))(9))
+generate_word_graph <- function(combined_text,
+                                minimum_n = 100,
+                                minimum_correlation = 0.2) {
 
-#filtered_bigram <- combined_bigram[combined_bigram$n >= 2, ]
-#wordcloud(words = filtered_bigram$word, freq = filtered_bigram$n, scale=c(3,0.5), min.freq = 2, random.order = #FALSE, colors=brewer.pal(8, "Dark2"))
+combined_text_count <- combined_text |>
+  count(word, sort = TRUE) |>
+  filter(n >= 200)
 
-#filtered_trigram <- combined_trigram[combined_trigram$n >= 2, ]
-#wordcloud(words = filtered_trigram$word, freq = filtered_trigram$n, scale=c(3,0.5), min.freq = 2, random.order #= FALSE, colors=brewer.pal(8, "Dark2"))
+combined_correlations <- combined_text |>
+  semi_join(combined_text_count, by = "word") |>
+  pairwise_cor(item = word, feature = book) |>
+  filter(correlation >= 0.2)
+
+graph_from_data_frame(d = combined_correlations, #create graphs of networks of words
+                      vertices = combined_text_count |>
+                      semi_join(combined_correlations, by = c("word" = "item1"))) |>
+  ggraph(layout = "fr") +
+  geom_edge_link(aes(alpha = correlation), alpha = 0.05) +
+  geom_node_point() +
+  geom_node_text(aes(color = n, label = name), repel = TRUE)
+
+}
+
+# Code sourced from
+# filtered_text <- combined_text[combined_text$n >= 100, ] #only select words with a frequency larger or equal to 100
+
+# wordcloud(words = filtered_text$word, freq = filtered_text$n, scale=c(3,0.5), min.freq = 2, random.order = FALSE, colors= colorRampPalette(c("lightgray", "black"))(9))
 
@@
 
+```{r}
+#| eval: true
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-word-networks
+#| layout-ncol: 2
+#| fig-cap: "Word Networks Measured by Frequency and Correlation when Combining All Stream of Consciousness Novels"
+#| fig-subcap: ["100 Frequency, 0.2 Correlation", "200 Frequency, 0.4 Correlation", "200 Frequency, 0.8 Correlation"] #Add "Two Words" and "Three Words" subsequently when uncovering the trends for the previous one
+
+combined_text|>
+  generate_word_graph(minimum_n = 100, minimum_correlation = 0.2)
+
+combined_text|>
+  generate_word_graph(minimum_n = 200, minimum_correlation = 0.4)
+
+combined_text|>
+  generate_word_graph(minimum_n = 200, minimum_correlation = 0.8)
+```
 # Discussion

### 2024-04-03 scripts/02-data_cleaning.R

@@ -15,7 +15,8 @@ library(tm)
 library(dplyr)
 library(ggplot2)
 library(scales)
-
+library(igraph)
+library(widyr)
 
 #### Clean data ####
 portrait <- read_csv("data/raw_data/portrait.csv")
@@ -73,44 +74,40 @@ prufrock_clean <- prufrock_tokenized |>
 
 # Use count() to quickly get the count of the most frequently used words
 portrait_words <- portrait_clean |>
-  count(word, sort = TRUE)
+  count(word, sort = TRUE) |>
+  filter(str_detect(word, "[:alpha:]")) |> # Remove rows where the word column does not contain any alphabetical characters.
+  distinct() # Return distinct words only
 
 swann_words <- swann_clean |>
-  count(word, sort = TRUE)
+  count(word, sort = TRUE) |>
+  filter(str_detect(word, "[:alpha:]")) |>
+  distinct()
 
 dalloway_words <- dalloway_clean |>
-  count(word, sort = TRUE)
+  count(word, sort = TRUE) |>
+  filter(str_detect(word, "[:alpha:]")) |> 
+  distinct()
 
 bliss_words <- bliss_clean |>
-  count(word, sort = TRUE)
+  count(word, sort = TRUE) |>
+  filter(str_detect(word, "[:alpha:]")) |> 
+  distinct()
 
 prufrock_words <- prufrock_clean |>
-  count(word, sort = TRUE)
+  count(word, sort = TRUE) |>
+  filter(str_detect(word, "[:alpha:]")) |> 
+  distinct()
 
 # Combine all texts to find the more general trend of word frequency
 
-## Bigram analysis or analysis by prepositioned words, codes sourced from https://www.youtube.com/watch?v=9T-hr3jinTw
-combined_trigram <-
-  rbind(portrait, swann, dalloway, prufrock, bliss)|>
-  na.omit(portrait, swann, dalloway, prufrock, bliss) |>
-  unnest_tokens(bigram, text, token = "ngrams", n = 3) |>
-  separate(bigram, c("word1", "word2", "word3"), sep = " ") #|>
-  #count(word1, word2, word3, sort = TRUE)
+## Calculate the correlations between all word combinations from all texts
 
-combined_bigram <-
-  rbind(portrait, swann, dalloway, prufrock, bliss)|>
-  na.omit(portrait, swann, dalloway, prufrock, bliss) |>
[…]

### 2024-04-03 paper/paper.qmd
update graphs

@@
 #| label: tbl-joyce
-#| tbl-cap: "Examplary Table of Unprocessed Stream of Consciousness Novels (James Joyce)"
+#| tbl-cap: "An Examplary Table Containing Unprocessed Novel Text (James Joyce)"
 #| warning: false
@@
 #| label: tbl-joyce-token
-#| tbl-cap: "Examplary Table of Tokenized Stream of Consciousness Novels (James Joyce)"
+#| tbl-cap: "An Examplary Table Containing Tokenzied Novel Text (James Joyce)"
 #| warning: false
@@
 #| label: tbl-joyce-words
-#| tbl-cap: "Examplary Table of Word Count of Stream of Consciousness Novels (James Joyce)"
+#| tbl-cap: "An Examplary Table Containing Word Count of Each Word within Novel Texts (James Joyce)"
 #| warning: false

### 2024-04-03 paper/paper.qmd
update paper

@@
 
-By analyzing these textual datasets, I seek to pose and answer crucial questions: What are some important factors contributing to this relationship between mental health, disassociation and stream of consciousness?
-Moreover, how does this relationship vary differently across differnet demographics of authors, for instance, authors with different geographical locations and genders?
-Understanding these dynamics is crucial in having an informed understanding of the West's late 19th to mid-20th century socio-political landscape, especially in regards to how authors and creative writers navigate and deal with then-taboo topics such as existential angst, mental health issues and disabilities.
+By analyzing these textual datasets, I seek to pose and answer crucial questions: \textit{What are some important factors contributing to this relationship between mental health, disassociation and stream of consciousness?
+Moreover, how does this relationship vary differently across differnet demographics of authors, for instance, authors with different geographical locations and genders?} Understanding these dynamics is crucial in having an informed understanding of the West's late 19th to mid-20th century socio-political landscape, especially in regards to how authors and creative writers navigate and deal with then-taboo topics such as existential angst, mental health issues and disabilities.
 

### 2024-04-03 paper/paper.qmd
update formatting

@@
 bibliography: references.bib
+toc: true
 ---
@@
 
+The goal of our modelling strategy is twofold. Firstly,...
+
+Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+
 ## Model set-up
 
+Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+
+\begin{align} 
+y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
+\mu_i &= \alpha + \beta_i + \gamma_i\\
+\alpha &\sim \mbox{Normal}(0, 2.5) \\
+\beta &\sim \mbox{Normal}(0, 2.5) \\
+\gamma &\sim \mbox{Normal}(0, 2.5) \\
+\sigma &\sim \mbox{Exponential}(1)
+\end{align}
+
+We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+
 ### Model justification
+
+We expect a negative relationship between average household income and the number of children per child care space by ward. In particular...
+
+We can use maths by including latex between dollar signs, for instance $\theta$.
 
@@
 ```
+\newpage
 
@@
 ```
-
+\newpage
 ## Gendered Mental Landscape of Stream of Consciousness Novels
@@
 
+\newpage
+
 ## Transnational Stream of Consciousness Novels and Mental Health Themes
@@
 
+\newpage
 ## Combined Texts: Trends, Word Networks, Bigram and Trigram Analsis

### 2024-04-03 paper/paper.qmd
update references

@@
 #| fig-cap: "Word Networks Measured by Frequency and Correlation when Combining All Stream of Consciousness Novels"
-#| fig-subcap: ["100 Frequency, 0.2 Correlation", "200 Frequency, 0.4 Correlation", "200 Frequency, 0.8 Correlation"] #Add "Two Words" and "Three Words" subsequently when uncovering the trends for the previous one
+#| fig-subcap: ["100 Frequency, 0.2 Correlation", "200 Frequency, 0.4 Correlation", "200 Frequency, 0.8 Correlation", "400 Frequency, 1 Correlation"]
 
@@
   generate_word_graph(minimum_n = 200, minimum_correlation = 0.8)
+
+combined_text|>
+  generate_word_graph(minimum_n = 400, minimum_correlation = 1)
 ```
@@
 ### Limiting Literature Works and Publication Editions
+
+### Uneven Novel Length and Categorization of Authors
 
@@
 ```{r fig.pos="H"}
-#| echo: false
-#| message: false
-#| label: tbl-reasons-strip-search
-#| tbl-cap: Summary Statistics of Reasons for Police Strip Searches
+##| echo: false
+##| message: false
+##| label: tbl-reasons-strip-search
+##| tbl-cap:
 
@@
 ```{r fig.pos="H"}
-#| echo: false
-#| message: false
-#| label: tbl-items-strip-search
-#| tbl-cap: Summary Statistics of Items Found During Police Strip Searches
+##| echo: false
+##| message: false
+##| label: tbl-items-strip-search
+##| tbl-cap:
 
@@
 
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-```
-
 ## Diagnostics
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-```
 

### 2024-04-03 paper/paper.qmd
update paper

@@
 By analyzing the word frequency of their famous works, I explore mental health themes like anxiety, depression, trauma, and existential angst.
-Through word frequency analysis of nine novels, I aim to uncover shared linguistic patterns and gain insights into the authors' mental states, offering a glimpse into themes of self-identity and existential contemplation."
+Through word frequency analysis of nine novels, I aim to uncover shared linguistic patterns and gain insights into the authors' mental states, offering a glimpse into themes of self-identity and existential contemplation.
+(add one sentence on main results"
 format: pdf
@@
 ```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
+##| eval: true
+##| echo: false
+##| message: false
+##| warning: false
 #combined_books 

### 2024-04-03 paper/paper.qmd
update graphs

@@
 abstract: "This project focuses on understanding the language used by renowned, transnational stream of consciousness authors James Joyce, Virginia Woolf, Marcel Proust, Katherine Mansfield and T.S Eliot.
-By analyzing the word frequency of their famous works, I explore mental health themes like anxiety, depression, trauma, and existential angst.
+By conducting word frequency analysis and sentiment analysis of their famous works, I explore mental health themes like anxiety, depression, trauma, and existential angst.
 Through word frequency analysis of nine novels, I aim to uncover shared linguistic patterns and gain insights into the authors' mental states, offering a glimpse into themes of self-identity and existential contemplation.
-(add one sentence on main results"
+(add one sentence on main results)"
 format: pdf
@@
 library(ggraph)
+library(textdata)
 
@@
 
-By analyzing these textual datasets, I seek to pose and answer crucial questions: \textit{What are some important factors contributing to this relationship between mental health, disassociation and stream of consciousness?
+By analyzing these textual datasets through word frequency and sentiment analysis, I seek to pose and answer crucial questions: \textit{What are some important factors contributing to this relationship between mental health, disassociation and stream of consciousness?
 Moreover, how does this relationship vary differently across differnet demographics of authors, for instance, authors with different geographical locations and genders?} Understanding these dynamics is crucial in having an informed understanding of the West's late 19th to mid-20th century socio-political landscape, especially in regards to how authors and creative writers navigate and deal with then-taboo topics such as existential angst, mental health issues and disabilities.
@@
 The novel texts used for analysis were sourced from Project Gutenberg under the library `gutenbergr` [@rGuten] [@gutenberg].
-Data was generated, extracted and cleaned using the open-source statistical programming language R [@r], leveraging functions from `tidyverse` [@rTidyverse], `tidytext` [@rTidytext], `rmarkdown` [@rMark], `dplyr` [@rDplyr], `ggplot2` [@rGgplot2], `scales` [@rScales], `here` [@rHere], `igraph` [@rIgraph], `widyr` [@rWidy], `ggraph` [@rGg] `tm` [@rTm] and `knitr` [@rKnitr].
+Data was generated, extracted and cleaned using the open-source statistical programming language R [@r], leveraging functions from `tidyverse` [@rTidyverse], `tidytext` [@rTidytext], `rmarkdown` [@rMark], `dplyr` [@rDplyr], `ggplot2` [@rGgplot2], `scales` [@rScales], `here` [@rHere], `igraph` [@rIgraph], `widyr` [@rWidy], `ggraph` [@rGg], `textdata` [@rTextdata], `tm` [@rTm] and `knitr` [@rKnitr].
 
@@
          "Text" = text,
-         "Book" = book) |>
+         "Book" = book,
+         "Author" = author) |>
   head(10) |>
   kable(
-    col.names = c("Book ID", "Text", "Book"),
+    col.names = c("Book ID", "Text", "Book", "Author"),
     booktabs = TRUE
@@
          "Book" = book,
+         "Author" = author,
          "Word" = word) |>
@@
   kable(
-    col.names = c("Book ID", "Book", "Word"),
+    col.names = c("Book ID", "Book", "Author", "Word"),
     booktabs = TRUE
@@
 
-## Great Britain as the Center of Stream of Consciousness Literature
+## Sentiment Analysis
+
+Leveraging setiment analysis from @sentiment
 
@@
 #| warning: false
-#| label: fig-frequency-brit
-#| fig-cap: "Comparative Analysis of Word Frequency in British Stream of Consciousness Authors"
+#| label: fig-sentiment
+#| fig-cap: "Sentiment Analysis of All Stream of Consciousness Books"
 
-frequency_britain <- bind_rows(
-                       mutate(dalloway_words, author = "Virginia Woolf"),
-                       mutate(prufrock_words, author = "T.S. Eliot"),
-                       mutate(portrait_words, author = "James Joyce")) |>
-  mutate(word = str_extract(word, "[a-z']+")) |>
-  count(author, word) |>
+sentiment_combined <- combined_text |>
+  inner_join(get_sentiments("bing")) |>
+  inner_join(get_sentiments("afinn")) |>
+  inner_join(get_sentiments("nrc"))
+
+sentiment_category <- sentiment_combined |>
+  inner_join(get_sentiments("bing")) |>
+  group_by(author, sentiment) |>
+  count() |>
   group_by(author) |>
-  mutate(proportion = n / sum(n)) |>
-  select(-n) |>
-  pivot_wider(names_from = author, values_from = proportion) |>
-  pivot_longer(`T.S. Eliot`:`James Joyce`,
-               names_to = "author", values_to = "proportion")
+  mutate(percentage = n / sum(n) * 100)
 
-ggplot(frequency_britain, aes(x = proportion, y = `Virginia Woolf`, color = abs(`Virginia Woolf` - proportion))) +
-  geom_abline(color = "gray50", lty = 2) +
-  geom_jitter(alpha = 0.05, size = 1.5, width = 0.7, height = 0.7) +  
-  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
-  scale_x_log10(labels = percent_format()) +
-  scale_y_log10(labels = percent_format()) +
-  scale_color_gradient(limits = c(0, 0.001), 
-                       low = "darkslateblue", high = "gray1") +
-  facet_wrap(~author, ncol = 2) +
-  theme(legend.position="none") +
-  labs(y = "Virginia Woolf", x = NULL)
+# Plot sentiment values
+ggplot(sentiment_category, aes(x = author, fill = sentiment, y = percentage)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  labs(
+       x = "Author", y = "Percentage") +
+  scale_fill_manual(values = c("positive" = "gray80", "negative" = "gray20", "neutral" = "gray50")) + 
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
 ```
+
+```{r}
+#| eval: true
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-sentiment-number
+#| fig-cap: "Sentiment Analysis of All Stream of Consciousness Books by Value"
+
+sentiment_value <- sentiment_combined |>
+  inner_join(get_sentiments("bing")) |>
+  group_by(author, value, sentiment) |>
+  count() |>
+  group_by(author) |>
+  mutate(percentage = n / sum(n) * 100)
+
+ggplot(sentiment_value, aes(x = value, y = percentage, color = sentiment, group = sentiment)) +
+  geom_line() +
+  facet_wrap(~ author, scales = "free") +
+  labs(title = "Sentiment Analysis of Books",
+       x = "Sentiment Value", y = "Percentage") +
+  scale_color_manual(values = c("negative" = "black", "postive" = "gray")) + # Adjust line colors
+  theme_minimal()
+```
+
 \newpage

### 2024-04-03 scripts/01-download_data.R

@@ -45,22 +45,27 @@ combined_books <- rbind(joyce_books, woolf_books, dostoevsky_books, proust_books
 
 portrait <- gutenberg_download(c('4217', '2817')) |>
   mutate(book = ifelse(gutenberg_id == 4217, "A Portrait of the Artist as a Young Man", "Chamber Music")) |>
+  mutate(author = ifelse(gutenberg_id == 4217, "James Joyce", "James Joyce")) |>
   slice(-(1:20))
 
 swann_way <- gutenberg_download(c('7178')) |>
   mutate(book = ifelse(gutenberg_id == 7178, "Swann's Way", "NA")) |>
+  mutate(author = ifelse(gutenberg_id == 7178, "Marcel Proust", "Marcel Proust")) |>
   slice(-(1:33))
 
 mrs_dalloway <- gutenberg_download(c('63107', '5670')) |>
   mutate(book = ifelse(gutenberg_id == 63107, "Mrs Dalloway", "Jacob's Room")) |>
+  mutate(author = ifelse(gutenberg_id == 63107, "Virigina Woolf", "Virginia Woolf")) |>
   slice(-(1:25))
 
 prufrock <- gutenberg_download(c('1321', '1459')) |>
   mutate(book = ifelse(gutenberg_id == 1321, "The Waste Land", "Prufrock and Other Observations")) |>
+  mutate(author = ifelse(gutenberg_id == 63107, "T.S. Eliot", "T.S. Eliot")) |>
   slice(-(1:35))
 
 bliss <- gutenberg_download(c('44385', '1429')) |>
   mutate(book = ifelse(gutenberg_id == 1321, "Bliss", "The Garden Party")) |>
+  mutate(author = ifelse(gutenberg_id == 63107, "Katherine Mansfield", "Katherine Mansfield")) |>
   slice(-(1:34))
 
 #### Save data ####

### 2024-04-03 scripts/02-data_cleaning.R

@@ -117,8 +117,6 @@ write_csv(dalloway_words, "data/analysis_data/dalloway_words.csv")
 write_csv(bliss_words, "data/analysis_data/bliss_words.csv")
 write_csv(prufrock_words, "data/analysis_data/prufrock_words.csv")
 write_csv(combined_text, "data/analysis_data/combined_text.csv")
-write_csv(combined_text_count, "data/analysis_data/combined_text_count.csv")
-write_csv(combined_correlations, "data/analysis_data/combined_correlations.csv")
 
 
 

### 2024-04-03 paper/paper.qmd
update title

@@
 title: "The Linguistic Landscape of Stream-of-Consciousness Literature"
-subtitle: "Exploring Word Frequency and Mental Health Themes in the Works of Joyce, Woolf, Proust, Mansfield and Eliot from Project Gutenberg"
+subtitle: "Exploring Word Frequency, Sentiment Value and Mental Health Themes in the Works of Joyce, Woolf, Proust, Mansfield and Eliot from Project Gutenberg"
 author: Quang Mai

### 2024-04-03 paper/paper.qmd
update paper

--- 2024-04-03 paper/paper.qmd
@@
 date-format: long
-abstract: "This project focuses on understanding the language used by renowned, transnational stream of consciousness authors James Joyce, Virginia Woolf, Marcel Proust, Katherine Mansfield and T.S Eliot.
-By conducting word frequency analysis and sentiment analysis of their famous works, I explore mental health themes like anxiety, depression, trauma, and existential angst.
-Through word frequency analysis of nine novels, I aim to uncover shared linguistic patterns and gain insights into the authors' mental states, offering a glimpse into themes of self-identity and existential contemplation.
+abstract: "This project focuses on understanding the language used by renowned stream of consciousness (SOC) authors James Joyce, Virginia Woolf, Marcel Proust, Katherine Mansfield and T.S Eliot.
+By conducting word frequency analysis and sentiment analysis of these authors' nine novels, I aim to uncover shared linguistic patterns and gain insights into the authors' mental states.
+Through the analysis on SOC literature, this paper thus attempts to offer insights into themes of self-identity, anxiety, disassciation and existential contemplation within Western society's from late 19th to mid-20th century.
 (add one sentence on main results)"
@@
 Its relationship to mental health lies in its ability to offer insight into the depths of consciousness, providing a platform to explore themes such as anxiety, trauma, and existential questioning.
-As such, this paper has mined the texts of a total of nine novels from the volunteer archive, Project Gutenberg, to examine the mental health themes of famous stream of consciousness authors, namely by Joyce, Woolf, Proust, Mansfield and Eliot, from the modernist era of literature, spanning from late 19th century to the mid-20th century [@gutenberg].
+As such, this paper has mined the texts of a total of nine novels from the volunteer archive, Project Gutenberg, to examine the mental health themes of famous stream of consciousness authors, namely by Joyce, Woolf, Proust, Mansfield and Eliot, from the modernist era of literature, spanning from late 19th century to mid-20th century [@gutenberg].
 (more stats and data mentioned here)
@@
          "Author" = author) |>
-  head(10) |>
+  head(5) |>
   kable(
@@
          "Word" = word) |>
-  tail(10) |>
+  tail(5) |>
   kable(
@@
   rename("Word" = word) |>
-  head(10) |>
+  head(5) |>
   kable(
@@
   rename("Word" = word) |>
-  head(12) |>
+  head(5) |>
   kable(
@@
 
-## The Dominant Vocabulary of Stream of Consciousness
+## The Dominant Vocabulary of SOC Literature
 
@@
 ggplot(sentiment_category, aes(x = author, fill = sentiment, y = percentage)) +
-  geom_bar(stat = "identity", position = "dodge") +
+  geom_bar(stat = "identity") +
   labs(
@@
   facet_wrap(~ author, scales = "free") +
-  labs(title = "Sentiment Analysis of Books",
+  labs(
        x = "Sentiment Value", y = "Percentage") +
-  scale_color_manual(values = c("negative" = "black", "postive" = "gray")) + # Adjust line colors
+  scale_color_manual(values = c("positive" = "black", "negative" = "gray")) + # Adjust line colors
   theme_minimal()
@@
 
-\newpage
 ## Gendered Mental Landscape of Stream of Consciousness Novels
@@
 
-### Comparing Female Stream of Consciousness Authors
+### Comparing SOC Literature's Female Authors
 
-### Comparing Male Stream of Consciousness Authors
+### Comparing SOC Literature's Male Authors
 
-\newpage
-
-## Transnational Stream of Consciousness Novels and Mental Health Themes
+## Transnational SOC Novels and Mental Health Themes
 

### 2024-04-03 paper/paper.qmd
update paper

@@
 #| layout-ncol: 2
-#| fig-cap: "Comparative Analysis of Top 20 Word Frequencies by Stream of Consciousness Authors' Novels"
+#| fig-cap: "Comparative Analysis of Top 20 Word Frequencies from Famous SOC Authors"
 #| fig-subcap: ["James Joyce", "Virginia Woolf", "Marcel Proust", "T.S. Eliot", "Katherine Mansfield"]
@@
 #| label: fig-sentiment
-#| fig-cap: "Sentiment Analysis of All Stream of Consciousness Books"
+#| fig-cap: "Sentiment Analysis of All SOC Novels"
 
@@
 #| label: fig-sentiment-number
-#| fig-cap: "Sentiment Analysis of All Stream of Consciousness Books by Value"
+#| fig-cap: "Sentiment Analysis of All SOC Novels by Sentiment Value"
 
@@
   scale_color_gradient(limits = c(0, 0.001), 
-                       low = "darkslateblue", high = "gray50") +
+                       low = "blue2", high = "gray50") +
   facet_wrap(~author, ncol = 1) +
@@
   scale_color_gradient(limits = c(0, 0.001), 
-                       low = "darkslateblue", high = "gray100") +
+                       low = "blue2", high = "gray100") +
   facet_wrap(~author, ncol = 2) +
@@
   scale_color_gradient(limits = c(0, 0.001), 
-                       low = "darkslateblue", high = "gray100") +
+                       low = "blue2", high = "gray100") +
   facet_wrap(~author, ncol = 1) +

### 2024-04-04 scripts/02-data_cleaning.R
Update 02-data_cleaning.R

@@ -4,7 +4,7 @@
 # Date: 30 March 2024 
 # Contact: quangpono@gmail.com
 # License: MIT
-# Pre-requisites: Read four raw datasets containing the novel texts from five authors
+# Pre-requisites: Read four raw datasets containing the novel texts from five authors, namely `portrait.csv`, `swann.csv`. `dalloway.csv`, `bliss.csv` and `prufrock.csv`
 
 #### Workspace setup ####
 library(tidyverse)

### 2024-04-04 scripts/02-data_cleaning.R
Update 02-data_cleaning.R

@@ -27,7 +27,7 @@ prufrock <- read_csv("data/raw_data/prufrock.csv")
 
 
 # Tokenize the words from the text, breaking the text down into individual words to insert them into individual datasets
-# Codes sourced from https://medium.com/the-data-nudge/nlp-basics-exploring-word-frequency-with-the-tidytext-r-package-ac35a6d805f4
+# Codes referenced from https://medium.com/the-data-nudge/nlp-basics-exploring-word-frequency-with-the-tidytext-r-package-ac35a6d805f4
 
 portrait_tokenized <- portrait |>
   na.omit(portrait) |>

### 2024-04-04 scripts/01-download_data.R
Update 01-download_data.R

@@ -17,7 +17,7 @@ library(ggplot2)
 
 
 # Collect stream-of-conscious authors main texts
-# Codes sourced from https://medium.com/the-data-nudge/nlp-basics-exploring-word-frequency-with-the-tidytext-r-package-ac35a6d805f4
+# Codes referenced from https://medium.com/the-data-nudge/nlp-basics-exploring-word-frequency-with-the-tidytext-r-package-ac35a6d805f4
 
 all_books <- gutenberg_metadata
 

### 2024-04-04 paper/paper.qmd
update paper

@@
 By conducting word frequency analysis and sentiment analysis of these authors' nine novels, I aim to uncover shared linguistic patterns and gain insights into the authors' mental states.
-Through the analysis on SOC literature, this paper thus attempts to offer insights into themes of self-identity, anxiety, disassciation and existential contemplation within Western society's from late 19th to mid-20th century.
+Through the analysis on SOC literature, this paper thus attempts to offer insights into themes of self-identity, anxiety, disassociation and existential contemplation within Western society's from late 19th to mid-20th century.
 (add one sentence on main results)"
@@
 
-Stream of consciousness is a narrative technique that aims to capture the continuous flow of thoughts, feelings, and sensations experienced by a character without conventional organization or punctuation  [@resampling].
+Stream of consciousness (SOC) is a narrative technique that aims to capture the continuous flow of thoughts, feelings, and sensations experienced by a character without conventional organization or punctuation  [@resampling].
 It mirrors the unpredictable and interconnected nature of human thought processes, often revealing the inner workings of the character's mind in an intimate and unfiltered manner [@flow].
 In literature, most scholars agree that stream of consciousness reveals the complexities of mental-scapes, shedding light on the nuances of characters' emotional well-being and psychological struggles [@pathology].
-Its relationship to mental health lies in its ability to offer insight into the depths of consciousness, providing a platform to explore themes such as anxiety, trauma, and existential questioning.
 As such, this paper has mined the texts of a total of nine novels from the volunteer archive, Project Gutenberg, to examine the mental health themes of famous stream of consciousness authors, namely by Joyce, Woolf, Proust, Mansfield and Eliot, from the modernist era of literature, spanning from late 19th century to mid-20th century [@gutenberg].
@@
 
-Thus, my estimand is the correlation between famous stream of consciousness novels, words related to mental health themes and their frequency.
-This is considered in terms of the nine selected novels only, namely Joyce's A Portrait of the Artist as a Young Man and Chamber Music; Woolf's Mrs Dalloway and Jacob's Room; Proust's Swann Way; Mansfield's Bliss and The Garden Party; and Eliot's The Waste Land and The Love Song of J.
+Thus, my estimand is the correlation between mental health-related words in SOC literature, their frequency and sentimental value as provided by @sentiment.
+This is considered in terms of nine selected SOC novels only, namely Joyce's A Portrait of the Artist as a Young Man and Chamber Music; Woolf's Mrs Dalloway and Jacob's Room; Proust's Swann Way; Mansfield's Bliss and The Garden Party; and Eliot's The Waste Land and The Love Song of J.
 Alfred Prufrock.
@@
 To further understand the correlation between stream of consciousness novels and mental health themes, in [Introduction], our paper briefly discusses the nature of stream of consciousness literature, relevant authors and the works that we've chosen to analyze.
-Subsequently, in [Data] and [Result], we talk about the nature of the data obtained and analyze the results garnered from the data with suitable tables and charts.
+Subsequently, in [Data] and [Results], we talk about the nature of the data obtained and analyze the results garnered from the data with suitable tables and charts.
 Next, [Discussion] provides further insights and future areas of study.
@@
 
-### Limiting Literature Works and Publication Editions
+### Decontextualized Literature Works and Limiting Publication Editions
 

### 2024-04-04 scripts/01-download_data.R

@@ -37,9 +37,7 @@ ts_books <- all_books %>%
   filter(str_detect(author, 'Eliot, T. S.'))
 
 
-
 combined_books <- rbind(joyce_books, woolf_books, dostoevsky_books, proust_books, ts_books, mansfield_books)
-
 
 # Download the more famous stream-of-consciousness work of each author:
 

### 2024-04-04 paper/paper.qmd
update title

@@
 # Calculate the word frequency over all unique words and bind all frequency rate of authors together to conduct a comparative comparison
-# Code sourced from https://medium.com/the-data-nudge/nlp-basics-exploring-word-frequency-with-the-tidytext-r-package-ac35a6d805f4
+# Code referenced from https://medium.com/the-data-nudge/nlp-basics-exploring-word-frequency-with-the-tidytext-r-package-ac35a6d805f4
 

### 2024-04-04 paper/paper.qmd
update paper

@@
 ---
-title: "The Linguistic Landscape of Stream-of-Consciousness Literature"
+title: "A Joyceless Stream-of-Consciousness Book" #The Linguistic Landscape of SOC Literature
 subtitle: "Exploring Word Frequency, Sentiment Value and Mental Health Themes in the Works of Joyce, Woolf, Proust, Mansfield and Eliot from Project Gutenberg"
@@
 It mirrors the unpredictable and interconnected nature of human thought processes, often revealing the inner workings of the character's mind in an intimate and unfiltered manner [@flow].
-In literature, most scholars agree that stream of consciousness reveals the complexities of mental-scapes, shedding light on the nuances of characters' emotional well-being and psychological struggles [@pathology].
+In literature, most scholars agree that stream of consciousness reveals the complexities of mind-scapes, shedding light on the nuances of characters' emotional well-being and psychological struggles [@pathology].
 As such, this paper has mined the texts of a total of nine novels from the volunteer archive, Project Gutenberg, to examine the mental health themes of famous stream of consciousness authors, namely by Joyce, Woolf, Proust, Mansfield and Eliot, from the modernist era of literature, spanning from late 19th century to mid-20th century [@gutenberg].
@@
 ## Measurement
+
+Two central variables in this paper are:
+
+* **Word Frequency:** This variable captures the repetition of a single word (unigram), two-words combination (bigram) or three-words combination (trigram) throughout a SOC novel text, providing us with a thematic understanding of SOC literature.
+* **Sentiment Value:** This variable enables us to analyze how every word is usually perceived emotionally, whether it be positively or negatively.
+In additon,    in childlessness over time, encompassing the chosen time frame for the survey.
+
+Out of two variables used, the first one, word frequency, is directly quantified through tokenizing the novel texts using the packages @rTidytext [@youtube1] [@youtube2].
+To do this, I first downloaded all nine novel texts from @gutenberg, and leveraged functions such as `unnest_tokens()` from @rTidytext to mine the texts, or separating it into individual words.
+Finally, I used `count()` to quantify the word frequency.
+
+The second variable used, sentiment value, is based on @sentiment's research, where  questioning to accurately capture respondents' backgrounds.
+For instance, the General Social Survey (GSS) obtains information on marital status through questions such as, "Are you currently married, widowed, divorced, separated, or have you never been married?", with responses categorized into predefined groups for further analysis [@gssCodebook].
+On the other hand, instead of relying on income sources, socioeconomic class is subjectively assessed based on the respondents' answers to this question, "If you were asked to use one of four names for your social class, which would you say you belong in: the lower class, the working class, the middle class, or the upper class?".
+Some are continuous variables, like age.
+With respondents' various age groups documented, we can understand the generational disparities in childbearing patterns.
+Systematic and data-driven, these measurement methods ensure that all data faithfully reflects the American population's childbearing tendencies.
+
+**Family and Childbearing**
+
+(@) Do you currently have any children?
+- Yes
+- No
+
+(@) If yes, how many children have you ever had? Please count all that were born alive at any time (including any you had from a previous marriage).
+(short answer text)
+
+Even though considerable attention has been given to the polarity of words (positive and negative) and thecreation of large polarity lexicons, research in emotion analysis has had to rely on limited and small emotionlexicons.
+In this paper, we show how the combined strength and wisdom of the crowds can be used to generate alarge, high-quality, word–emotion and word–polarity association lexicon quickly and inexpensively.
+We enumeratethe challenges in emotion annotation in a crowdsourcing scenario and propose solutions to address them.
+Mostnotably, in addition to questions about emotions associated with terms, we show how the inclusion of a word choicequestion can discourage malicious data entry, help to identify instances where the annotator may not be familiarwith the target term (allowing us to reject such annotations), and help to obtain annotations at sense level (ratherthan at word level).
+We conducted experiments on how to formulate the emotion-annotation questions, and showthat asking if a term isassociatedwith an emotion leads to markedly higher interannotator agreement than thatobtained by asking if a termevokesan emotion
 

### 2024-04-04 scripts/00-simulate_data.R

@@ -1,11 +1,11 @@
 #### Preamble ####
-# Purpose: Simulates... [...UPDATE THIS...]
+# Purpose: Simulates
 # Author: Rohan Alexander [...UPDATE THIS...]
 # Date: 11 February 2023 [...UPDATE THIS...]
 # Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites:
+# Any other information needed? No
 
 
 #### Workspace setup ####

### 2024-04-04 paper/paper.tex
update measurement

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={The Joyceless Landscape of Stream of Consciousness Literature},
  pdfauthor={Quang Mai},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{The Joyceless Landscape of Stream of Consciousness
Literature\thanks{Code and data are available at:
https://github.com/ponolite/stream-consciousness-language.git}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Exploring Word Frequency, Sentiment Value and Mental Health
Themes in the Works of Joyce, Woolf, Proust, Mansfield and Eliot from
Project Gutenberg}
\author{Quang Mai}
\date{April 4, 2024}

\begin{document}
\maketitle
\begin{abstract}
This project focuses on understanding the language used by renowned
stream of consciousness (SOC) authors James Joyce, Virginia Woolf,
Marcel Proust, Katherine Mansfield and T.S Eliot. By conducting word
frequency analysis and sentiment analysis of these authors' nine novels,
I aim to uncover shared linguistic patterns and gain insights into the
authors' mental states. Through the analysis on SOC literature, this
paper thus attempts to offer insights into themes of self-identity,
anxiety, disassociation and existential contemplation within Western
society's from late 19th to mid-20th century. (add one sentence on main
results)
\end{abstract}

\renewcommand*\contentsname{Table of contents}
{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{3}
\tableofcontents
}
\section{Introduction}\label{introduction}

Stream of consciousness (SOC) is a narrative technique that aims to
capture the continuous flow of thoughts, feelings, and sensations
experienced by a character without conventional organization or
punctuation (Bernini and Fernyhough 2022). It mirrors the unpredictable
and interconnected nature of human thought processes, often revealing
the inner workings of the character's mind in an intimate and unfiltered
manner (Long and So 2016). In literature, most scholars agree that
stream of consciousness reveals the complexities of mind-scapes,
shedding light on the nuances of characters' emotional well-being and
psychological struggles (Nyongesa 2023). As such, this paper has mined
the texts of a total of nine novels from the volunteer archive, Project
Gutenberg, to examine the mental health themes of famous stream of
consciousness authors, namely by Joyce, Woolf, Proust, Mansfield and
Eliot, from the modernist era of literature, spanning from late 19th
century to mid-20th century ({``Project Gutenberg,''} n.d.). (more stats
and data mentioned here)

By analyzing these textual datasets through word frequency and sentiment
analysis, I seek to pose and answer crucial questions:
\textit{What are some important factors contributing to this relationship between mental health, disassociation and stream of consciousness? Moreover, how does this relationship vary differently across differnet demographics of authors, for instance, authors with different geographical locations and genders?}
Understanding these dynamics is crucial in having an informed
understanding of the West's late 19th to mid-20th century literature and
even socio-political landscape, especially in regards to how authors and
creative writers navigate and deal with then-taboo topics such as
existential angst, mental health issues and disabilities.

Thus, my estimand is the correlation between mental health-related words
in SOC literature, their frequency and sentimental value as provided by
Mohammad and Turney (2013). This is considered in terms of nine selected
SOC novels only, namely Joyce's A Portrait of the Artist as a Young Man
and Chamber Music; Woolf's Mrs Dalloway and Jacob's Room; Proust's Swann
Way; Mansfield's Bliss and The Garden Party; and Eliot's The Waste Land
and The Love Song of J. Alfred Prufrock. Through our analysis, we found
that (percentage, number and data here, main results)\ldots{}

To further understand the correlation between stream of consciousness
novels and mental health themes, in
\hyperref[introduction]{Introduction}, the paper briefly discusses the
nature of stream of consciousness literature, relevant authors and the
works that I've chosen to analyze. Subsequently, in
\hyperref[sec-data]{Data} and \hyperref[results]{Results}, I talk about
the nature of the data obtained and analyze the results garnered from
the data with suitable tables and charts. Next,
\hyperref[discussion]{Discussion} provides further insights and future
areas of study. Finally, {[}Conclusion{]} summarizes our main findings.
To complete the paper, \hyperref[appendix]{Appendix} clarifies how each
variable within each dataset is generated with relevant tables to
accordingly demonstrate this.

The novel texts used for analysis were sourced from Project Gutenberg
under the library \texttt{gutenbergr} (Johnston and Robinson 2023)
({``Project Gutenberg,''} n.d.). Data was generated, extracted and
cleaned using the open-source statistical programming language R (R Core
Team 2022), leveraging functions from \texttt{tidyverse} (Wickham et al.
2019), \texttt{tidytext} (Julia Silge and Robinson 2016),
\texttt{rmarkdown} (Allaire et al. 2024), \texttt{dplyr} (Wickham et al.
2022), \texttt{ggplot2} (Wickham 2016), \texttt{scales} (Wickham,
Pedersen, and Seidel 2023), \texttt{here} (Müller 2020), \texttt{igraph}
(J. Silge and Robinson 2006), \texttt{widyr} (J. Silge and Robinson
2022), \texttt{ggraph} (Pedersen 2024), \texttt{textdata} (Hvitfeldt
2022), \texttt{tm} (Feinerer, Hornik, and Meyer 2008) and \texttt{knitr}
(Xie 2014).

\section{Data}\label{sec-data}

\subsection{Measurement}\label{measurement}

Two central variables in this paper are:

\begin{itemize}
\tightlist
\item
  \textbf{Word Frequency:} This variable captures the repetition of a
  single word (unigram), two-words combination (bigram) or three-words
  combination (trigram) throughout a SOC novel text, providing us with a
  thematic understanding of SOC literature.
\item
  \textbf{Sentiment Value:} This variable enables us to analyze how
  every word is usually perceived emotionally, whether it be positively
  or negatively. In additon, in childlessness over time, encompassing
  the chosen time frame for the survey.
\end{itemize}

Out of two variables used, the first one, `Word Frequency' usually
captured as \texttt{n} or \texttt{frequency} in datasets, is directly
quantified through tokenizing the novel texts using the packages Julia
Silge and Robinson (2016; Robertson 2021) (Henry 2021). To do this, I
first downloaded all nine novel texts from {``Project Gutenberg''}
(n.d.), and leveraged functions such as \texttt{unnest\_tokens()} from
Julia Silge and Robinson (2016) to mine the texts, or separating it into
individual words. Finally, I used \texttt{count()} to quantify the word
frequency.

The second variable used, `Sentiment Value', is based on three
English-based, general-purpose ``word-emotion and word-polarity
association lexicons'', specifically Mohammad and Turney (2013)'s
expansive research along with efforts from Finn Årup Nielsen and Bing
Liu and collaborators (Julia Silge and Robinson 2016).

The three general-purpose lexicon that contributes to my sentiment
analysis are (Julia Silge and Robinson 2016):

\begin{itemize}[label={$\square$}]
    \item AFINN from Finn Årup Nielsen, which assigns a numerical value from '-5 to 5' 
    \item bing from Bing Liu and collaborators, which assigns if a word is 'positive' or negative'
    \item nrc from Saif Mohammad and Peter Turney, which assigns a core emotional value to a word, such as 'fear', 'anger', 'sadness' or 'trust'.
\end{itemize}

In terms of measuring `Sentiment Value', all three general-purpose
lexicons are compiled through crowd-sourcing and directly surveying the
public on how each word is emotionally perceived

A survey sample of how the word `startle' is compiled within the `nrc'
lexicon is presented below (Mohammad and Turney 2013):

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\tightlist
\item
  Which word is closest in meaning (most related) to startle?
\end{enumerate}

\begin{itemize}
\tightlist
\item
  automobile
\item
  shake
\item
  honesty
\item
  entertain
\end{itemize}

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\setcounter{enumi}{1}
\tightlist
\item
  How positive (good, praising) is the word startle?:
\end{enumerate}

\begin{itemize}
\tightlist
\item
  startle is not positive
\item
  startle is weakly positive
\item
  startle is moderately positive
\item
  startle is strongly positive
\end{itemize}

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\setcounter{enumi}{2}
\tightlist
\item
  How negative (bad, criticizing) is the word startle?
\end{enumerate}

\begin{itemize}
\tightlist
\item
  startle is not negative
\item
  startle is weakly negative
\item
  startle is moderately negativ
\item
  startle is strongly negative
\end{itemize}

After the survey results are garnered, researchers average the answers
to sort each surveyed word into pre-defined categories, specifically `-5
to 5' for `AFINN', `postive' or `negative' for `bing', and `anger' or
`fear' for `nrc'. With the continuous work of compiling these lexicons
spanning years and decades (Mohammad and Turney 2013) comes these
functions: \texttt{get\_sentiments("bing")} and
\texttt{get\_sentiments("nrc")} and \texttt{get\_sentiments("nrc")} in
which I can use \texttt{inner\_join} to categorize my exisitng datasets
of novel texts into existing sentiment value. Systematic and
data-driven, these measurement methods ensure that all lexicons
faithfully reflects each word's emotionality.

However, I do recognize how decontextualized and reductive this
quantification of language can be. When it comes to understand such
social and human artifacts as language or literary texts, much is
dependent on their contextuality. As such, I will further discuss these
weaknesses of the datasets under \hyperref[discussion]{Discussion}.

\subsection{Source Data}\label{source-data}

\subsection{Data Cleaning and Word
Tokenization}\label{data-cleaning-and-word-tokenization}

\begin{table}

\caption{\label{tbl-genderclass}Table of Number of Classes Students
Considered for Regrade Requests by Students' Gender}

\centering{

}

\end{table}%

\subsubsection{Word Count}\label{word-count}

\begin{longtable}[]{@{}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.0988}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.5802}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1728}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1481}}@{}}

\caption{\label{tbl-joyce}An Examplary Table Containing Unprocessed
Novel Text (James Joyce)}

\tabularnewline

\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedleft
Book ID
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Text
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Book
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Author
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
2817 & To deep and deeper blue, & Chamber Music & James Joyce \\
2817 & NA & Chamber Music & James Joyce \\
2817 & III At that hour when all things have repose, & Chamber Music &
James Joyce \\
2817 & O lonely watcher of the skies, & Chamber Music & James Joyce \\
2817 & NA & Chamber Music & James Joyce \\

\end{longtable}

\begin{longtable}[]{@{}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1176}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.5882}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1765}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1176}}@{}}

\caption{\label{tbl-joyce-token}An Examplary Table Containing Tokenzied
Novel Text (James Joyce)}

\tabularnewline

\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedleft
Book ID
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Book
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Author
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Word
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
4217 & A Portrait of the Artist as a Young Man & James Joyce & stead \\
4217 & A Portrait of the Artist as a Young Man & James Joyce & dublin \\
4217 & A Portrait of the Artist as a Young Man & James Joyce & 1904 \\
4217 & A Portrait of the Artist as a Young Man & James Joyce &
trieste \\
4217 & A Portrait of the Artist as a Young Man & James Joyce & 1914 \\

\end{longtable}

\newpage

\begin{longtable}[]{@{}lr@{}}

\caption{\label{tbl-joyce-words}An Examplary Table Containing Word Count
of Each Word within Novel Texts (James Joyce)}

\tabularnewline

\toprule\noalign{}
Word & Count \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
stephen & 373 \\
god & 194 \\
eyes & 180 \\
soul & 178 \\
father & 151 \\

\end{longtable}

\subsubsection{Comparative Word
Frequency}\label{comparative-word-frequency}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1429}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1429}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2381}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1310}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1786}}@{}}

\caption{\label{tbl-frequency}Word Frequency of Stream of Consciousness
Novels, A Comparison Between Five Authors}

\tabularnewline

\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Word
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
James Joyce
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Katherine Mansfield
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Marcel Proust
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
T.S. Eliot
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Virignia Woolf
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
abandon & 0.000112 & NA & 8.19e-05 & NA & NA \\
abandoned & 0.000112 & NA & 8.19e-05 & NA & NA \\
abandonment & 0.000112 & NA & 8.19e-05 & NA & 0.0001421 \\
abase & 0.000112 & NA & NA & NA & NA \\
abased & 0.000112 & NA & NA & NA & NA \\

\end{longtable}

\subsubsection{Generating Word Networks}\label{generating-word-networks}

\section{Model}\label{model}

The goal of our modelling strategy is twofold. Firstly,\ldots{}

Here we briefly describe the Bayesian analysis model used to
investigate\ldots{} Background details and diagnostics are included in
Appendix~\ref{sec-model-details}.

\subsection{Model set-up}\label{model-set-up}

Define \(y_i\) as the number of seconds that the plane remained aloft.
Then \(\beta_i\) is the wing width and \(\gamma_i\) is the wing length,
both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R (R Core Team 2023) using the \texttt{rstanarm}
package of (\textbf{rstanarm?}). We use the default priors from
\texttt{rstanarm}.

\subsubsection{Model justification}\label{model-justification}

We expect a negative relationship between average household income and
the number of children per child care space by ward. In
particular\ldots{}

We can use maths by including latex between dollar signs, for instance
\(\theta\).

\section{Results}\label{results}

\subsection{The Dominant Vocabulary of SOC
Literature}\label{the-dominant-vocabulary-of-soc-literature}

\begin{figure}

\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics{paper_files/figure-pdf/fig-count-1.pdf}

}

\subcaption{\label{fig-count-1}James Joyce}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics{paper_files/figure-pdf/fig-count-2.pdf}

}

\subcaption{\label{fig-count-2}Virginia Woolf}

\end{minipage}%
\newline
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics{paper_files/figure-pdf/fig-count-3.pdf}

}

\subcaption{\label{fig-count-3}Marcel Proust}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics{paper_files/figure-pdf/fig-count-4.pdf}

}

\subcaption{\label{fig-count-4}T.S. Eliot}

\end{minipage}%
\newline
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics{paper_files/figure-pdf/fig-count-5.pdf}

}

\subcaption{\label{fig-count-5}Katherine Mansfield}

\end{minipage}%

\caption{\label{fig-count}Comparative Analysis of Top 20 Word
Frequencies from Famous SOC Authors}

\end{figure}%

\newpage

\subsection{Sentiment Analysis}\label{sentiment-analysis}

Leveraging setiment analysis from Mohammad and Turney (2013)

\begin{figure}

\centering{

\includegraphics{paper_files/figure-pdf/fig-sentiment-1.pdf}

}

\caption{\label{fig-sentiment}Sentiment Analysis of All SOC Novels}

\end{figure}%

\begin{figure}

\centering{

\includegraphics{paper_files/figure-pdf/fig-sentiment-number-1.pdf}

}

\caption{\label{fig-sentiment-number}Sentiment Analysis of All SOC
Novels by Sentiment Value}

\end{figure}%

\subsection{Gendered Mental Landscape of Stream of Consciousness
Novels}\label{gendered-mental-landscape-of-stream-of-consciousness-novels}

\begin{figure}

\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics{paper_files/figure-pdf/fig-word-gender-1.pdf}

}

\subcaption{\label{fig-word-gender-1}Female Authors}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics{paper_files/figure-pdf/fig-word-gender-2.pdf}

}

\subcaption{\label{fig-word-gender-2}Male Authors}

\end{minipage}%

\caption{\label{fig-word-gender}Comparative Analysis of Word Frequency
in Female and Male Stream of Consciousness Authors}

\end{figure}%

\subsubsection{Comparing SOC Literature's Female
Authors}\label{comparing-soc-literatures-female-authors}

\subsubsection{Comparing SOC Literature's Male
Authors}\label{comparing-soc-literatures-male-authors}

\subsection{Transnational SOC Novels and Mental Health
Themes}\label{transnational-soc-novels-and-mental-health-themes}

\begin{figure}

\centering{

\includegraphics{paper_files/figure-pdf/fig-frequency-int-1.pdf}

}

\caption{\label{fig-frequency-int}Comparative Analysis of Word Frequency
in Transnational Stream of Consciousness Authors}

\end{figure}%

\newpage

\subsection{Combined Texts: Trends, Word Networks, Bigram and Trigram
Analsis}\label{combined-texts-trends-word-networks-bigram-and-trigram-analsis}

\begin{figure}

\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics{paper_files/figure-pdf/fig-word-networks-1.pdf}

}

\subcaption{\label{fig-word-networks-1}100 Frequency, 0.2 Correlation}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics{paper_files/figure-pdf/fig-word-networks-2.pdf}

}

\subcaption{\label{fig-word-networks-2}200 Frequency, 0.4 Correlation}

\end{minipage}%
\newline
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics{paper_files/figure-pdf/fig-word-networks-3.pdf}

}

\subcaption{\label{fig-word-networks-3}200 Frequency, 0.8 Correlation}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics{paper_files/figure-pdf/fig-word-networks-4.pdf}

}

\subcaption{\label{fig-word-networks-4}400 Frequency, 1 Correlation}

\end{minipage}%

\caption{\label{fig-word-networks}Word Networks Measured by Frequency
and Correlation when Combining All Stream of Consciousness Novels}

\end{figure}%

\section{Discussion}\label{discussion}

\subsection{Mental Health Vocabulary: Patterns and
Trends}\label{mental-health-vocabulary-patterns-and-trends}

Discuss vocabulary patterns and word trends

\subsection{Insights into Socio-Political Landscape of the West's
Modernist
Era}\label{insights-into-socio-political-landscape-of-the-wests-modernist-era}

The novels' linguistic patterns reflect the socio-political landscape of
the Western hemisphere, from late 19th century to the mid-20th century.

\subsection{Schizophrenic and Disassociative Tendencies in Female Stream
of
Consciousness}\label{schizophrenic-and-disassociative-tendencies-in-female-stream-of-consciousness}

\subsection{Weaknesses}\label{weaknesses}

\subsubsection{Lack of Thorough Word
Cleaning}\label{lack-of-thorough-word-cleaning}

\subsubsection{Decontextualized Literature Works and Limiting
Publication
Editions}\label{decontextualized-literature-works-and-limiting-publication-editions}

\subsubsection{Uneven Novel Length and Categorization of
Authors}\label{uneven-novel-length-and-categorization-of-authors}

\subsubsection{Project Gutenberg's Focus on the
Canon}\label{project-gutenbergs-focus-on-the-canon}

\subsection{Moving Forward and Next
Steps}\label{moving-forward-and-next-steps}

\newpage

\section{Appendix}\label{appendix}

\subsection{Additional Data Details}\label{additional-data-details}

\begin{Shaded}
\begin{Highlighting}[]
\DocumentationTok{\#\#| eval: true}
\DocumentationTok{\#\#| echo: false}
\DocumentationTok{\#\#| message: false}
\DocumentationTok{\#\#| warning: false}
\CommentTok{\#combined\_books }
\end{Highlighting}
\end{Shaded}

\subsubsection{Data Gathering}\label{data-gathering}

\begin{Shaded}
\begin{Highlighting}[]
\DocumentationTok{\#\#| echo: false}
\DocumentationTok{\#\#| message: false}
\DocumentationTok{\#\#| label: tbl{-}reasons{-}strip{-}search}
\DocumentationTok{\#\#| tbl{-}cap:}
\end{Highlighting}
\end{Shaded}

\subsubsection{Data Cleaning}\label{data-cleaning}

\begin{Shaded}
\begin{Highlighting}[]
\DocumentationTok{\#\#| echo: false}
\DocumentationTok{\#\#| message: false}
\DocumentationTok{\#\#| label: tbl{-}items{-}strip{-}search}
\DocumentationTok{\#\#| tbl{-}cap: }
\end{Highlighting}
\end{Shaded}

\subsection{Model Details}\label{sec-model-details}

\subsection{Posterior predictive
check}\label{posterior-predictive-check}

\subsection{Diagnostics}\label{diagnostics}

\newpage

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-rMark}
Allaire, J., Y. Xie, C. Dervieux, J. McPherson, J. Luraschi, K. Ushey,
A. Atkins, et al. 2024. \emph{Rmarkdown: Dynamic Documents for r}. R
package version 2.26. \url{https://github.com/rstudio/rmarkdown}.

\bibitem[\citeproctext]{ref-resampling}
Bernini, M., and C. Fernyhough. 2022. {``Resampling (Narrative) Stream
of Consciousness: Mind Wandering, Inner Speech, and Reading as Reversed
Introspection.''} \emph{Modern Fiction Studies} 68 (4): 639--67.
https://doi.org/\url{https://doi.org/10.1353/mfs.2022.0045}.

\bibitem[\citeproctext]{ref-rTm}
Feinerer, I., K. Hornik, and D. Meyer. 2008. {``Text Mining
Infrastructure in r.''} \emph{Journal of Statistical Software} 25 (5):
1--54. \url{https://doi.org/10.18637/jss.v025.i05}.

\bibitem[\citeproctext]{ref-youtube2}
Henry, T. 2021. {``Https://Www.youtube.com/Watch?v=ae\_XVhjHd\_o.''}
\url{https://www.youtube.com/watch?v=ae_XVhjHd_o}.

\bibitem[\citeproctext]{ref-rTextdata}
Hvitfeldt, Emil. 2022. \emph{Textdata: Download and Load Various Text
Datasets}. \url{https://github.com/EmilHvitfeldt/textdata}.

\bibitem[\citeproctext]{ref-rGuten}
Johnston, Myfanwy, and David Robinson. 2023. \emph{Gutenbergr: Download
and Process Public Domain Works from Project Gutenberg}.
\url{https://docs.ropensci.org/gutenbergr/}.

\bibitem[\citeproctext]{ref-flow}
Long, H., and J. So R. 2016. {``Turbulent Flow: A Computational Model of
World Literature.''} \emph{Modern Language Quarterly} 77 (3): 345--67.
https://doi.org/\url{https://doi.org/10.1215/00267929-3570656}.

\bibitem[\citeproctext]{ref-sentiment}
Mohammad, Saif M., and Peter D. Turney. 2013. {``Crowdsourcing a
Word-Emotion Association Lexicon.''} \emph{Computational Intelligence}
29 (3): 436--65. \url{https://doi.org/10.1111/j.1467-8640.2012.00460.x}.

\bibitem[\citeproctext]{ref-rHere}
Müller, Kirill. 2020. \emph{Here: A Simpler Way to Find Your Files}.
\url{https://CRAN.R-project.org/package=here}.

\bibitem[\citeproctext]{ref-pathology}
Nyongesa, A. 2023. {``The Centre and Pathology: Postmodernist Reading of
Madness in the Oppressor in Contemporary Fiction.''} \emph{Cogent Arts
\& Humanities} 10 (1): 1--12.
https://doi.org/\url{https://doi.org/10.1080/23311983.2023.2249280}.

\bibitem[\citeproctext]{ref-rGg}
Pedersen, L., T. 2024. \emph{Ggraph: An Implementation of Grammar of
Graphics for Graphs and Networks}.
\url{https://ggraph.data-imaginist.com}.

\bibitem[\citeproctext]{ref-gutenberg}
{``Project Gutenberg.''} n.d. \url{https://www.gutenberg.org}.

\bibitem[\citeproctext]{ref-r}
R Core Team. 2022. \emph{R: A Language and Environment for Statistical
Computing}. Vienna, Austria: R Foundation for Statistical Computing.
\url{https://www.R-project.org/}.

\bibitem[\citeproctext]{ref-citeR}
---------. 2023. \emph{R: A Language and Environment for Statistical
Computing}. Vienna, Austria: R Foundation for Statistical Computing.
\url{https://www.R-project.org/}.

\bibitem[\citeproctext]{ref-youtube1}
Robertson, C. 2021. {``Text Tokenization in r.''}
\url{https://www.youtube.com/watch?v=9T-hr3jinTw}.

\bibitem[\citeproctext]{ref-rIgraph}
Silge, J., and D. Robinson. 2006. {``The Igraph Software Package for
Complex Network Research.''} \emph{InterJournal,*Complex Systems*},
1695. \url{https://igraph.org}.

\bibitem[\citeproctext]{ref-rWidy}
---------. 2022. \emph{Widyr: Widen, Process, Then Re-Tidy Data}.
\url{https://github.com/juliasilge/widyr}.

\bibitem[\citeproctext]{ref-rTidytext}
Silge, Julia, and David Robinson. 2016. {``Tidytext: Text Mining and
Analysis Using Tidy Data Principles in r.''} \emph{JOSS} 1 (3).
\url{https://doi.org/10.21105/joss.00037}.

\bibitem[\citeproctext]{ref-rGgplot2}
Wickham, Hadley. 2016. \emph{Ggplot2: Elegant Graphics for Data
Analysis}. Springer-Verlag New York.
\url{https://ggplot2.tidyverse.org}.

\bibitem[\citeproctext]{ref-rTidyverse}
Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy
D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019.
{``Welcome to the {tidyverse}.''} \emph{Journal of Open Source Software}
4 (43): 1686. \url{https://doi.org/10.21105/joss.01686}.

\bibitem[\citeproctext]{ref-rDplyr}
Wickham, Hadley, Romain François, Lionel Henry, and Kirill Müller. 2022.
\emph{Dplyr: A Grammar of Data Manipulation}.
\url{https://CRAN.R-project.org/package=dplyr}.

\bibitem[\citeproctext]{ref-rScales}
Wickham, Hadley, Thomas Lin Pedersen, and Dana Seidel. 2023.
\emph{Scales: Scale Functions for Visualization}.
\url{https://scales.r-lib.org}.

\bibitem[\citeproctext]{ref-rKnitr}
Xie, Yihui. 2014. {``Knitr: A Comprehensive Tool for Reproducible
Research in {R}.''} In \emph{Implementing Reproducible Computational
Research}, edited by Victoria Stodden, Friedrich Leisch, and Roger D.
Peng. Chapman; Hall/CRC.
\url{http://www.crcpress.com/product/isbn/9781466561595}.

\end{CSLReferences}

\end{document}

### 2024-04-04 paper/paper.qmd
update measurement

--- 2024-04-04 paper/paper.qmd
@@
 ---
-title: "The Linguistic Landscape of Stream-of-Consciousness Literature"
+title: "The Joyceless Landscape of Stream of Consciousness Literature" 
 subtitle: "Exploring Word Frequency, Sentiment Value and Mental Health Themes in the Works of Joyce, Woolf, Proust, Mansfield and Eliot from Project Gutenberg"
@@
 By analyzing these textual datasets through word frequency and sentiment analysis, I seek to pose and answer crucial questions: \textit{What are some important factors contributing to this relationship between mental health, disassociation and stream of consciousness?
-Moreover, how does this relationship vary differently across differnet demographics of authors, for instance, authors with different geographical locations and genders?} Understanding these dynamics is crucial in having an informed understanding of the West's late 19th to mid-20th century socio-political landscape, especially in regards to how authors and creative writers navigate and deal with then-taboo topics such as existential angst, mental health issues and disabilities.
+Moreover, how does this relationship vary differently across differnet demographics of authors, for instance, authors with different geographical locations and genders?} Understanding these dynamics is crucial in having an informed understanding of the West's late 19th to mid-20th century literature and even socio-political landscape, especially in regards to how authors and creative writers navigate and deal with then-taboo topics such as existential angst, mental health issues and disabilities.
 
-Thus, my estimand is the correlation between mental health-related words in SOC literature, their frequency and sentimental value as provided by @sentiment.
+Thus, my estimand is the correlation between mental health-related words in SOC literature, their frequency and sentiment value as provided by @sentiment.
 This is considered in terms of nine selected SOC novels only, namely Joyce's A Portrait of the Artist as a Young Man and Chamber Music; Woolf's Mrs Dalloway and Jacob's Room; Proust's Swann Way; Mansfield's Bliss and The Garden Party; and Eliot's The Waste Land and The Love Song of J.
@@
 
-To further understand the correlation between stream of consciousness novels and mental health themes, in [Introduction], our paper briefly discusses the nature of stream of consciousness literature, relevant authors and the works that we've chosen to analyze.
-Subsequently, in [Data] and [Results], we talk about the nature of the data obtained and analyze the results garnered from the data with suitable tables and charts.
+To further understand the correlation between stream of consciousness novels and mental health themes, in [Introduction], the paper briefly discusses the nature of stream of consciousness literature, relevant authors and the works that I've chosen to analyze.
+Subsequently, in [Data] and [Results], I talk about the nature of the data obtained and analyze the results garnered from the data with suitable tables and charts.
 Next, [Discussion] provides further insights and future areas of study.
 Finally, [Conclusion] summarizes our main findings.
-To complete the paper, [Appendix] clarifies how each variable within each dataset is generated and tables to accordingly demonstrate this.
+To complete the paper, [Appendix] clarifies how each variable within each dataset is generated with relevant tables to accordingly demonstrate this.
 
@@
 
-## Source Data
+Two central variables in this paper are:
 
-## Data Cleaning and Word Tokenization
+* **Word Frequency:** This variable captures the repetition of a single word (unigram), two-words combination (bigram) or three-words combination (trigram) throughout a SOC novel text, providing us with a thematic understanding of SOC literature.
+* **Sentiment Value:** This variable enables us to analyze how every word is usually perceived emotionally, whether it be positively or negatively.
+In additon,    in childlessness over time, encompassing the chosen time frame for the survey.
 
-```{r}
-#| echo: false
-#| message: false
-#| label: tbl-genderclass
-#| tbl-cap: Table of Number of Classes Students Considered for Regrade Requests by Students’ Gender
+Out of two variables used, the first one, 'Word Frequency' usually captured as `n` or `frequency` in datasets, is directly quantified through tokenizing the novel texts using the packages @rTidytext [@youtube1] [@youtube2].
+To do this, I first downloaded all nine novel texts from @gutenberg, and leveraged functions such as `unnest_tokens()` from @rTidytext to mine the texts, or separating it into individual words.
+Finally, I used `count()` to quantify the word frequency.
 
-```
+The second variable used, 'Sentiment Value', is based on three English-based, general-purpose  "word-emotion and word-polarity association lexicons", sourced from @sentiment's expansive research along with efforts from Finn Årup Nielsen and Bing Liu and collaborators [@rTidytext].
 
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
+The three general-purpose lexicon that contributes to my sentiment analysis are [@rTidytext]:
 
-```
+- 'AFINN' from Finn Årup Nielsen, which assigns a numerical value from '-5 to 5' 
+- 'bing' from Bing Liu and collaborators, which assigns if a word is 'positive' or negative'
+- 'nrc' from Saif Mohammad and Peter Turney, which assigns a core emotional value to a word, such as 'fear', 'anger', 'sadness' or 'trust'.
+
+In terms of measuring 'Sentiment Value', all three general-purpose lexicons are compiled through crowd-sourcing and directly surveying the public on how each word is emotionally perceived
+
+A survey sample of how the word 'startle' is compiled within the 'nrc' lexicon is presented below [@sentiment]:
+
+(@) Which word is closest in meaning (most related) to startle?
+- automobile
+- shake
+- honesty
+- entertain
+
+(@) How positive (good, praising) is the word startle?:
+- startle is not positive
+- startle is weakly positive
+- startle is moderately positive
+- startle is strongly positive
+
+(@) How negative (bad, criticizing) is the word startle?
+- startle is not negative
+- startle is weakly negative
+- startle is moderately negative
+- startle is strongly negative
+
+After the survey results are garnered, researchers average the answers to sort each surveyed word into pre-defined categories, specifically '-5 to 5' for 'AFINN', 'postive' or 'negative' for 'bing', and 'anger' or 'fear' for 'nrc'.
+With the continuous work of compiling these lexicons spanning years and decades [@sentiment] comes these functions: `get_sentiments("bing")` and `get_sentiments("nrc")` and `get_sentiments("afinn")` in which I can use `inner_join` to categorize my existing datasets of novel texts into their sentiment value.
+Systematic and data-driven, these measurement methods ensure that all lexicons faithfully reflects each word's emotionality.
+
+However, I do recognize how decontextualized and reductive this quantification of language can be.
+When it comes to understand such social and human artifacts as language or literary texts, much is dependent on their contextuality.
+As such, I will further discuss these weaknesses of the datasets under [Discussion].
+
+## Source Data: Data Cleaning and Word Tokenization
 

### 2024-04-04 paper/paper.qmd
graph updates

@@
 ---
-title: "The Joyceless Landscape of Stream of Consciousness Literature" 
+title: "Stream of Consciousness Literature: A Joyceless Linguistic Landscape" 
 subtitle: "Exploring Word Frequency, Sentiment Value and Mental Health Themes in the Works of Joyce, Woolf, Proust, Mansfield and Eliot from Project Gutenberg"
@@
 abstract: "This project focuses on understanding the language used by renowned stream of consciousness (SOC) authors James Joyce, Virginia Woolf, Marcel Proust, Katherine Mansfield and T.S Eliot.
-By conducting word frequency analysis and sentiment analysis of these authors' nine novels, I aim to uncover shared linguistic patterns and gain insights into the authors' mental states.
-Through the analysis on SOC literature, this paper thus attempts to offer insights into themes of self-identity, anxiety, disassociation and existential contemplation within Western society's from late 19th to mid-20th century.
+By conducting word frequency analysis and sentiment analysis of these authors' nine novels, this paper aims to uncover shared linguistic patterns and gain insights into the authors' mental states.
+Ultimatlely, this paper attempts to offer insights into themes of self-identity, anxiety, disassociation and existential contemplation within Western society and its literary circle from late 19th to mid-20th century.
 (add one sentence on main results)"
@@
 #| warning: false
-#| label: fig-sentiment
-#| fig-cap: "Sentiment Analysis of All SOC Novels"
+#| label: fig-sentiment-category
+#| layout-ncol: 2
+#| fig-cap: "Categorical Sentiment Analysis of All SOC Novel Texts by Authors"
+#| fig-subcap: ["Bing lexicon", "NRC lexicon"]
+
+bing_lexicon <- get_sentiments("bing") %>%
+  rename(sentiment_binary = sentiment)
 
 sentiment_combined <- combined_text |>
-  inner_join(get_sentiments("bing")) |>
+  inner_join(bing_lexicon) |>
   inner_join(get_sentiments("afinn")) |>
@@
 
+sentiment_category_binary <- sentiment_combined |>
+  group_by(author, sentiment_binary) |>
+  count() |>
+  group_by(author) |>
+  mutate(percentage = n / sum(n) * 100)
+
+# Plot sentiment values
+ggplot(sentiment_category_binary, aes(x = author, fill = sentiment_binary, y = percentage)) +
+  geom_bar(stat = "identity") +
+  labs(
+       x = "Author", y = "Percentage") +
+  scale_fill_manual(values = c("positive" = "gray80", "negative" = "gray20", "neutral" = "gray50")) + 
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
 sentiment_category <- sentiment_combined |>
-  inner_join(get_sentiments("bing")) |>
   group_by(author, sentiment) |>
@@
   group_by(author) |>
-  mutate(percentage = n / sum(n) * 100)
+  mutate(percentage = n / sum(n) * 100)  |>
+  group_by(author) |>
+  arrange(desc(percentage))
 
@@
        x = "Author", y = "Percentage") +
-  scale_fill_manual(values = c("positive" = "gray80", "negative" = "gray20", "neutral" = "gray50")) + 
+  scale_fill_manual(values = c("negative" = "gray1", "sadness" = "gray10", "anger" = "gray20", "disgust" = "gray30", "fear" = "gray40", "joy" = "gray50", "trust" = "gray60", "anticipation" = "gray70", "positive" = "gray80", "surprise" = "gray90")) +
   theme_minimal() +
@@
 #| label: fig-sentiment-number
-#| fig-cap: "Sentiment Analysis of All SOC Novels by Sentiment Value"
+#| fig-cap: "Numerical Sentiment Analysis of All SOC Novel Texts by Authors (AFINN lexicon)"
 
@@
 
-\newpage
 ## Combined Texts: Trends, Word Networks, Bigram and Trigram Analsis

### 2024-04-04 paper/paper.qmd
simulation updated

@@
 ---
-title: "Stream of Consciousness Literature: A Joyceless Linguistic Landscape" 
+title: "Stream of Consciousness Literature: A 'Joyceless' Linguistic Landscape" 
 subtitle: "Exploring Word Frequency, Sentiment Value and Mental Health Themes in the Works of Joyce, Woolf, Proust, Mansfield and Eliot from Project Gutenberg"
@@
 Thus, my estimand is the correlation between mental health-related words in SOC literature, their frequency and sentiment value as provided by @sentiment.
-This is considered in terms of nine selected SOC novels only, namely Joyce's A Portrait of the Artist as a Young Man and Chamber Music; Woolf's Mrs Dalloway and Jacob's Room; Proust's Swann Way; Mansfield's Bliss and The Garden Party; and Eliot's The Waste Land and The Love Song of J.
-Alfred Prufrock.
+This is considered in terms of nine selected SOC novels only, namely Joyce's \textit{A Portrait of the Artist as a Young Man and Chamber Music}; Woolf's \textit{Mrs Dalloway and Jacob's Room}; Proust's \textit{Swann Way}; Mansfield's \textit{Bliss and The Garden Party}; and Eliot's \textit{The Waste Land and The Love Song of J.
+Alfred Prufrock}.
 Through our analysis, we found that (percentage, number and data here, main results)...
@@
 ## Source Data: Data Cleaning and Word Tokenization
+
+Discuss the authenticity/validity of Project Gutenberg
 

### 2024-04-04 scripts/00-simulate_data.R

@@ -1,19 +1,56 @@
 #### Preamble ####
-# Purpose: Simulates
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Simulates datasets of word frequency
+# Author: Quang Mai
+# Date: 4 April 2024 
+# Contact: quangpono@gmail.com
 # License: MIT
-# Pre-requisites:
+# Pre-requisites: None
 # Any other information needed? No
 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+library(tibble)
 
 #### Simulate data ####
-# [...ADD CODE HERE...]
+set.seed(85) #ensure simulated data's reproducibility
 
+# self-generate random words since R doesn't have a built-in dataset of random English words
+random_words <- c("paper", "on", "five", "stream", "of", "consciousness", "authors", "and", "nine", "their", "most", "famous", "novels")
+# generate a dataset of selected novels
+novels <- c("A Portrait of the Artist as a Young Man", "Chamber Music", "Mrs Dalloway", "Jacob's Room", "Swann Way", "Bliss", "The Garden Party", "The Waste Land", "The Love Song of J. Alfred Prufrock")
+
+simulated_data <-
+  tibble(
+    # Randomly select novels from the generated dataset of chosen novels where the unique words are capped at 12
+    novel = sample(novels, 12, replace = TRUE, prob = NULL),  # Randomly select novels from the pre-defined novels dataset, repeatable since work can change
+    # Randomly select words from the randomly generated dataset of words that are capped at 12
+    word = sample(random_words, 12, replace = FALSE, prob = NULL),  # Randomly select words from the English words dataset, not repeated since this is the count of unique words
+    #use 1 through 870 to represent each count of a word since 870 is the maximum number of a unique word appears out of all nine novels
+    n = sample(1:870, 12, replace = TRUE))
+
+# Print the simulated data 
+print(simulated_data)
+
+#### Data Validation ####
+
+# Code referenced from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
+
+# Check that there are no more than 870 counts of a word #
+simulated_data |>
+  group_by(n) |>
+  count() |>
+  filter(n > n) |>
+  sum() == 0
+
[…]

### 2024-04-04 scripts/02-data_cleaning.R
update data cleaning

@@ -98,7 +98,7 @@ prufrock_words <- prufrock_clean |>
   filter(str_detect(word, "[:alpha:]")) |> 
   distinct()
 
-# Combine all texts to find the more general trend of word frequency
+# Combine all texts to find the more general trend of word frequency and sentiment value
 
 ## Calculate the correlations between all word combinations from all texts
 

### 2024-04-04 paper/paper.qmd
paper update

@@
 abstract: "This project focuses on understanding the language used by renowned stream of consciousness (SOC) authors James Joyce, Virginia Woolf, Marcel Proust, Katherine Mansfield and T.S Eliot.
-By conducting word frequency analysis and sentiment analysis of these authors' nine novels, this paper aims to uncover shared linguistic patterns and gain insights into the authors' mental states.
-Ultimatlely, this paper attempts to offer insights into themes of self-identity, anxiety, disassociation and existential contemplation within Western society and its literary circle from late 19th to mid-20th century.
+By conducting word frequency analysis and sentiment analysis on these authors' nine novels, this paper aims to uncover shared linguistic patterns and gain insights into the authors' mental states.
+Ultimately, this paper attempts to offer insights into themes of self-identity, anxiety, disassociation and existential contemplation within Western society and its literary circle from late 19th to mid-20th century.
 (add one sentence on main results)"
@@
 
+This includes words such as chapter titles, Roman numbers and personal names, affecting the integrity of data analysis ("Swann" being the most common word)
+
 ### Decontextualized Literature Works and Limiting Publication Editions
+
+Words are singled out and analyzed without context which could have affected their intended meanings, especially in such a complex genre such as stream of consciousness.
+The limiting novel editions also doesn't make sure that their literary integrity are maintained and the analysis might have missed important texts of other editions.
 
@@
 
+Since the novels are chosen based on their worldwide reception and canon, practical constraints like how all novels should be of similar length are ignored.
+This constraint can prevent one novel having more words than others, which can critically affect the integrity of the word frequency analysis, sentiment value and word networks where one novel dominates the others and skews the results.
+
+In addition, these authors work are expansive and so choosing a select few to bind them to the SOC genre can be limiting as the SOC genre in itself is already an amalgamation of different literary trends.
+Thus, this can affect the integrity of the datasets.
+
 ### Project Gutenberg's Focus on the Canon
+
+Along the search for other stream of consciousness authors to include in the generated datasets, only a few-those that are in the SOC canon, mostly white authors-are present in Project Gutenberg, without including lesser known SOC authors, which could have steared the data analysis, still, towards their preconception: being rueful and angst-ridden
 

### 2024-04-05 paper/paper.qmd
paper update

@@
 By conducting word frequency analysis and sentiment analysis on these authors' nine novels, this paper aims to uncover shared linguistic patterns and gain insights into the authors' mental states.
-Ultimately, this paper attempts to offer insights into themes of self-identity, anxiety, disassociation and existential contemplation within Western society and its literary circle from late 19th to mid-20th century.
+This paper attempts to offer insights into themes of self-identity, anxiety, disassociation and existential contemplation within Western society and its literary circle from late 19th to mid-20th century.
 (add one sentence on main results)"
@@
 * **Word Frequency:** This variable captures the repetition of a single word (unigram), two-words combination (bigram) or three-words combination (trigram) throughout a SOC novel text, providing us with a thematic understanding of SOC literature.
-* **Sentiment Value:** This variable enables us to analyze how every word is usually perceived emotionally, whether it be positively or negatively.
-In additon,    in childlessness over time, encompassing the chosen time frame for the survey.
+* **Sentiment Value:** This variable enables us to analyze how every word is usually perceived emotionally, whether it be a qualitative feeling or if it is conveyed through a numerical value.
 
-Out of two variables used, the first one, 'Word Frequency' usually captured as `n` or `frequency` in datasets, is directly quantified through tokenizing the novel texts using the packages @rTidytext [@youtube1] [@youtube2].
+Out of two variables used, the first one, 'Word Frequency' usually captured as `n` or `frequency` in datasets, is directly quantified through tokenizing the novel texts using the package `tidytext` and its function `unnest_tokens()` [@rTidytext].
 To do this, I first downloaded all nine novel texts from @gutenberg, and leveraged functions such as `unnest_tokens()` from @rTidytext to mine the texts, or separating it into individual words.
@@
 
-The second variable used, 'Sentiment Value', is based on three English-based, general-purpose  "word-emotion and word-polarity association lexicons", sourced from @sentiment's expansive research along with efforts from Finn Årup Nielsen and Bing Liu and collaborators [@rTidytext].
-
+The second variable used, 'Sentiment Value', is based on three English-based, general-purpose  "word-emotion and word-polarity association lexicons", sourced from Mohammad and Turney's expansive research along with efforts from Finn Årup Nielsen and Bing Liu and collaborators [@rTidytext] [@sentiment].
 The three general-purpose lexicon that contributes to my sentiment analysis are [@rTidytext]:
@@
 
-In terms of measuring 'Sentiment Value', all three general-purpose lexicons are compiled through crowd-sourcing and directly surveying the public on how each word is emotionally perceived
-
+In terms of measuring 'Sentiment Value', all three general-purpose lexicons are compiled through crowd-sourcing and directly surveying the public on how each word is emotionally perceived.
 A survey sample of how the word 'startle' is compiled within the 'nrc' lexicon is presented below [@sentiment]:

### 2024-04-05 paper/paper.qmd
paper update
update introduction

@@
 
-Thus, my estimand is the correlation between mental health-related words in SOC literature, their frequency and sentiment value as provided by @sentiment.
+Thus, my estimand is the correlation between mental health-related words in SOC literature, their frequency and sentiment value as provided by @Tidytext.
 This is considered in terms of nine selected SOC novels only, namely Joyce's \textit{A Portrait of the Artist as a Young Man and Chamber Music}; Woolf's \textit{Mrs Dalloway and Jacob's Room}; Proust's \textit{Swann Way}; Mansfield's \textit{Bliss and The Garden Party}; and Eliot's \textit{The Waste Land and The Love Song of J.
@@
 ### Generating Word Networks
+
+## Data Limitations
 

### 2024-04-06 scripts/01-download_data.R
Update 01-download_data.R

@@ -21,19 +21,19 @@ library(ggplot2)
 
 all_books <- gutenberg_metadata
 
-joyce_books <- all_books %>%
+joyce_books <- all_books |>
   filter(str_detect(author, 'Joyce, James'))
 
-woolf_books <- all_books %>%
+woolf_books <- all_books |>
   filter(str_detect(author, 'Woolf, Virginia'))
 
-mansfield_books <- all_books %>%
+mansfield_books <- all_books |>
   filter(str_detect(author, 'Mansfield, Katherine'))
 
-proust_books <- all_books %>%
+proust_books <- all_books |>
   filter(str_detect(author, 'Proust, Marcel'))
 
-ts_books <- all_books %>%
+ts_books <- all_books |>
   filter(str_detect(author, 'Eliot, T. S.'))
 
 

### 2024-04-06 scripts/01-download_data.R
Update 01-download_data.R

@@ -36,7 +36,6 @@ proust_books <- all_books |>
 ts_books <- all_books |>
   filter(str_detect(author, 'Eliot, T. S.'))
 
-
 combined_books <- rbind(joyce_books, woolf_books, dostoevsky_books, proust_books, ts_books, mansfield_books)
 
 # Download the more famous stream-of-consciousness work of each author:

### 2024-04-06 scripts/02-data_cleaning.R
Update 02-data_cleaning.R

@@ -4,7 +4,7 @@
 # Date: 30 March 2024 
 # Contact: quangpono@gmail.com
 # License: MIT
-# Pre-requisites: Read four raw datasets containing the novel texts from five authors, namely `portrait.csv`, `swann.csv`. `dalloway.csv`, `bliss.csv` and `prufrock.csv`
+# Pre-requisites: Read four raw datasets containing the literary texts from five authors, namely `portrait.csv`, `swann.csv`. `dalloway.csv`, `bliss.csv` and `prufrock.csv`
 
 #### Workspace setup ####
 library(tidyverse)

### 2024-04-21 paper/paper.qmd
update paper

@@
 library(textdata)
+library(arrow)
 
+portrait_clean <- read_parquet(here::here("data/analysis_data/portrait_clean.parquet"))
+portrait_words <- read_parquet(here::here("data/analysis_data/portrait_words.parquet"))
+swann_words <- read_parquet(here::here("data/analysis_data/swann_words.parquet"))
+dalloway_words <- read_parquet(here::here("data/analysis_data/dalloway_words.parquet"))
+bliss_words <- read_parquet(here::here("data/analysis_data/bliss_words.parquet"))
+prufrock_words <- read_parquet(here::here("data/analysis_data/prufrock_words.parquet"))
+combined_text <- read_parquet(here::here("data/analysis_data/combined_text.parquet"))
 portrait <- read_csv(here::here("data/raw_data/portrait.csv"))
-portrait_clean <- read_csv(here::here("data/analysis_data/portrait_clean.csv"))
 all_books <- read_csv(here::here("data/raw_data/all_books.csv"))
 combined_books <- read_csv(here::here("data/raw_data/combined_books.csv"))
-portrait_words <- read_csv(here::here("data/analysis_data/portrait_words.csv"))
-swann_words <- read_csv(here::here("data/analysis_data/swann_words.csv"))
-dalloway_words <- read_csv(here::here("data/analysis_data/dalloway_words.csv"))
-bliss_words <- read_csv(here::here("data/analysis_data/bliss_words.csv"))
-prufrock_words <- read_csv(here::here("data/analysis_data/prufrock_words.csv"))
-combined_text <- read_csv(here::here("data/analysis_data/combined_text.csv"))
 
@@
 The novel texts used for analysis were sourced from Project Gutenberg under the library `gutenbergr` [@rGuten] [@gutenberg].
-Data was generated, extracted and cleaned using the open-source statistical programming language R [@r], leveraging functions from `tidyverse` [@rTidyverse], `tidytext` [@rTidytext], `rmarkdown` [@rMark], `dplyr` [@rDplyr], `ggplot2` [@rGgplot2], `scales` [@rScales], `here` [@rHere], `igraph` [@rIgraph], `widyr` [@rWidy], `ggraph` [@rGg], `textdata` [@rTextdata], `tm` [@rTm] and `knitr` [@rKnitr].
+Data was generated, extracted and cleaned using the open-source statistical programming language R [@r], leveraging functions from `tidyverse` [@rTidyverse], `tidytext` [@rTidytext], `rmarkdown` [@rMark], `dplyr` [@rDplyr], `ggplot2` [@rGgplot2], `scales` [@rScales], `here` [@rHere], `igraph` [@rIgraph], `widyr` [@rWidy], `ggraph` [@rGg], `textdata` [@rTextdata], `tm` [@rTm], `here` [@here], `arrow` [@arrow], and `knitr` [@rKnitr].
 
@@
 ## Source Data: Data Cleaning and Word Tokenization
+
+Founded in December 1, 1971 by Michael S.
+Hart, Project Gutenberg exclusively publishes literature in the public domain within the United States.
+Typically, submissions to Project Gutenberg are digitized editions of printed books, the majority of which were published over 95 years ago.
+To confirm public domain status, authors can utilize the copy.pglaf.org website [@gutenberg].
+
+Since the archive is largely supported by voluntary works, its literature is also selected by volunteers.
+However, the archive doesn't accept copyrighted or other contemporary items, even if copyright or licensing permission would be granted.
+
+To submit work to the archive involves obtaining copyright, then scanning or harvesting to obtain images of a book's pages, then engaging in various hours of proofreading and formatting, and finally, ensuring the eBook is fully valid HTML, correctly spelled, and otherwise compliant with Project Gutenberg’s requirements,
+Collection development focuses on literature and other written works that have enduring value for readers.
+Selections are made by volunteers with diverse interests, and essentially all eligible submissions are welcome.
+
+COMMON POINTS OF FAILURE
+
+Crop marks or other printer's marks should not be used on any files.
+The book description should be in compliance with the rules listed here.
+Missing Pages
+Title missing on the front cover
+Incorrect pagination
+Books with typographical errors, such as misspellings or poor grammar.
+
+Project Gutenberg does not avoid difficult or unpopular topics, or topics for which societal views, or state-of-the-art knowledge, is vastly different from contemporary literature.
+
+Project Gutenberg does not accept copyrighted or other contemporary items, even if copyright or licensing permission would be granted (this includes the various “open” licenses).
+Instead, see the self-publishing portal described in the Submitting Your own Work How-To.
+In the past, a greater variety of non-public domain and non-literary works was added to Project Gutenberg, including copyrighted & other donated works, non-print formats, and different encodings and file types.
+Today, there are many outlets for such other items, and Project Gutenberg is entirely focused on works in the public domain.
+This includes literature, reference works, and variations such as children’s books and travelogues.
+
+Who should submit an eBook?
+
+Creating an eBook is a lot of work, and Project Gutenberg’s requirements are rather strict.
+If these steps seem daunting, you might be better off working with Distributed Proofreaders, where every volunteer contributes to a portion of the effort.
+Distributed Proofreaders has its own guidelines on submissions, as well as other policy, guidance, and community.
+
+Being a “solo” producer involves obtaining copyright, then scanning or harvesting to obtain images of a book’s pages, then engaging in many hours of proofreading and formatting, and, finally, ensuring the eBook is fully valid HTML, correctly spelled, and otherwise compliant with Project Gutenberg’s requirements.
+The detailed requirements are on the submission page, which also includes tools for automatically checking compliance.
+Any new submitter is invited to get in touch before to getting started.
+Contact the “copyright” or “whitewashers” team &emdash; email addresses are at the bottom of the copyright and upload sites.
+
+Copyright
+
+The first step in any eBook submission is to confirm that Project Gutenberg may legally distribute the eBook.
+Visit our Copyright How-To for details.
+Project Gutenberg will not accept any eBook without confirming copyright status.
+Generally, this means that a printed book has entered the public domain in the United States, typically because copyright has expired.
+
+File Formats
+
+The master format for nearly all new Project Gutenberg eBooks is HTML.
+Project Gutenberg insists the HTML is fully valid, and any cascading style sheets (CSS) in the HTML are part of the published standards by the World Wide Web consortium (W3C).
+
+Whenever possible, Project Gutenberg also requires a plain text version of an eBook.
+We stress the inclusion of plain text because of its longevity: Project Gutenberg includes numerous text files that are over 30 years old.
+In that time, dozens of widely used file formats have come and gone.
+Text is accessible on all computers, and is also insurance against future obsolescence.
+
+The only times when Project Gutenberg distributes an eBook without a plain text version are when plain text is impossible or impractical — for example, for our movies and MP3 audio files, and for some of our mathematical works.
+
+This isn’t as hard as it sounds, if you start with HTML, and continuously confirm validity using the W3C’s online validator (see the upload page for details).
+Modern eBook producers almost always start with (valid) HTML, and then derive plain text from the HTML.
+These two “master” formats, text and HTML, are then submitted to Project Gutenberg.
+Automated tools then create derivative formats, including epub and mobi (the common e-reader formats).
+
+Hand-crafted epub and mobi files are not accepted, currently.
+This limitation is mainly to allow easy editing when fixes are applied, because application of fixes occurs many times over the lifetime of every eBook.
+Project Gutenberg tries to limit the number of master formats, and then automate derived formats, to facilitate continuous improvement of items in the collection.
+
+A small number of new eBooks utilize LaTeX or TEI as the master format - mostly those including lots of mathematical notation.
+Other less frequently utilized formats include ReSstructured Text (RST), Rich Data Format (RDF), and a few others.
+Note that PDF, Word, and other word processor formats are not utilized as master formats, because they do not convert easily to valid HTML.
+In addition, they are much more challenging to fix/update than HTML and plain text.
 

### 2024-04-21 scripts/02-data_cleaning.R

@@ -4,7 +4,7 @@
 # Date: 30 March 2024 
 # Contact: quangpono@gmail.com
 # License: MIT
-# Pre-requisites: Read four raw datasets containing the literary texts from five authors, namely `portrait.csv`, `swann.csv`. `dalloway.csv`, `bliss.csv` and `prufrock.csv`
+# Pre-requisites: Read four raw datasets containing the novel texts from five authors, namely `portrait.csv`, `swann.csv`. `dalloway.csv`, `bliss.csv` and `prufrock.csv`
 
 #### Workspace setup ####
 library(tidyverse)
@@ -16,6 +16,7 @@ library(dplyr)
 library(scales)
 library(igraph)
 library(widyr)
+library(arrow)
 
 #### Clean data ####
 portrait <- read_csv("data/raw_data/portrait.csv")
@@ -109,14 +110,13 @@ combined_text <-
   filter(str_detect(word, "[:alpha:]")) 
 
 #### Save data ####
-write_csv(portrait_clean, "data/analysis_data/portrait_clean.csv")
-write_csv(portrait_words, "data/analysis_data/portrait_words.csv")
-write_csv(swann_words, "data/analysis_data/swann_words.csv")
-write_csv(dalloway_words, "data/analysis_data/dalloway_words.csv")
-write_csv(bliss_words, "data/analysis_data/bliss_words.csv")
-write_csv(prufrock_words, "data/analysis_data/prufrock_words.csv")
-write_csv(combined_text, "data/analysis_data/combined_text.csv")
-
+write_parquet(portrait_clean, "data/analysis_data/portrait_clean.parquet")
+write_parquet(portrait_words, "data/analysis_data/portrait_words.parquet")
+write_parquet(swann_words, "data/analysis_data/swann_words.parquet")
+write_parquet(dalloway_words, "data/analysis_data/dalloway_words.parquet")
+write_parquet(bliss_words, "data/analysis_data/bliss_words.parquet")
+write_parquet(prufrock_words, "data/analysis_data/prufrock_words.parquet")
+write_parquet(combined_text, "data/analysis_data/combined_text.parquet")
 
 
 

### 2024-04-22 paper/paper.qmd
data section and data testing updated

--- 2024-04-05 paper/paper.qmd
@@
 library(textdata)
+library(arrow)
+library(kableExtra)
 
+portrait_clean <- read_parquet(here::here("data/analysis_data/portrait_clean.parquet"))
+portrait_words <- read_parquet(here::here("data/analysis_data/portrait_words.parquet"))
+swann_words <- read_parquet(here::here("data/analysis_data/swann_words.parquet"))
+dalloway_words <- read_parquet(here::here("data/analysis_data/dalloway_words.parquet"))
+bliss_words <- read_parquet(here::here("data/analysis_data/bliss_words.parquet"))
+prufrock_words <- read_parquet(here::here("data/analysis_data/prufrock_words.parquet"))
+combined_text <- read_parquet(here::here("data/analysis_data/combined_text.parquet"))
 portrait <- read_csv(here::here("data/raw_data/portrait.csv"))
-portrait_clean <- read_csv(here::here("data/analysis_data/portrait_clean.csv"))
 all_books <- read_csv(here::here("data/raw_data/all_books.csv"))
 combined_books <- read_csv(here::here("data/raw_data/combined_books.csv"))
-portrait_words <- read_csv(here::here("data/analysis_data/portrait_words.csv"))
-swann_words <- read_csv(here::here("data/analysis_data/swann_words.csv"))
-dalloway_words <- read_csv(here::here("data/analysis_data/dalloway_words.csv"))
-bliss_words <- read_csv(here::here("data/analysis_data/bliss_words.csv"))
-prufrock_words <- read_csv(here::here("data/analysis_data/prufrock_words.csv"))
-combined_text <- read_csv(here::here("data/analysis_data/combined_text.csv"))
 
@@
 As such, this paper has mined the texts of a total of nine novels from the volunteer archive, Project Gutenberg, to examine the mental health themes of famous stream of consciousness authors, namely by Joyce, Woolf, Proust, Mansfield and Eliot, from the modernist era of literature, spanning from late 19th century to mid-20th century [@gutenberg].
-(more stats and data mentioned here)
 
@@
 
-Thus, my estimand is the correlation between mental health-related words in SOC literature, their frequency and sentiment value as provided by @Tidytext.
+Thus, the estimand is the correlation between mental health-related words in SOC literature, their frequency and sentiment value as provided by @rTidytext.
 This is considered in terms of nine selected SOC novels only, namely Joyce's \textit{A Portrait of the Artist as a Young Man and Chamber Music}; Woolf's \textit{Mrs Dalloway and Jacob's Room}; Proust's \textit{Swann Way}; Mansfield's \textit{Bliss and The Garden Party}; and Eliot's \textit{The Waste Land and The Love Song of J.
@@
 The novel texts used for analysis were sourced from Project Gutenberg under the library `gutenbergr` [@rGuten] [@gutenberg].
-Data was generated, extracted and cleaned using the open-source statistical programming language R [@r], leveraging functions from `tidyverse` [@rTidyverse], `tidytext` [@rTidytext], `rmarkdown` [@rMark], `dplyr` [@rDplyr], `ggplot2` [@rGgplot2], `scales` [@rScales], `here` [@rHere], `igraph` [@rIgraph], `widyr` [@rWidy], `ggraph` [@rGg], `textdata` [@rTextdata], `tm` [@rTm] and `knitr` [@rKnitr].
+Data was generated, extracted and cleaned using the open-source statistical programming language R [@r], leveraging functions from `tidyverse` [@rTidyverse], `tidytext` [@rTidytext], `rmarkdown` [@rMark], `dplyr` [@rDplyr], `ggplot2` [@rGgplot2], `scales` [@rScales], `here` [@rHere], `igraph` [@rIgraph], `widyr` [@rWidy], `ggraph` [@rGg], `textdata` [@rTextdata], `tm` [@rTm], `here` [@here], `kableExtra` [@kableExtra], `arrow` [@arrow], and `knitr` [@rKnitr].
 
@@
 
-## Source Data: Data Cleaning and Word Tokenization
+## Source Data: Project Gutenberg
 
-Discuss the authenticity/validity of Project Gutenberg
+Founded in December 1, 1971 by Michael S.
+Hart, Project Gutenberg exclusively publishes literature in the public domain within the United States.
+Typically, submissions to Project Gutenberg are digitized editions of printed books, the majority of which were published over 95 years ago.
+To confirm public domain status, authors can use the copy.pglaf.org website [@gutenberg].
 
-### Word Count
+The archive depends mainly on volunteer efforts for support, and its literature is chosen and proofread by volunteers.
+However, the archive doesn't take copyrighted or modern items, even with permission.
+To submit work to the archive, you need to get copyright, scan or harvest images of book pages, spend hours proofreading and formatting, and ensure the eBook meets Project Gutenberg’s requirements, including being valid HTML, correctly spelled, and compliant  with Project Gutenberg’s requirement.
+Common points of failures for submitted works include:
+
+- Crop marks or other printer's marks should not be used on any files.
+- The book description should be in compliance with the rules listed here.
+- Missing Pages
+- Title missing on the front cover
+- Incorrect pagination
+- Books with typographical errors, such as misspellings or poor grammar.
+
+In terms of format, Most new Project Gutenberg eBooks are in HTML format.
+Project Gutenberg will check each work's the HTML validity using the W3C’s online validator.
+The archive also askss for a plain text version whenever possible.
+This is because plain text has been around for a long time, is widely accessible on all devices and ensures long-term usability.
+PDF, Word, and other word processor formats are not used as master formats because they are harder to convert to HTML and update  [@gutenberg].
+
+With these points being addressed and the impartial nature of the archive being publicly disclosed on its website, the source data in which I garner my literary texts are selected with strict quality and accuracy.
+However, there are potential spaces for biases such as selection bias where Project Gutenberg primarily hosts older works that have entered the public domain.
+Thus, archived works might gear towards certain genres, authors, or time periods, potentially limiting the diversity of the datasets for analysis, which will be further discussed below [@gutenberg].
+
+### Data Cleaning and Word Tokenization
 
@@
          "Author" = author) |>
-  head(5) |>
+  head(3) |>
   kable(
@@
     booktabs = TRUE
-  )
+  ) |>
+  kable_styling(latex_options = c("striped", "hold_position"), 
+                bootstrap_options = "striped", 
+                full_width = FALSE)
 
 ```
+
+To garner the word count of SOC literature  using the library `gutenbergr`, I first downloaded the index of each respective SOC author's body of works using each author's name as outlined in Project Gutenberg database, namely Joyce, Mansfield, Eliot, Woolf, and Proust [@rGuten] ([Appendix]).
+Then, to create my own datasets containing the word frequency within each SOC author's texts, I used the identification number of the desired literary work and the function `gutenberg_download` to upload the HTML texts into individual csv-and later on, parquet-datasets, as observed in @tbl-joyce for James Joyce.
 
@@
 #| label: tbl-joyce-token
-#| tbl-cap: "An Examplary Table Containing Tokenzied Novel Text (James Joyce)"
+#| tbl-cap: "Tokenized Novel Text & Word Count (James Joyce)"
+#| tbl-subcap: 
+#|   - "Tokenized Text"
+#|   - "Word Count"
+#| layout-ncol: 2
 #| warning: false
+#| message: false
 
@@
          "Word" = word) |>
-  tail(5) |>
+  tail(3) |>
   kable(
@@
     booktabs = TRUE
-  )
-
-```
-\newpage
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-joyce-words
-#| tbl-cap: "An Examplary Table Containing Word Count of Each Word within Novel Texts (James Joyce)"
-#| warning: false
+  ) |>
+  kable_styling(latex_options = c("striped", "hold_position"), 
+                bootstrap_options = "striped", 
+                full_width = FALSE)
 
@@
     booktabs = TRUE
-  )
+  ) |>
+  kable_styling(latex_options = c("striped", "hold_position"), 
+                bootstrap_options = "striped", 
+                full_width = FALSE)
+
+```
+
+Next, observable in @tbl-joyce-token, to measure the word frequency of each literary text, I first tokenized or separated the texts into individual chunks of words using `unnest_tokens()` [@rTidytext].In addition, to account for stop words like, a, of, and, etc., I leveraged the dataset `stop_words`and the function `anti_join` from @rTidytext to exclude all stop words from my self-generated datasets of literary texts, ensuring data validity and meaningful analysis.
+
+Then, to quantify the number of times each word appears in each text, I used `count()` (@tbl-joyce-token).
+The result manifests into @fig-count.
+The figure demonstrates that beside the top frequencies of personal names, most of the words are abstract nouns that revolve around the nature of time, and life, e.g.
+god (appearing 194 times in Joyce's), time (appearing in 4 authors' texts, 121 times for Joyce's, 77 times for Woolf's, 421 times for Proust's, 23 times for Eliot), mind (108 times in Joyce's and 196 times in Proust's), life (appearing 266 times in Proust's and 131 times in Joyce's), etc.
+This further exemplifies the sense of disassociation, temporality and non-linearity commonly associated to stream-of-consciousness literature and more broadly with mental health themes ([@flow], [@pathology], [@resampling]).
+
+\newpage
+
+```{r}
+#| eval: true
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-count
+#| layout-ncol: 2
+#| fig-cap: "Comparative Analysis of Top 20 Word Frequencies from Famous SOC Authors"
+#| fig-subcap: ["James Joyce", "Virginia Woolf", "Marcel Proust", "T.S. Eliot", "Katherine Mansfield"]
+
+portrait_words |> head(20) |>
+ggplot(aes(x = n, y = reorder(word, n))) +
+  geom_bar(stat = "identity", fill = "darkgrey") +
+  geom_text(aes(label = n), hjust = -0.3, vjust = 0.5, color = "grey40", size = 3) +
+  labs(x = "Frequency", y = "Word")
+
+dalloway_words |> head(20) |>
+ggplot(aes(x = n, y = reorder(word, n))) +
+  geom_bar(stat = "identity", fill = "darkgrey") +
+  geom_text(aes(label = n), hjust = -0.3, vjust = 0.5, color = "grey40", size = 3) +
+  labs(x = "Frequency", y = "Word")
+
+swann_words |> head(20) |>
+ggplot(aes(x = n, y = reorder(word, n))) +
+  geom_bar(stat = "identity", fill = "darkgrey") +
+  geom_text(aes(label = n), hjust = -0.3, vjust = 0.5, color = "grey40", size = 3) +
+  labs(x = "Frequency", y = "Word")
+
+prufrock_words |> head(20) |>
+ggplot(aes(x = n, y = reorder(word, n))) +
+  geom_bar(stat = "identity", fill = "darkgrey") +
+  geom_text(aes(label = n), hjust = -0.3, vjust = 0.5, color = "grey40", size = 3) +
+  labs(x = "Frequency", y = "Word")
+
+bliss_words |> head(20) |>
+ggplot(aes(x = n, y = reorder(word, n))) +
+  geom_bar(stat = "identity", fill = "darkgrey") +
+  geom_text(aes(label = n), hjust = -0.3, vjust = 0.5, color = "grey40", size = 3) +
+  labs(x = "Frequency", y = "Word")
 
@@
   select(-n) |>
-  pivot_wider(names_from = author, values_from = proportion)
+  pivot_wider(names_from = author, values_from = proportion) |>
+  filter(if_all(everything(), ~is.na(.) | !is.na(.)))
 
@@
     booktabs = TRUE
-  )
+  ) |>
+    kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
 
@@
 
-### Generating Word Networks
+To accurately compare word frequency among SOC authors, from @tbl-frequency, I convert each word frequency into a percentage.
+This is done by using `mutate` to calculate the word frequency of each word relative to the total word frequency of each author.
+Due to the different number of words present in each author's literary works, this method accounts for the inconsistencies of between each author's word count.
 
-## Data Limitations
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-boxplot 
+#| tbl-cap: "Overall Boxplot Summary Statistics of all Word Frequency Percentages (AFINN Sentiment)"
+#| warning: false
 
-# Model
+# Calculate summary statistics
+frequency_long |>
+  group_by(frequency) %>%
+  summarise(
+    min = min(value),
+    q1 = quantile(value, probs = 0.25),
+    median = median(value),
+    mean = mean(value),
+    q3 = quantile(value, probs = 0.75),
+    max = max(value),
+    sd = sd(value)
+  ) |>
+    head(10) |>
+  kable(
+    col.names = c("Frequency", "Min", "Q1", "Median", "Mean", "Q3", "Max", "SD"),
+    booktabs = TRUE
+  ) |>
+  kable_styling(latex_options = c("striped", "hold_position"), 
+                bootstrap_options = "striped", 
+                full_width = FALSE)
 
-The goal of our modelling strategy is twofold. Firstly,...
+```
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+@fig-sentiment-frequency and @tbl-boxplot captures the data garnered.
+For the NRC index, for all authors, the lower word frequency percentage ranges from 0.0000819 to 0004255 and are largely words that carry a mixture of emotional connotations such as, joy, surprise, positive, disgust and anticipation.
+However, the higher word frequency percentage, ranging from 0,0004255 to 0.0008511, mostly contains words with negative connotations like fear and anger, that weighs a -1.875 median on a scale between -5 to 5 in the AFINN index.
+Similarly, for the Bing index, the higher word frequency ranging from 0,0004255 to 0.0008511 percentages largely contains words that carry negative connotations, that weighs approximately a -1.875 median on a scale between -5 to 5 in the AFINN index.
 
-## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-We expect a negative relationship between average household income and the number of children per child care space by ward. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
-# Results
-
-## The Dominant Vocabulary of SOC Literature
+Overseeing the entire dataset, @tbl-boxplot demonstrates that, for all distinct word frequency percentages, the median sentiment value is -2 on the scale between -5 and 5.
+More specifically, words that appear most frequently, or words with the frequency percentage of 0.0008511, have the lowest mean sentiment value-or the most negative connotation of-2.33 on the scale between -5 and 5 in the AFINN sentiment index.
+This summary statistics of the gathered data only further shows how the SOC literary landscape mostly associates with words that carry more negative or even extreme connotations often associated with mental health themes, and topics such as depression, disassociation or schizophrenia [@pathology].
 
@@
 #| warning: false
-#| label: fig-count
+#| label: fig-sentiment-frequency
 #| layout-ncol: 2
-#| fig-cap: "Comparative Analysis of Top 20 Word Frequencies from Famous SOC Authors"
-#| fig-subcap: ["James Joyce", "Virginia Woolf", "Marcel Proust", "T.S. Eliot", "Katherine Mansfield"]
+#| fig-cap: "Overall Word Frequency and Sentiment Analysis"
+#| fig-subcap: ["Bing lexicon", "NRC lexicon"]
 
-portrait_words |> head(20) |>
-ggplot(aes(x = n, y = reorder(word, n))) +
-  geom_bar(stat = "identity", fill = "darkgrey") +
-  labs(x = "Frequency", y = "Word")
+# Reshape data for plotting
+bing_lexicon <- get_sentiments("bing") %>%
+  rename(sentiment_binary = sentiment)
 
-dalloway_words |> head(20) |>
-ggplot(aes(x = n, y = reorder(word, n))) +
-  geom_bar(stat = "identity", fill = "darkgrey") +
-  labs(x = "Frequency", y = "Word")
+frequency_long <- pivot_longer(frequency, 
+                               cols = c("James Joyce", "Marcel Proust", "Virginia Woolf", "Katherine Mansfield", "T.S. Eliot"), 
+                               names_to = "author", values_to = "frequency") |>
+  inner_join(bing_lexicon) |>
+  inner_join(get_sentiments("afinn")) |>
+  inner_join(get_sentiments("nrc"))
 
-swann_words |> head(20) |>
-ggplot(aes(x = n, y = reorder(word, n))) +
-  geom_bar(stat = "identity", fill = "darkgrey") +
-  labs(x = "Frequency", y = "Word")
+frequency_long <- na.omit(frequency_long)
 
-prufrock_words |> head(20) |>
-ggplot(aes(x = n, y = reorder(word, n))) +
-  geom_bar(stat = "identity", fill = "darkgrey") +
-  labs(x = "Frequency", y = "Word")
+ggplot(frequency_long, aes(x = frequency, y = value, fill = sentiment_binary)) +
+  geom_boxplot() +
+  labs(title = "Boxplot of Sentiment Scores by Author (Bing)",
+       x = "Word Frequency Percentage",
+       y = "Sentiment Value (AFINN)",
+       fill = "Sentiment Associated") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
-bliss_words |> head(20) |>
-ggplot(aes(x = n, y = reorder(word, n))) +
-  geom_bar(stat = "identity", fill = "darkgrey") +
-  labs(x = "Frequency", y = "Word")
+ggplot(frequency_long, aes(x = frequency, y = value, fill = sentiment)) +
+  geom_boxplot() +
+  labs(title = "Boxplot of Sentiment Scores by Author (NRC)",
+       x = "Word Frequency Percentage",
+       y = "Sentiment Value (AFINN)",
+       fill = "Sentiment Associated") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
 ```
-\newpage
+
+## Data Limitations
+
+### Lack of Thorough Word Cleaning
+
+With the large amount of different literary texts need processing by myself, it's hard to truly conduct thorough word processing.
+Certain textual elements, while can have their frequencies accounted for, can't have sentiment analysis due to their factual nature.
+These textual elements might include things like chapter titles, Roman numbers and personal names, which can strongly skew and affect the data integrity.
+As observed in @fig-count, "Swann", "Stephen" or "Jacob" are the most repeated words in Proust's, Joyce's and Woolf's, proving this data limitation.
+
+### Potential Measurement Errors
+
+The actual raw text data sourced from Project Gutenberg may be susceptible to measurement errors due to various factors, such as selection bias towards older, Western and canonical texts, social desirability bias that edits out certain political elements, memory lapses, or misinterpretations of literary texts by voluntary proofreaders.
+While the archive has strict guidelines to minimize these errors, they cannot be entirely eliminated.
+
+### Uneven Text Length
+
+I chose literary texts based on their global reception and canonical status within the stream-of-consciousness genre.
+This decision involved forgoing certain practical considerations, such as ensuring similar text lengths across all works by each author.
+While this approach enables a thorough exploration of word frequency analysis, sentiment value, and word networks within the SOC genre, it also has the potential to emphasize the frequencies of certain words from specific authors more than others due to the longer text length of those authors.
+This could critically affect the integrity of the data analysis.
+
+## Biased Author Selection
+
+Furthermore, SOC literature is expansive, spanning a wide range of themes, styles, and narrative techniques.
+By exclusively representing the genre through a select group of canonical Western authors, I risk constraining the genre's diversity.
+By oversimplifying the complexity of the the dynamic literary genre, I also affect the integrity of the datasets and its correlation to mental health themes.
+
+# Results
+
+## The Dominant Vocabulary of SOC Literature
 
@@
 
+pathological consequences of the constant
+shifts of identity in migrant characters. Most postcolonial studies contend that
+marginalization occasioned by the dominant group result in neurotic conditions in
+members of the minority group.
+
 The novels' linguistic patterns reflect the socio-political landscape of the Western hemisphere, from  late 19th century to the mid-20th century.
@@
 
+While these datasets aim to illuminate on the general time trends of childlessness with each variable measured, they are often de-contextualized for the same purpose.
+For instance, in Table 6, there is no clarifying note on whether people who declared themselves as ‘divorced’ are divorced once or more.
+This can be extremely narrowing, and exclude nuances when analyzing the contexts behind childlessness.
 ### Uneven Novel Length and Categorization of Authors
@@
 
-## Model Details {#sec-model-details}
-
-## Posterior predictive check
-
-## Diagnostics
-
 \newpage

### 2024-04-22 scripts/00-simulate_data.R

@@ -26,8 +26,8 @@ simulated_data <-
     novel = sample(novels, 12, replace = TRUE, prob = NULL),  # Randomly select novels from the pre-defined novels dataset, repeatable since work can change
     # Randomly select words from the randomly generated dataset of words that are capped at 12
     word = sample(random_words, 12, replace = FALSE, prob = NULL),  # Randomly select words from the English words dataset, not repeated since this is the count of unique words
-    #use 1 through 870 to represent each count of a word since 870 is the maximum number of a unique word appears out of all nine novels
-    n = sample(1:870, 12, replace = TRUE))
+    #use 1 through 602 to represent each count of a word since 602 is the maximum number of a unique word appears out of all nine novels
+    n = sample(1:602, 12, replace = TRUE))
 
 # Print the simulated data 
 print(simulated_data)

### 2024-04-22 scripts/01-download_data.R

@@ -51,7 +51,7 @@ swann_way <- gutenberg_download(c('7178')) |>
 
 mrs_dalloway <- gutenberg_download(c('63107', '5670')) |>
   mutate(book = ifelse(gutenberg_id == 63107, "Mrs Dalloway", "Jacob's Room")) |>
-  mutate(author = ifelse(gutenberg_id == 63107, "Virigina Woolf", "Virginia Woolf")) |>
+  mutate(author = ifelse(gutenberg_id == 63107, "Virginia Woolf", "Virginia Woolf")) |>
   slice(-(1:25))
 
 prufrock <- gutenberg_download(c('1321', '1459')) |>
@@ -60,7 +60,7 @@ prufrock <- gutenberg_download(c('1321', '1459')) |>
   slice(-(1:35))
 
 bliss <- gutenberg_download(c('44385', '1429')) |>
-  mutate(book = ifelse(gutenberg_id == 1321, "Bliss", "The Garden Party")) |>
+  mutate(book = ifelse(gutenberg_id == 44385, "Bliss", "The Garden Party")) |>
   mutate(author = ifelse(gutenberg_id == 63107, "Katherine Mansfield", "Katherine Mansfield")) |>
   slice(-(1:34))
 

### 2024-04-22 scripts/03-test_data.R

@@ -1,15 +1,61 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Tests for the six cleaned datasets to make sure that they can be analyzed
+# Author: Quang Mai
+# Date: 21 April 2024
+# Contact: quangpono@gmail.com
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Pre-requisites: combined_text.parquet, portrait_words.parquet, swann_words.parquet, dalloway_words.parquet, bliss_words.parquet, prufrock_words.parquet, combined_text.parquet
+# Any other information needed? No
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+library(arrow)
+
+portrait_words <- read_parquet(here::here("data/analysis_data/portrait_words.parquet"))
+swann_words <- read_parquet(here::here("data/analysis_data/swann_words.parquet"))
+dalloway_words <- read_parquet(here::here("data/analysis_data/dalloway_words.parquet"))
+bliss_words <- read_parquet(here::here("data/analysis_data/bliss_words.parquet"))
+prufrock_words <- read_parquet(here::here("data/analysis_data/prufrock_words.parquet"))
+combined_text <- read_parquet(here::here("data/analysis_data/combined_text.parquet"))
+
 
 #### Test data ####
+# check that there are nine unique ids to represent 9 books
+length(unique(combined_text$gutenberg_id)) == 9
+length(unique(combined_text$book)) == 9
+
+# check that there are five unique author names to represent five authors
+length(unique(combined_text$author)) == 5
+
+# check that the number of words in all nine texts is only from 1 to 167,516
+length(combined_text$word) == 167516
+
+# check that the number of unique words in Mansfield's texts is 7829
+length(bliss_words$word) == 7829
+
+# check that the number of unique words in Proust's texts is 7829
+length(swann_words$word) == 12203
+
+# check that the number of unique words in Eliot's texts is 2350
+length(prufrock_words$word) == 2350
+
+# check that the number of unique words in Joyce's texts is 8927
[…]

### 2024-04-22 paper/paper.qmd
update results

@@
 
-Thus, the estimand is the correlation between mental health-related words in SOC literature, their frequency and sentiment value as provided by @rTidytext.
+Thus, the estimand is the correlation between mental health-related themes and words in SOC literature, their frequency and sentiment value as provided by @rTidytext.
 This is considered in terms of nine selected SOC novels only, namely Joyce's \textit{A Portrait of the Artist as a Young Man and Chamber Music}; Woolf's \textit{Mrs Dalloway and Jacob's Room}; Proust's \textit{Swann Way}; Mansfield's \textit{Bliss and The Garden Party}; and Eliot's \textit{The Waste Land and The Love Song of J.
@@
 
-However, I do recognize how decontextualized and reductive this quantification of language can be.
-When it comes to understand such social and human artifacts as language or literary texts, much is dependent on their contextuality.
+However, I do recognize how reductive this quantification of language can be.
+When it comes to understand such social and human artifacts as language or literary texts, much is dependent on their context.
 As such, I will further discuss these weaknesses of the datasets under [Discussion].
@@
 
-The archive depends mainly on volunteer efforts for support, and its literature is chosen and proofread by volunteers.
-However, the archive doesn't take copyrighted or modern items, even with permission.
-To submit work to the archive, you need to get copyright, scan or harvest images of book pages, spend hours proofreading and formatting, and ensure the eBook meets Project Gutenberg’s requirements, including being valid HTML, correctly spelled, and compliant  with Project Gutenberg’s requirement.
+The archive relies heavily on volunteers for support and content selection, but it does not accept copyrighted or modern works, even with permission.
+To submit work, individuals must obtain copyright clearance, scan or capture book images, dedicate hours to proofreading and formatting, and ensure compliance with Project Gutenberg’s guidelines, including valid HTML and correct spelling.
 Common points of failures for submitted works include:
@@
 
-In terms of format, Most new Project Gutenberg eBooks are in HTML format.
-Project Gutenberg will check each work's the HTML validity using the W3C’s online validator.
-The archive also askss for a plain text version whenever possible.
-This is because plain text has been around for a long time, is widely accessible on all devices and ensures long-term usability.
-PDF, Word, and other word processor formats are not used as master formats because they are harder to convert to HTML and update  [@gutenberg].
+Most new Project Gutenberg eBooks are in HTML format, which is validated using the W3C’s online validator.
+Additionally, the archive prefers plain text versions whenever possible due to their universal accessibility, and long-term usability.
+PDF, Word, and other word processor formats are avoided as master formats due to the challenges involved in converting them to HTML and updating them [@gutenberg].
 
@@
 
+\newpage
+
 ```{r}
@@
 #| label: tbl-joyce-token
-#| tbl-cap: "Tokenized Novel Text & Word Count (James Joyce)"
-#| tbl-subcap: 
-#|   - "Tokenized Text"
-#|   - "Word Count"
+#| tbl-cap: Tokenized Novel Text & Word Count after Tokenization (James Joyce)
+#| tbl-subcap: ["Tokenized Text", "Word Count"]
 #| layout-ncol: 2
@@
 
+portrait_clean <- portrait_clean |>
+  select(-book)
+
 portrait_clean |>
   rename("Book ID" = gutenberg_id,
-         "Book" = book,
          "Author" = author,
@@
   kable(
-    col.names = c("Book ID", "Book", "Author", "Word"),
+    col.names = c("Book ID", "Author", "Word"),
     booktabs = TRUE
@@
 god (appearing 194 times in Joyce's), time (appearing in 4 authors' texts, 121 times for Joyce's, 77 times for Woolf's, 421 times for Proust's, 23 times for Eliot), mind (108 times in Joyce's and 196 times in Proust's), life (appearing 266 times in Proust's and 131 times in Joyce's), etc.
-This further exemplifies the sense of disassociation, temporality and non-linearity commonly associated to stream-of-consciousness literature and more broadly with mental health themes ([@flow], [@pathology], [@resampling]).
+This further exemplifies the sense of disassociation, temporality and non-linearity commonly associated to stream-of-consciousness literature and more broadly with mental health themes [@flow] [@pathology] [@resampling].
 
@@
 #| warning: false
+
+#reshape the frequency datasets and add sentiment values
+bing_lexicon <- get_sentiments("bing") %>%
+  rename(sentiment_binary = sentiment)
+
+frequency_long <- pivot_longer(frequency, 
+                               cols = c("James Joyce", "Marcel Proust", "Virginia Woolf", "Katherine Mansfield", "T.S. Eliot"), 
+                               names_to = "author", values_to = "frequency") |>
+  inner_join(bing_lexicon) |>
+  inner_join(get_sentiments("afinn")) |>
+  inner_join(get_sentiments("nrc"))
+
+frequency_long <- na.omit(frequency_long)
 
@@
 
+\newpage
+```{r}
+#| eval: true
+#| echo: false
+#| message: false
+#| warning: false
+
+## Code sourced from https://www.youtube.com/watch?v=ae_XVhjHd_o
+
+generate_word_graph <- function(combined_text,
+                                minimum_n = 100,
+                                minimum_correlation = 0.2) {
+
+combined_text_count <- combined_text |>
+  count(word, sort = TRUE) |>
+  filter(n >= 200)
+
+combined_correlations <- combined_text |>
+  semi_join(combined_text_count, by = "word") |>
+  pairwise_cor(item = word, feature = book) |>
+  filter(correlation >= 0.2)
+
+graph_from_data_frame(d = combined_correlations, #create graphs of networks of words
+                      vertices = combined_text_count |>
+                      semi_join(combined_correlations, by = c("word" = "item1"))) |>
+  ggraph(layout = "fr") +
+  geom_edge_link(aes(alpha = correlation), alpha = 0.05) +
+  geom_node_point() +
+  geom_node_text(aes(color = n, label = name), repel = TRUE)
+
+}
+
+# Code sourced from
+# filtered_text <- combined_text[combined_text$n >= 100, ] #only select words with a frequency larger or equal to 100
+
+# wordcloud(words = filtered_text$word, freq = filtered_text$n, scale=c(3,0.5), min.freq = 2, random.order = FALSE, colors= colorRampPalette(c("lightgray", "black"))(9))
+
+```
+
+```{r}
+#| eval: true
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-word-networks
+#| layout-ncol: 2
+#| fig-cap: "Word Networks Measured by Frequency and Correlation when Combining All Stream of Consciousness Novels"
+#| fig-subcap: ["100 Frequency, 0.2 Correlation", "200 Frequency, 0.4 Correlation", "200 Frequency, 0.8 Correlation", "400 Frequency, 1 Correlation"]
+
+combined_text|>
+  generate_word_graph(minimum_n = 100, minimum_correlation = 0.2)
+
+combined_text|>
+  generate_word_graph(minimum_n = 200, minimum_correlation = 0.4)
+
+combined_text|>
+  generate_word_graph(minimum_n = 200, minimum_correlation = 0.8)
+
+combined_text|>
+  generate_word_graph(minimum_n = 400, minimum_correlation = 1)
+```
+
 ## Sentiment Analysis
 
-Leveraging setiment analysis from @sentiment
+To further understand the emotional landscape and mental health themes related to SOC literature, I've combined all nine SOC texts to conduct an oveall sentiment analysis [@sentiment].
+
+For Bing index, @fig-sentiment-category shows that, SOC authors' literary works are predominantly negative.
+All authors, except for Mansfield, demonstrates this.
+68% of Joyce's works are negative compared to his 32% positive, similar to Eliot's, which is also the highest statistic out of all five authors.
+Woolf comes second with 60% words carrying negative meanings, while 40% are negative.
+Proust comes fourth with  57% negative and 43% positive.
+While Mansfield has words with positive connotations, at 51%, this isn't much different from her 49% negative words.
+
+The NRC index in @fig-sentiment-category further details the emotional nuances of SOC literature.
+The dominant emotional connotation in SOC literature's vocabulary is negative, standing at 16% to 22%.
+While the second highest percentage range belong to words with positive connotation, ranging from 11% to 17%, most of the other bars are occupied by words with negative-leaning emotions, such as sadness, fear, and anger.
+Sadness-related words, in particular, ranges the third highest, from 11% to 15%.
 
@@
 
-bing_lexicon <- get_sentiments("bing") %>%
+bing_lexicon <- get_sentiments("bing") |>
   rename(sentiment_binary = sentiment)
@@
   group_by(author) |>
-  mutate(percentage = n / sum(n) * 100)
+  mutate(percentage = round(n / sum(n) * 100))
 
@@
   geom_bar(stat = "identity") +
+    geom_text(aes(label = percentage),position = position_stack(vjust = 0.5), color = "white", size = 3) +
   labs(
-       x = "Author", y = "Percentage") +
+       x = "Author", y = "Percentage", fill = "Sentiment") +
   scale_fill_manual(values = c("positive" = "gray80", "negative" = "gray20", "neutral" = "gray50")) + 
   theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),  legend.key.size = unit(0.5, "cm"))
 
@@
   group_by(author) |>
-  mutate(percentage = n / sum(n) * 100)  |>
+  mutate(percentage = round(n / sum(n) * 100))  |>
   group_by(author) |>
@@
   geom_bar(stat = "identity") +
+   geom_text(aes(label = percentage),  position = position_stack(vjust = 0.5), color = "white", size = 3) +
   labs(
-       x = "Author", y = "Percentage") +
+       x = "Author", y = "Percentage", fill = "Sentiment") +
   scale_fill_manual(values = c("negative" = "gray1", "sadness" = "gray10", "anger" = "gray20", "disgust" = "gray30", "fear" = "gray40", "joy" = "gray50", "trust" = "gray60", "anticipation" = "gray70", "positive" = "gray80", "surprise" = "gray90")) +
   theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),
+       legend.key.size = unit(0.5, "cm"))
 
 ```
+
+The AFINN index in @fig-sentiment-number portrays a similar trend with its more quantitative sentiment analysis.
+On a scale of -5 to 5, SOC literature from these five authors mostly carry a negative sentiment, observable in the significant difference in height between the negative trendlines and the positive trendlines.
+Eliot's works are the most negative with 37% of words being -2 on the emotional scale between -5 and 5.
+Respectively, Woolf, Proust and Joyce tail beind at 33%, 32% and 31%, averaging out to be ~-2.2 on the emotional scale between -5 and 5.
+While most a large percentage Mansfield's works are positive, her highest percentage is negative on the AFINN index, at 24% for a scale of about -2.2 between -5 and 5.
+
+This further suggests that the SOC literary landscape is inundated with negative, disgust, fear and sadness-leaning vocabulary.
+These trends reflect the incoherent mental spaces, depressive perceptions, and existential angst inherent to SOC literature.
+These are the central themes often conveyed by SOC authors through their exploration of the unconscious mind and repressed emotions, reflecting the societal concerns of the late 19th to mid-20th century West [@pathology] [@resampling].
 
@@
   geom_line() +
+    geom_text(aes(label = percentage),position = position_stack(vjust = 0.5), color = "black", size = 2) +
   facet_wrap(~ author, scales = "free") +
@@
 
-## Combined Texts: Trends, Word Networks, Bigram and Trigram Analsis
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-
-## Code sourced from https://www.youtube.com/watch?v=ae_XVhjHd_o
-
-generate_word_graph <- function(combined_text,
-                                minimum_n = 100,
-                                minimum_correlation = 0.2) {
-
-combined_text_count <- combined_text |>
-  count(word, sort = TRUE) |>
-  filter(n >= 200)
-
-combined_correlations <- combined_text |>
-  semi_join(combined_text_count, by = "word") |>
-  pairwise_cor(item = word, feature = book) |>
-  filter(correlation >= 0.2)
-
-graph_from_data_frame(d = combined_correlations, #create graphs of networks of words
-                      vertices = combined_text_count |>
-                      semi_join(combined_correlations, by = c("word" = "item1"))) |>
-  ggraph(layout = "fr") +
-  geom_edge_link(aes(alpha = correlation), alpha = 0.05) +
-  geom_node_point() +
-  geom_node_text(aes(color = n, label = name), repel = TRUE)
-
-}
-
-# Code sourced from
-# filtered_text <- combined_text[combined_text$n >= 100, ] #only select words with a frequency larger or equal to 100
-
-# wordcloud(words = filtered_text$word, freq = filtered_text$n, scale=c(3,0.5), min.freq = 2, random.order = FALSE, colors= colorRampPalette(c("lightgray", "black"))(9))
-
-```
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-word-networks
-#| layout-ncol: 2
-#| fig-cap: "Word Networks Measured by Frequency and Correlation when Combining All Stream of Consciousness Novels"
-#| fig-subcap: ["100 Frequency, 0.2 Correlation", "200 Frequency, 0.4 Correlation", "200 Frequency, 0.8 Correlation", "400 Frequency, 1 Correlation"]
-
-combined_text|>
-  generate_word_graph(minimum_n = 100, minimum_correlation = 0.2)
-
-combined_text|>
-  generate_word_graph(minimum_n = 200, minimum_correlation = 0.4)
-
-combined_text|>
-  generate_word_graph(minimum_n = 200, minimum_correlation = 0.8)
-
-combined_text|>
-  generate_word_graph(minimum_n = 400, minimum_correlation = 1)
-```
 # Discussion
@@
 The novels' linguistic patterns reflect the socio-political landscape of the Western hemisphere, from  late 19th century to the mid-20th century.
+
+Stream of consciousness literature emerged as a significant literary technique in the early 20th century, particularly associated with the modernist movement in literature.
+Several factors contributed to its rise and prominence:
+
+Psychological Realism: One of the primary motivations for the development of stream of consciousness was the desire to represent the inner workings of the human mind more accurately.
+Writers sought to capture the complexity of thought processes, emotions, and perceptions in a more realistic and nuanced manner.
+Response to Realism and Naturalism: Stream of consciousness literature can be seen as a reaction against the more structured and external focus of realism and naturalism.
+Instead of portraying characters and events objectively, authors wanted to delve deeper into subjective experiences and individual consciousness.
+Technological and Cultural Changes: The early 20th century was a period of rapid technological advancement and social change, including the rise of psychology as a discipline.
+These developments influenced writers to explore new narrative techniques and modes of representation.
+Literary Experimentation: Modernist writers were often interested in pushing the boundaries of traditional narrative forms and experimenting with language, style, and structure.
+Stream of consciousness offered a way to challenge conventional storytelling techniques and engage readers in more innovative ways.
+Exploration of Identity and Perception: Stream of consciousness literature allowed writers to explore themes related to identity, consciousness, and perception.
+By immersing readers in the internal thoughts and sensations of characters, authors could interrogate questions about the nature of reality, self-awareness, and the construction of meaning.
+Overall, stream of consciousness literature became a prominent feature of modernist writing due to its capacity to convey the complexity of human experience and its ability to reflect the uncertainties and ambiguities of the modern world.
+
+The rise of modernism in literature, art, and culture during the late 19th and early 20th centuries was influenced by several interconnected factors:
+
+Industrialization and Urbanization: The rapid industrialization and urbanization of society brought about significant changes in the way people lived, worked, and interacted.
+This upheaval in traditional social structures and lifestyles led to a sense of dislocation and fragmentation, which modernist artists and writers sought to capture in their work.
+Technological Advances: Advances in technology, particularly in transportation, communication, and mass media, reshaped human experience and perception of time and space.
+Innovations such as the telegraph, telephone, and photography altered how people communicated, experienced distance, and understood reality.
+Crisis of Faith and Meaning: The late 19th and early 20th centuries witnessed a crisis of faith and meaning, as traditional religious and philosophical beliefs came into question.
+Developments such as Darwin's theory of evolution, Freudian psychology, and existential philosophy challenged established notions of human existence, morality, and purpose.
+World Wars and Political Turmoil: The turmoil of World War I and the interwar period brought about profound changes in society, politics, and culture.
+The devastation of war, the collapse of empires, and the emergence of new ideologies (such as communism and fascism) shattered old certainties and created an atmosphere of disillusionment and uncertainty.
+Intellectual and Cultural Exchange: The modernist movement was characterized by a spirit of intellectual and cultural exchange, with artists and writers drawing inspiration from diverse sources, including non-Western cultures, folk traditions, and avant-garde movements.
+This openness to new ideas and influences fostered experimentation and innovation in the arts.
+Quest for Individuality and Authenticity: Modernism rejected the conventions and constraints of traditional artistic forms and sought to express individual subjectivity and experience in new and unconventional ways.
+Artists and writers embraced themes of alienation, identity, and the search for authenticity in an increasingly fragmented and alienating world.
 
@@
 This can be extremely narrowing, and exclude nuances when analyzing the contexts behind childlessness.
-### Uneven Novel Length and Categorization of Authors
+
+### Reliance on Three Different Indices
 
@@
 Thus, this can affect the integrity of the datasets.
-
-### Project Gutenberg's Focus on the Canon
-
-Along the search for other stream of consciousness authors to include in the generated datasets, only a few-those that are in the SOC canon, mostly white authors-are present in Project Gutenberg, without including lesser known SOC authors, which could have steared the data analysis, still, towards their preconception: being rueful and angst-ridden
 

### 2024-04-22 paper/paper.qmd
results updated

@@
 
-Then, to quantify the number of times each word appears in each text, I used `count()` (@tbl-joyce-token).
-The result manifests into @fig-count.
-The figure demonstrates that beside the top frequencies of personal names, most of the words are abstract nouns that revolve around the nature of time, and life, e.g.
-god (appearing 194 times in Joyce's), time (appearing in 4 authors' texts, 121 times for Joyce's, 77 times for Woolf's, 421 times for Proust's, 23 times for Eliot), mind (108 times in Joyce's and 196 times in Proust's), life (appearing 266 times in Proust's and 131 times in Joyce's), etc.
-This further exemplifies the sense of disassociation, temporality and non-linearity commonly associated to stream-of-consciousness literature and more broadly with mental health themes [@flow] [@pathology] [@resampling].
-
-\newpage
+Then, to quantify the number of times each word appears in each text, I used `count()` (@tbl-joyce-token). The result manifests into @fig-count.
 
@@
 
-frequency |>
-  rename("Word" = word) |>
-  head(5) |>
-  kable(
-    col.names = c("Word", "James Joyce", "Katherine Mansfield", "Marcel Proust", "T.S. Eliot", "Virignia Woolf"),
-    booktabs = TRUE
-  ) |>
-    kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
-
-```
-
-To accurately compare word frequency among SOC authors, from @tbl-frequency, I convert each word frequency into a percentage.
-This is done by using `mutate` to calculate the word frequency of each word relative to the total word frequency of each author.
-Due to the different number of words present in each author's literary works, this method accounts for the inconsistencies of between each author's word count.
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-boxplot 
-#| tbl-cap: "Overall Boxplot Summary Statistics of all Word Frequency Percentages (AFINN Sentiment)"
-#| warning: false
-
 #reshape the frequency datasets and add sentiment values
@@
 
-frequency_long <- na.omit(frequency_long)
+frequency_long |> 
+  rename("Word" = word) |>
+  head(5) |>
+  kable(
+    col.names = c("Word", "Author", "Frequency", "Sentiment Binary", "Value", "Sentiment"),
+    booktabs = TRUE
+  ) |>  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
+```
+
+To accurately compare word frequency among SOC authors, from @tbl-frequency, I convert each word frequency from each author into a percentage.
+This is done by using `mutate` to calculate the word frequency of each word relative to the total word frequency of each author.
+Due to the different number of words present in each author's literary works, this method accounts for the inconsistencies of between each author's word count.
+
+\newpage
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-boxplot 
+#| tbl-cap: "Overall Boxplot Summary Statistics of all Word Frequency Percentages (AFINN Sentiment)"
+#| warning: false
 
@@
   ) |>
-    head(10) |>
+    head(8) |>
   kable(
@@
 
-@fig-sentiment-frequency and @tbl-boxplot captures the data garnered.
-For the NRC index, for all authors, the lower word frequency percentage ranges from 0.0000819 to 0004255 and are largely words that carry a mixture of emotional connotations such as, joy, surprise, positive, disgust and anticipation.
-However, the higher word frequency percentage, ranging from 0,0004255 to 0.0008511, mostly contains words with negative connotations like fear and anger, that weighs a -1.875 median on a scale between -5 to 5 in the AFINN index.
-Similarly, for the Bing index, the higher word frequency ranging from 0,0004255 to 0.0008511 percentages largely contains words that carry negative connotations, that weighs approximately a -1.875 median on a scale between -5 to 5 in the AFINN index.
-
-Overseeing the entire dataset, @tbl-boxplot demonstrates that, for all distinct word frequency percentages, the median sentiment value is -2 on the scale between -5 and 5.
+@fig-sentiment-frequency and @tbl-boxplot captures the text data garnered and their emotional connotations.
+Overseeing the entire dataset, @tbl-boxplot in particular demonstrates that, for all distinct word frequency percentages, the median sentiment value is -2 on the scale between -5 and 5.
 More specifically, words that appear most frequently, or words with the frequency percentage of 0.0008511, have the lowest mean sentiment value-or the most negative connotation of-2.33 on the scale between -5 and 5 in the AFINN sentiment index.
-This summary statistics of the gathered data only further shows how the SOC literary landscape mostly associates with words that carry more negative or even extreme connotations often associated with mental health themes, and topics such as depression, disassociation or schizophrenia [@pathology].
+This summary statistics of the gathered data only further shows how the SOC literary landscape mostly associates with words that carry more negative or even extreme connotations often associated with mental health themes, and topics such as depression, disassociation or schizophrenia [@pathology], which will be further discussed in [Results].
 
@@
        y = "Sentiment Value (AFINN)",
-       fill = "Sentiment Associated") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+       fill = "Sentiment") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_fill_manual(values = c("lightblue", "darkblue"))
 
@@
        y = "Sentiment Value (AFINN)",
-       fill = "Sentiment Associated") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+       fill = "Sentiment") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_fill_manual(values = c(
+    "negative" = "darkblue", "sadness" = "steelblue4", "anger" = "royalblue4",
+    "disgust" = "cadetblue", "fear" = "blue4", "joy" = "skyblue",
+    "trust" = "lightblue", "anticipation" = "deepskyblue", "positive" = "dodgerblue",
+    "surprise" = "lightsteelblue"
+  ))
 
@@
 With the large amount of different literary texts need processing by myself, it's hard to truly conduct thorough word processing.
-Certain textual elements, while can have their frequencies accounted for, can't have sentiment analysis due to their factual nature.
+Certain texts, while can have their frequencies accounted for, can't have sentiment analysis due to their factual nature.
 These textual elements might include things like chapter titles, Roman numbers and personal names, which can strongly skew and affect the data integrity.
@@
 
-## The Dominant Vocabulary of SOC Literature
+## The Dominant Vocabulary of SOC Literature, A Word Frequency Analysis
+
+@fig-count demonstrates that beside the top frequencies of personal names, most of the words are abstract nouns that revolve around the nature of time, and life.
+For instance, these words are, god (appearing 194 times in Joyce's), time (appearing in 4 authors' texts, 121 times for Joyce's, 77 times for Woolf's, 421 times for Proust's, 23 times for Eliot), mind (108 times in Joyce's and 196 times in Proust's), life (appearing 266 times in Proust's and 131 times in Joyce's), etc.
+This further exemplifies the sense of disassociation, temporality and non-linearity commonly associated to stream-of-consciousness literature and more broadly with mental health themes [@flow] [@pathology] [@resampling].
+
+Within @fig-sentiment-frequency, for the NRC index, for all authors, the lower word frequency percentage ranges from 0.0000819 to 0004255 and are largely words that carry a mixture of emotional connotations such as, joy, surprise, positive, disgust and anticipation.
+However, the higher word frequency percentage, ranging from 0,0004255 to 0.0008511, mostly contains words with negative connotations like fear and anger, that weighs a -1.875 median on a scale between -5 to 5 in the AFINN index.
+Similarly, for the Bing index, the higher word frequency ranging from 0,0004255 to 0.0008511 percentages largely contains words that carry negative connotations, that weighs approximately a -1.875 median on a scale between -5 to 5 in the AFINN index.
 
@@
 combined_text|>
-  generate_word_graph(minimum_n = 100, minimum_correlation = 0.2)
+  generate_word_graph(minimum_n = 50, minimum_correlation = 0.2)
 
@@
 combined_text|>
-  generate_word_graph(minimum_n = 200, minimum_correlation = 0.8)
+  generate_word_graph(minimum_n = 600, minimum_correlation = 0.8)
 
 combined_text|>
-  generate_word_graph(minimum_n = 400, minimum_correlation = 1)
+  generate_word_graph(minimum_n = 1000, minimum_correlation = 1)
 ```
@@
 
-To further understand the emotional landscape and mental health themes related to SOC literature, I've combined all nine SOC texts to conduct an oveall sentiment analysis [@sentiment].
+To further understand the emotional landscape and mental health themes related to SOC literature, I've combined all nine SOC texts to conduct an overall sentiment analysis [@sentiment].
 
@@
 
+\newpage
 ```{r}
@@
   scale_color_gradient(limits = c(0, 0.001), 
-                       low = "blue2", high = "gray50") +
+                       low = "lightblue", high = "darkblue") +
   facet_wrap(~author, ncol = 1) +
@@
   scale_color_gradient(limits = c(0, 0.001), 
-                       low = "blue2", high = "gray100") +
+                       low = "lightblue", high = "darkblue") +
   facet_wrap(~author, ncol = 2) +
@@
   scale_color_gradient(limits = c(0, 0.001), 
-                       low = "blue2", high = "gray100") +
+                       low = "lightblue", high = "darkblue") +
   facet_wrap(~author, ncol = 1) +

### 2024-04-22 paper/paper.qmd
results updated

@@
 
-- 'AFINN' from Finn Årup Nielsen, which assigns a numerical value from '-5 to 5' 
-- 'bing' from Bing Liu and collaborators, which assigns if a word is 'positive' or negative'
+- 'AFINN' from Finn Årup Nielsen, which assigns each word a numerical value from '-5 to 5', judging its emotionality from negative to postive
+- 'bing' from Bing Liu and collaborators, which assigns if a word is either 'positive' or negative'
 - 'nrc' from Saif Mohammad and Peter Turney, which assigns a core emotional value to a word, such as 'fear', 'anger', 'sadness' or 'trust'.
@@
 
-To garner the word count of SOC literature  using the library `gutenbergr`, I first downloaded the index of each respective SOC author's body of works using each author's name as outlined in Project Gutenberg database, namely Joyce, Mansfield, Eliot, Woolf, and Proust [@rGuten] ([Appendix]).
-Then, to create my own datasets containing the word frequency within each SOC author's texts, I used the identification number of the desired literary work and the function `gutenberg_download` to upload the HTML texts into individual csv-and later on, parquet-datasets, as observed in @tbl-joyce for James Joyce.
+To collect text data from five SOC authors using the `gutenbergr` package, I began by downloading the index of each author's works from the Project Gutenberg database.
+These authors include Joyce, Mansfield, Eliot, Woolf, and Proust [@rGuten] ([Appendix]).
+
+Next, I created custom datasets containing each author's raw texts.
+This involved using the identification number of the desired literary work and the `gutenberg_download` function to retrieve the HTML texts.
+Subsequently, I saved the data into individual CSV files and later converted them into Parquet datasets, as demonstrated in @tbl-joyce example for James Joyce.
 
@@
 
-Next, observable in @tbl-joyce-token, to measure the word frequency of each literary text, I first tokenized or separated the texts into individual chunks of words using `unnest_tokens()` [@rTidytext].In addition, to account for stop words like, a, of, and, etc., I leveraged the dataset `stop_words`and the function `anti_join` from @rTidytext to exclude all stop words from my self-generated datasets of literary texts, ensuring data validity and meaningful analysis.
+Next, observable in @tbl-joyce-token, to measure the word count of each literary text, I first tokenized or separated the texts into individual chunks of words using `unnest_tokens()` [@rTidytext].In addition, to account for stop words like, a, of, and, etc., I leveraged the dataset `stop_words`and the function `anti_join` from @rTidytext to exclude all stop words from my self-generated datasets of literary texts, ensuring data validity and meaningful analysis.
 
@@
 ggplot(aes(x = n, y = reorder(word, n))) +
-  geom_bar(stat = "identity", fill = "darkgrey") +
+  geom_bar(stat = "identity", fill = "steelblue4") +
   geom_text(aes(label = n), hjust = -0.3, vjust = 0.5, color = "grey40", size = 3) +
@@
   kable(
-    col.names = c("Word", "Author", "Frequency", "Sentiment Binary", "Value", "Sentiment"),
+    col.names = c("Word", "Author", "Frequency Percent", "Binary Sentiment", "Value", "Sentiment"),
     booktabs = TRUE
@@
 
-To accurately compare word frequency among SOC authors, from @tbl-frequency, I convert each word frequency from each author into a percentage.
-This is done by using `mutate` to calculate the word frequency of each word relative to the total word frequency of each author.
+To accurately compare word count among SOC authors, from @tbl-frequency, I convert each word count from each author into a percentage.
+This is done by using `mutate` to calculate the word frequency of each word relative to the total word count of each author.
 Due to the different number of words present in each author's literary works, this method accounts for the inconsistencies of between each author's word count.
+As explained, the variables within @tbl-frequency are:
+
+- **Word:** This variable is the word sourced from SOC literary texts.
+- **Author:** This variable accounts for the author whose books contain the word in question.
+- **Frequency:** This variable is the word count percentage of the word in question against each author's total word count.
+- **Binary Sentiment (Bing):** This variable, as explained, comes from the Bing sentiment index, enabling us to analyze how every word is usually perceived emotionally, whether negatively or positively.
+- **Value of Sentiment (AFINN):** This variable, as explained, comes from the AFINN sentiment index, enabling us to analyze how every word is usually perceived emotionally, on a scale between -5 and 5.
+- **Sentiment (NRC):** This variable, as explained, comes from the NRC sentiment index, enabling us to analyze how every word is usually perceived emotionally, whether it's associated with anger, anticipation, disgust, fear, joy, negative, positive, sadness, surprise or trust.
 
@@
 
-@fig-sentiment-frequency and @tbl-boxplot captures the text data garnered and their emotional connotations.
-Overseeing the entire dataset, @tbl-boxplot in particular demonstrates that, for all distinct word frequency percentages, the median sentiment value is -2 on the scale between -5 and 5.
-More specifically, words that appear most frequently, or words with the frequency percentage of 0.0008511, have the lowest mean sentiment value-or the most negative connotation of-2.33 on the scale between -5 and 5 in the AFINN sentiment index.
-This summary statistics of the gathered data only further shows how the SOC literary landscape mostly associates with words that carry more negative or even extreme connotations often associated with mental health themes, and topics such as depression, disassociation or schizophrenia [@pathology], which will be further discussed in [Results].
+@tbl-boxplot captures the text data garnered and their emotional connotations.
+Overseeing the entire dataset, @tbl-boxplot in particular demonstrates that, for all distinct word frequency, the median sentiment value is -2 on the scale between -5 and 5.
+More specifically, words that appear most frequently, or words with the frequency  of 0.0008511%, have the lowest mean sentiment value-or the most negative connotation of -2.33 on the scale between -5 and 5 in the AFINN sentiment index.
+This summary statistics further shows how SOC literature mostly associates with words that carry more negative or even extreme connotations often associated with mental health themes, and topics such as depression, disassociation or schizophrenia [@pathology], a central argument that I'll further discuss in [Results].
+
+## Data Limitations
+
+### Lack of Thorough Word Cleaning
+
+With the large amount of different literary texts need processing by myself, it's hard to truly conduct thorough word processing.
+Certain texts, while can have their frequencies accounted for, can't have sentiment analysis due to their factual nature.
+These textual elements might include things like chapter titles, Roman numbers and personal names, which can strongly skew and affect the data integrity.
+As observed in @fig-count, "Swann", "Stephen" or "Jacob" are the most repeated words in Proust's, Joyce's and Woolf's, proving this data limitation.
+
+### Potential Measurement Errors
+
+The actual raw text data sourced from Project Gutenberg may be susceptible to measurement errors due to various factors, such as selection bias towards older, Western and canonical texts, social desirability bias that edits out certain political elements, memory lapses, or misinterpretations of literary texts by voluntary proofreaders.
+While the archive has strict guidelines to minimize these errors, they cannot be entirely eliminated.
+
+### Uneven Text Length
+
+I chose literary texts based on their global reception and canonical status within the stream-of-consciousness genre.
+This decision involved forgoing certain practical considerations, such as ensuring similar text lengths across all works by each author.
+While this approach enables a thorough exploration of word frequency analysis, sentiment value, and word networks within the SOC genre, it also has the potential to emphasize the frequencies of certain words from specific authors more than others due to the longer text length of those authors.
+This could critically affect the integrity of the data analysis.
+
+### Biased Author Selection
+
+Furthermore, SOC literature is expansive, spanning a wide range of themes, styles, and narrative techniques.
+By exclusively representing the genre through a select group of canonical Western authors, I risk constraining the genre's diversity.
+By oversimplifying the complexity of the the dynamic literary genre, I also affect the integrity of the datasets and its correlation to mental health themes.
+
+# Results
+
+## The Dominant Vocabulary of SOC Literature, A Word Frequency Analysis
+
+@fig-count demonstrates that beside personal names, words with the highest counts are mostly abstract nouns that revolve around the nature of time, and life.
+For instance, these words are, \textit{god} (appearing 194 times in Joyce's), \textit{time} (appearing in 4 authors' texts, 121 times for Joyce's, 77 times for Woolf's, 421 times for Proust's, 23 times for Eliot), \textit{mind} (108 times in Joyce's and 196 times in Proust's), \textit{life} (appearing 266 times in Proust's and 131 times in Joyce's), etc.
+This further exemplifies the sense of disassociation, temporality and non-linearity commonly associated to stream-of-consciousness literature and more broadly with mental health themes [@flow] [@pathology] [@resampling].
 
@@
 
-## Data Limitations
+Within @fig-sentiment-frequency, for the NRC index, between all authors, the words with lower frequency-those ranging from 0.0000819% to 0004255%-are largely words that carry a mixture of emotional connotations such as, joy, surprise, positive, disgust and anticipation.
+However, words with high frequency-ranging from 0,0004255% to 0.0008511%-mostly contain negative connotations like fear and negative They also have a -1.875 median on a scale between -5 to 5 in the AFINN index.
+Similarly, for the Bing index, words with higher frequency-those ranging from 0,0004255% to 0.0008511%-largely carry negative connotations.
+They also have, approximately, a -1.875 median on a scale between -5 to 5 in the AFINN index.
+This further emphasizes how SOC literature mostly contains negativity-inclined vocabulary and linguistic patterns
 
-### Lack of Thorough Word Cleaning
+To further explore the contextual relationship between each word, @fig-word-networks creates word networks of all nine of these SOC texts.
+These networks will capture how often each word frequents in these SOC texts and how closely each correlates with others, whether as single words (unigrams), pairs of words (bigrams), or triplets of words (trigrams).
+In particular, the darker the connecting lines, the more correlative these combinations of words are, and the bluer the words, the more frequent they appear.
 
-With the large amount of different literary texts need processing by myself, it's hard to truly conduct thorough word processing.
-Certain texts, while can have their frequencies accounted for, can't have sentiment analysis due to their factual nature.
-These textual elements might include things like chapter titles, Roman numbers and personal names, which can strongly skew and affect the data integrity.
-As observed in @fig-count, "Swann", "Stephen" or "Jacob" are the most repeated words in Proust's, Joyce's and Woolf's, proving this data limitation.
-
-### Potential Measurement Errors
-
-The actual raw text data sourced from Project Gutenberg may be susceptible to measurement errors due to various factors, such as selection bias towards older, Western and canonical texts, social desirability bias that edits out certain political elements, memory lapses, or misinterpretations of literary texts by voluntary proofreaders.
-While the archive has strict guidelines to minimize these errors, they cannot be entirely eliminated.
-
-### Uneven Text Length
-
-I chose literary texts based on their global reception and canonical status within the stream-of-consciousness genre.
-This decision involved forgoing certain practical considerations, such as ensuring similar text lengths across all works by each author.
-While this approach enables a thorough exploration of word frequency analysis, sentiment value, and word networks within the SOC genre, it also has the potential to emphasize the frequencies of certain words from specific authors more than others due to the longer text length of those authors.
-This could critically affect the integrity of the data analysis.
-
-## Biased Author Selection
-
-Furthermore, SOC literature is expansive, spanning a wide range of themes, styles, and narrative techniques.
-By exclusively representing the genre through a select group of canonical Western authors, I risk constraining the genre's diversity.
-By oversimplifying the complexity of the the dynamic literary genre, I also affect the integrity of the datasets and its correlation to mental health themes.
-
-# Results
-
-## The Dominant Vocabulary of SOC Literature, A Word Frequency Analysis
-
-@fig-count demonstrates that beside the top frequencies of personal names, most of the words are abstract nouns that revolve around the nature of time, and life.
-For instance, these words are, god (appearing 194 times in Joyce's), time (appearing in 4 authors' texts, 121 times for Joyce's, 77 times for Woolf's, 421 times for Proust's, 23 times for Eliot), mind (108 times in Joyce's and 196 times in Proust's), life (appearing 266 times in Proust's and 131 times in Joyce's), etc.
-This further exemplifies the sense of disassociation, temporality and non-linearity commonly associated to stream-of-consciousness literature and more broadly with mental health themes [@flow] [@pathology] [@resampling].
-
-Within @fig-sentiment-frequency, for the NRC index, for all authors, the lower word frequency percentage ranges from 0.0000819 to 0004255 and are largely words that carry a mixture of emotional connotations such as, joy, surprise, positive, disgust and anticipation.
-However, the higher word frequency percentage, ranging from 0,0004255 to 0.0008511, mostly contains words with negative connotations like fear and anger, that weighs a -1.875 median on a scale between -5 to 5 in the AFINN index.
-Similarly, for the Bing index, the higher word frequency ranging from 0,0004255 to 0.0008511 percentages largely contains words that carry negative connotations, that weighs approximately a -1.875 median on a scale between -5 to 5 in the AFINN index.
+ providing us with a thematic understanding of SOC literature.
 
@@
 
+##first, n >=250
 generate_word_graph <- function(combined_text,
@@
   count(word, sort = TRUE) |>
-  filter(n >= 200)
+  filter(n >= 150)
 
@@
   pairwise_cor(item = word, feature = book) |>
-  filter(correlation >= 0.2)
+  filter(correlation >= 1)
 
@@
   geom_node_text(aes(color = n, label = name), repel = TRUE)
+}
 
+##second, n >=300
+generate_word_graph_second <- function(combined_text,
+                                minimum_n = 100,
+                                minimum_correlation = 0.2) {
+
+combined_text_count <- combined_text |>
+  count(word, sort = TRUE) |>
+  filter(n >= 250)
+
+combined_correlations <- combined_text |>
+  semi_join(combined_text_count, by = "word") |>
+  pairwise_cor(item = word, feature = book) |>
+  filter(correlation >= 1)
+
+graph_from_data_frame(d = combined_correlations, #create graphs of networks of words
+                      vertices = combined_text_count |>
+                      semi_join(combined_correlations, by = c("word" = "item1"))) |>
+  ggraph(layout = "fr") +
+  geom_edge_link(aes(alpha = correlation), alpha = 0.05) +
+  geom_node_point() +
+  geom_node_text(aes(color = n, label = name), repel = TRUE)
+}
+
+##third, n >=350
+generate_word_graph_third <- function(combined_text,
+                                minimum_n = 100,
+                                minimum_correlation = 0.2) {
+
+combined_text_count <- combined_text |>
+  count(word, sort = TRUE) |>
+  filter(n >= 350)
+
+combined_correlations <- combined_text |>
+  semi_join(combined_text_count, by = "word") |>
+  pairwise_cor(item = word, feature = book) |>
+  filter(correlation >= 0.6)
+
+graph_from_data_frame(d = combined_correlations, #create graphs of networks of words
+                      vertices = combined_text_count |>
+                      semi_join(combined_correlations, by = c("word" = "item1"))) |>
+  ggraph(layout = "fr") +
+  geom_edge_link(aes(alpha = correlation), alpha = 0.05) +
+  geom_node_point() +
+  geom_node_text(aes(color = n, label = name), repel = TRUE)
+}
+
+## fourth, n>= 400
+generate_word_graph_final <- function(combined_text,
+                                minimum_n = 100,
+                                minimum_correlation = 0.2) {
+
+combined_text_count <- combined_text |>
+  count(word, sort = TRUE) |>
+  filter(n >= 400)
+
+combined_correlations <- combined_text |>
+  semi_join(combined_text_count, by = "word") |>
+  pairwise_cor(item = word, feature = book) |>
+  filter(correlation >= 0.8)
+
+graph_from_data_frame(d = combined_correlations, #create graphs of networks of words
+                      vertices = combined_text_count |>
+                      semi_join(combined_correlations, by = c("word" = "item1"))) |>
+  ggraph(layout = "fr") +
+  geom_edge_link(aes(alpha = correlation), alpha = 0.05) +
+  geom_node_point() +
+  geom_node_text(aes(color = n, label = name), repel = TRUE)
 }
@@
 combined_text|>
-  generate_word_graph(minimum_n = 50, minimum_correlation = 0.2)
+  generate_word_graph(minimum_n = 250, minimum_correlation = 0.2)
 
 combined_text|>
-  generate_word_graph(minimum_n = 200, minimum_correlation = 0.4)
+  generate_word_graph_second(minimum_n = 300, minimum_correlation = 0.4)
 
 combined_text|>
-  generate_word_graph(minimum_n = 600, minimum_correlation = 0.8)
+  generate_word_graph_third(minimum_n = 350, minimum_correlation = 0.6)
 
 combined_text|>
-  generate_word_graph(minimum_n = 1000, minimum_correlation = 1)
+  generate_word_graph_final(minimum_n = 400, minimum_correlation = 0.8)
 ```
@@
        x = "Author", y = "Percentage", fill = "Sentiment") +
-  scale_fill_manual(values = c("positive" = "gray80", "negative" = "gray20", "neutral" = "gray50")) + 
+  scale_fill_manual(values = c("positive" = "lightblue", "negative" = "darkblue")) + 
   theme_minimal() +
@@
        x = "Author", y = "Percentage", fill = "Sentiment") +
-  scale_fill_manual(values = c("negative" = "gray1", "sadness" = "gray10", "anger" = "gray20", "disgust" = "gray30", "fear" = "gray40", "joy" = "gray50", "trust" = "gray60", "anticipation" = "gray70", "positive" = "gray80", "surprise" = "gray90")) +
+   scale_fill_manual(values = c(
+    "negative" = "darkblue", "sadness" = "steelblue4", "anger" = "royalblue4",
+    "disgust" = "cadetblue", "fear" = "blue4", "joy" = "skyblue",
+    "trust" = "lightblue", "anticipation" = "deepskyblue", "positive" = "dodgerblue",
+    "surprise" = "lightsteelblue"
+  )) + 
   theme_minimal() +
@@
        x = "Sentiment Value", y = "Percentage") +
-  scale_color_manual(values = c("positive" = "black", "negative" = "gray")) + # Adjust line colors
+  scale_color_manual(values = c("positive" = "lightblue", "negative" = "darkblue")) + # Adjust line colors
   theme_minimal()

### 2024-04-23 paper/paper.qmd
results and abstract updates

@@
 abstract: "This project focuses on understanding the language used by renowned stream of consciousness (SOC) authors James Joyce, Virginia Woolf, Marcel Proust, Katherine Mansfield and T.S Eliot.
-By conducting word frequency analysis and sentiment analysis on these authors' nine novels, this paper aims to uncover shared linguistic patterns and gain insights into the authors' mental states.
-This paper attempts to offer insights into themes of self-identity, anxiety, disassociation and existential contemplation within Western society and its literary circle from late 19th to mid-20th century.
-(add one sentence on main results)"
+By conducting word frequency analysis and sentiment analysis on these authors' nine novels, we aim to uncover shared linguistic patterns and gain insights into the authors' mental states.
+Our analysis of the SOC genre shows that higher frequency words are mostly negative and carry sadness or fear-inclined connotations.
+With these trends uncovered, we attempt to offer insights into themes of self-identity, anxiety, disassociation and existential contemplation within Western society and its literary circle from late 19th to mid-20th century."
 format: pdf
@@
 
-By analyzing these textual datasets through word frequency and sentiment analysis, I seek to pose and answer crucial questions: \textit{What are some important factors contributing to this relationship between mental health, disassociation and stream of consciousness?
+By analyzing these textual datasets through word frequency and sentiment analysis, this paper seeks to pose and answer crucial questions: \textit{What are some important factors contributing to this relationship between mental health, disassociation and stream of consciousness?
 Moreover, how does this relationship vary differently across differnet demographics of authors, for instance, authors with different geographical locations and genders?} Understanding these dynamics is crucial in having an informed understanding of the West's late 19th to mid-20th century literature and even socio-political landscape, especially in regards to how authors and creative writers navigate and deal with then-taboo topics such as existential angst, mental health issues and disabilities.
@@
 Through our analysis, we found that (percentage, number and data here, main results)...
+Analysis of the SOC genre shows that higher frequency words are mostly negative (-1.875 median on AFINN index); sentiment-wise, analysis reveals 11% to 15% of words in SOC texts are sadness-inclined (NRC index), and over 60% of words in selected SOC literature are negative, with Eliot's works being the most negative at 68% (Bing index).
+Temporal nouns like "time," "day," "life," "moment," and "night" are most frequent, correlating with SOC's non-linear nature,
+
+Analysis of SOC literature reveals lower frequency words have mixed emotional connotations, while higher frequency words tend to be negative (-1.875 median on AFINN index).
+Temporal nouns like "time," "day," "life," "moment," and "night" are most frequent.
+Sentiment analysis shows SOC literature is predominantly negative, with Eliot's works being the most negative (37% on AFINN index), followed closely by Joyce, Woolf, Proust, and Mansfield.
+The NRC index indicates a prevalence of negative emotions, particularly sadness-related words.
 
 To further understand the correlation between stream of consciousness novels and mental health themes, in [Introduction], the paper briefly discusses the nature of stream of consciousness literature, relevant authors and the works that I've chosen to analyze.
-Subsequently, in [Data] and [Results], I talk about the nature of the data obtained and analyze the results garnered from the data with suitable tables and charts.
+Subsequently, in [Data] and [Results], we talk about the nature of the data obtained and analyze the results garnered from the data with suitable tables and charts.
 Next, [Discussion] provides further insights and future areas of study.
 Finally, [Conclusion] summarizes our main findings.
-To complete the paper, [Appendix] clarifies how each variable within each dataset is generated with relevant tables to accordingly demonstrate this.
-
-The novel texts used for analysis were sourced from Project Gutenberg under the library `gutenbergr` [@rGuten] [@gutenberg].
-Data was generated, extracted and cleaned using the open-source statistical programming language R [@r], leveraging functions from `tidyverse` [@rTidyverse], `tidytext` [@rTidytext], `rmarkdown` [@rMark], `dplyr` [@rDplyr], `ggplot2` [@rGgplot2], `scales` [@rScales], `here` [@rHere], `igraph` [@rIgraph], `widyr` [@rWidy], `ggraph` [@rGg], `textdata` [@rTextdata], `tm` [@rTm], `here` [@here], `kableExtra` [@kableExtra], `arrow` [@arrow], and `knitr` [@rKnitr].
 
 # Data {#sec-data}
+
+The SOC texts used for analysis were sourced from Project Gutenberg under the library `gutenbergr` [@rGuten] [@gutenberg].
+Data was generated, extracted and cleaned using the open-source statistical programming language R [@r], leveraging functions from `tidyverse` [@rTidyverse], `tidytext` [@rTidytext], `rmarkdown` [@rMark], `dplyr` [@rDplyr], `ggplot2` [@rGgplot2], `scales` [@rScales], `here` [@rHere], `igraph` [@rIgraph], `widyr` [@rWidy], `ggraph` [@rGg], `textdata` [@rTextdata], `tm` [@rTm], `here` [@here], `kableExtra` [@kableExtra], `arrow` [@arrow], and `knitr` [@rKnitr].
 
@@
 
-* **Word Frequency:** This variable captures the repetition of a single word (unigram), two-words combination (bigram) or three-words combination (trigram) throughout a SOC novel text, providing us with a thematic understanding of SOC literature.
+* **Word Frequency:** This variable captures the repetition of a single word (unigram) or a pair of words combination (bigram) throughout a SOC novel text, providing us with a thematic understanding of SOC literature.
 * **Sentiment Value:** This variable enables us to analyze how every word is usually perceived emotionally, whether it be a qualitative feeling or if it is conveyed through a numerical value.
 
-Out of two variables used, the first one, 'Word Frequency' usually captured as `n` or `frequency` in datasets, is directly quantified through tokenizing the novel texts using the package `tidytext` and its function `unnest_tokens()` [@rTidytext].
-To do this, I first downloaded all nine novel texts from @gutenberg, and leveraged functions such as `unnest_tokens()` from @rTidytext to mine the texts, or separating it into individual words.
-Finally, I used `count()` to quantify the word frequency.
+Out of two variables used, the first one, 'Word Frequency' usually captured as `n/sum(n)` or `frequency` in datasets, is directly quantified through tokenizing the novel texts using the package `tidytext` and its function `unnest_tokens()` [@rTidytext].
+To do this, we downloaded all nine SOC texts from @gutenberg, using functions such as `unnest_tokens()` from @rTidytext to mine the texts, or separating them into individual words.
+Finally, I used `count()` to quantify the word count (`n`), and later divide the each word's count over the total word count of each author to garner the word frequency percentages.
 
 The second variable used, 'Sentiment Value', is based on three English-based, general-purpose  "word-emotion and word-polarity association lexicons", sourced from Mohammad and Turney's expansive research along with efforts from Finn Årup Nielsen and Bing Liu and collaborators [@rTidytext] [@sentiment].
-The three general-purpose lexicon that contributes to my sentiment analysis are [@rTidytext]:
+The three general-purpose lexicon that contributes to this paper's sentiment analysis are [@rTidytext]:
 
@@
 - 'bing' from Bing Liu and collaborators, which assigns if a word is either 'positive' or negative'
-- 'nrc' from Saif Mohammad and Peter Turney, which assigns a core emotional value to a word, such as 'fear', 'anger', 'sadness' or 'trust'.
+- 'nrc' from Saif Mohammad and Peter Turney, which assigns a core emotional value to a word, such as 'fear', 'anger', 'sadness' or 'trust'
 
@@
 
-However, I do recognize how reductive this quantification of language can be.
+However, this paper does recognize how reductive this quantification of language can be.
 When it comes to understand such social and human artifacts as language or literary texts, much is dependent on their context.
-As such, I will further discuss these weaknesses of the datasets under [Discussion].
+As such, this paper will further discuss these weaknesses of the datasets under [Discussion].
 
@@
 
-With these points being addressed and the impartial nature of the archive being publicly disclosed on its website, the source data in which I garner my literary texts are selected with strict quality and accuracy.
+With these points being addressed and the impartial nature of the archive being publicly disclosed on its website, the source data seems to archive SOC literary texts with strict quality control and accuracy.
 However, there are potential spaces for biases such as selection bias where Project Gutenberg primarily hosts older works that have entered the public domain.
@@
 
-To collect text data from five SOC authors using the `gutenbergr` package, I began by downloading the index of each author's works from the Project Gutenberg database.
+To collect text data from five SOC authors using the `gutenbergr` package, we first downloaded the index of each author's works from the Project Gutenberg database.
 These authors include Joyce, Mansfield, Eliot, Woolf, and Proust [@rGuten] ([Appendix]).
 
-Next, I created custom datasets containing each author's raw texts.
+Next, custom datasets containing each author's raw texts were created.
 This involved using the identification number of the desired literary work and the `gutenberg_download` function to retrieve the HTML texts.
-Subsequently, I saved the data into individual CSV files and later converted them into Parquet datasets, as demonstrated in @tbl-joyce example for James Joyce.
+Subsequently, the data was saved into individual CSV files and later converted into Parquet datasets, as demonstrated in @tbl-joyce example for James Joyce.
 
@@
 
-Next, observable in @tbl-joyce-token, to measure the word count of each literary text, I first tokenized or separated the texts into individual chunks of words using `unnest_tokens()` [@rTidytext].In addition, to account for stop words like, a, of, and, etc., I leveraged the dataset `stop_words`and the function `anti_join` from @rTidytext to exclude all stop words from my self-generated datasets of literary texts, ensuring data validity and meaningful analysis.
+Next, observable in @tbl-joyce-token, to measure the word count of each literary text, we first tokenized or separated into individual chunks of words using `unnest_tokens()` [@rTidytext].In addition, to account for stop words like, a, of, and, etc., the dataset `stop_words`and the function `anti_join` from @rTidytext were used to exclude all stop words from the self-generated datasets of literary texts, ensuring data validity and meaningful analysis.
 
-Then, to quantify the number of times each word appears in each text, I used `count()` (@tbl-joyce-token). The result manifests into @fig-count.
+Then, to quantify the number of times each word appears in each text, we used `count()` (@tbl-joyce-token). The result manifests into @fig-count.
 
@@
 
-To accurately compare word count among SOC authors, from @tbl-frequency, I convert each word count from each author into a percentage.
+To accurately compare word count among SOC authors, from @tbl-frequency, qw converted each word count from each author into a percentage.
 This is done by using `mutate` to calculate the word frequency of each word relative to the total word count of each author.
@@
 Overseeing the entire dataset, @tbl-boxplot in particular demonstrates that, for all distinct word frequency, the median sentiment value is -2 on the scale between -5 and 5.
-More specifically, words that appear most frequently, or words with the frequency  of 0.0008511%, have the lowest mean sentiment value-or the most negative connotation of -2.33 on the scale between -5 and 5 in the AFINN sentiment index.
-This summary statistics further shows how SOC literature mostly associates with words that carry more negative or even extreme connotations often associated with mental health themes, and topics such as depression, disassociation or schizophrenia [@pathology], a central argument that I'll further discuss in [Results].
+More specifically, words that appear most frequently, or words with the frequency  of 0.0008511%, have the lowest mean of sentiment value-or the most negative connotation-of -2.33 on the scale between -5 and 5 in the AFINN sentiment index.
+This summary statistic further shows how SOC literature mostly associates with words that carry more negative or even extreme connotations often associated with mental health themes, and topics such as depression, disassociation or schizophrenia [@pathology], a central argument that we will further discuss in [Results].
 
@@
 
-With the large amount of different literary texts need processing by myself, it's hard to truly conduct thorough word processing.
+With the large amount of different literary texts need processing, it's hard to truly conduct thorough word processing.
 Certain texts, while can have their frequencies accounted for, can't have sentiment analysis due to their factual nature.
@@
 
-I chose literary texts based on their global reception and canonical status within the stream-of-consciousness genre.
-This decision involved forgoing certain practical considerations, such as ensuring similar text lengths across all works by each author.
-While this approach enables a thorough exploration of word frequency analysis, sentiment value, and word networks within the SOC genre, it also has the potential to emphasize the frequencies of certain words from specific authors more than others due to the longer text length of those authors.
+We chose literary texts based on their global reception and status within the stream-of-consciousness genre.
+This decision didn't consider certain practical constraints, such as ensuring similar text lengths across all works by each author.
+While this approach enables us to thoroughly explore word frequency analysis, sentiment value, and word networks within the SOC genre, it also has a downside: privileging the frequencies of certain words from specific authors more than others due to the longer text length of those authors.
 This could critically affect the integrity of the data analysis.
@@
 Furthermore, SOC literature is expansive, spanning a wide range of themes, styles, and narrative techniques.
-By exclusively representing the genre through a select group of canonical Western authors, I risk constraining the genre's diversity.
-By oversimplifying the complexity of the the dynamic literary genre, I also affect the integrity of the datasets and its correlation to mental health themes.
+By exclusively representing the genre through a select group of canonical Western authors, we risk constraining the genre's diversity.
+By oversimplifying the complexity of the the dynamic literary genre, we also affect the integrity of the datasets and its correlation to mental health themes.
 
@@
 ## The Dominant Vocabulary of SOC Literature, A Word Frequency Analysis
+
+In terms of word frequency, most words depicted in the nine selected SOC texts are abstract nouns with temporal or philosophical tendencies and negative connotations.
 
@@
 
-Within @fig-sentiment-frequency, for the NRC index, between all authors, the words with lower frequency-those ranging from 0.0000819% to 0004255%-are largely words that carry a mixture of emotional connotations such as, joy, surprise, positive, disgust and anticipation.
-However, words with high frequency-ranging from 0,0004255% to 0.0008511%-mostly contain negative connotations like fear and negative They also have a -1.875 median on a scale between -5 to 5 in the AFINN index.
+Additionally, within @fig-sentiment-frequency, for the NRC index, between all authors, the words with lower frequency-those ranging from 0.0000819% to 0004255%-are largely words that carry a mixture of emotional connotations such as, joy, surprise, positive, disgust and anticipation.
+However, words with high frequency-ranging from 0,0004255% to 0.0008511%-mostly contain negative connotations like fear and negative.
+They also have a -1.875 median on a scale between -5 to 5 in the AFINN index.
 Similarly, for the Bing index, words with higher frequency-those ranging from 0,0004255% to 0.0008511%-largely carry negative connotations.
 They also have, approximately, a -1.875 median on a scale between -5 to 5 in the AFINN index.
-This further emphasizes how SOC literature mostly contains negativity-inclined vocabulary and linguistic patterns
+This further emphasizes how SOC literature mostly contains negativity-inclined vocabulary and linguistic patterns.
 
 To further explore the contextual relationship between each word, @fig-word-networks creates word networks of all nine of these SOC texts.
-These networks will capture how often each word frequents in these SOC texts and how closely each correlates with others, whether as single words (unigrams), pairs of words (bigrams), or triplets of words (trigrams).
+These networks will capture how often each word frequents in these SOC texts and how closely each correlates with others, whether as single words (unigrams) or pairs of words (bigrams).
 In particular, the darker the connecting lines, the more correlative these combinations of words are, and the bluer the words, the more frequent they appear.
+When combined all SOC texts, similarly, temporal nouns like \textit{time, day, life, moment, night} appear the most, at, respectively, 870, 597, 567, 502 and 376 times.
+This further correlates to the non-linear nature of SOC literature and their exploration of cognitive incoherence.
+In terms of correlation, forgoing personal names like \textit{swann, odette or mme}, bigrams with a correlation of 1 include those like \textit{dark, love}; \textit{day, heart}; \textit{moment, life}; \textit{mind, life}; \textit{moment, mind}, among others.
+Observable from  @fig-word-networks, with a correlation and a word count of at least 0.8 and 400, \textit{people, moment, life} and \textit{mind} are the most predominant word combinations throughout all 9 SOC texts.
 
- providing us with a thematic understanding of SOC literature.
-
+This word frequency and correlation analysis then has enabled us to grasp the thematic framework of SOC literature, where most correlated and frequently present words are those with philosophical undertones, a sense of temporality and tinge of mental incoherence like \textit{dark, love}.
+While not entirely direct, this has somewhat portrayed SOC literature as being intricately linked to cognitive dissonance or the mind, and especially mental health issues in which SOC authors have portrayed through their characters [@flow] [@pathology] [@resampling].
 \newpage
@@
   count(word, sort = TRUE) |>
-  filter(n >= 150)
+  filter(n >= 250)
 
@@
   pairwise_cor(item = word, feature = book) |>
-  filter(correlation >= 1)
+  filter(correlation >= 0.2)
 
@@
   count(word, sort = TRUE) |>
-  filter(n >= 250)
+  filter(n >= 300)
 
@@
   pairwise_cor(item = word, feature = book) |>
-  filter(correlation >= 1)
+  filter(correlation >= 0.4)
 
@@
 #| fig-cap: "Word Networks Measured by Frequency and Correlation when Combining All Stream of Consciousness Novels"
-#| fig-subcap: ["100 Frequency, 0.2 Correlation", "200 Frequency, 0.4 Correlation", "200 Frequency, 0.8 Correlation", "400 Frequency, 1 Correlation"]
+#| fig-subcap: ["250 Frequency, 0.2 Correlation", "300 Frequency, 0.4 Correlation", "350 Frequency, 0.6 Correlation", "400 Frequency, 0.8 Correlation"]
 
@@
 
-To further understand the emotional landscape and mental health themes related to SOC literature, I've combined all nine SOC texts to conduct an overall sentiment analysis [@sentiment].
+To further understand the emotional landscape and mental health themes related to SOC literature, all nine SOC texts are combined to conduct an overall sentiment analysis [@sentiment].
 
@@
 Respectively, Woolf, Proust and Joyce tail beind at 33%, 32% and 31%, averaging out to be ~-2.2 on the emotional scale between -5 and 5.
-While most a large percentage Mansfield's works are positive, her highest percentage is negative on the AFINN index, at 24% for a scale of about -2.2 between -5 and 5.
+While a large percentage Mansfield's works are positive, her highest percentage is negative on the AFINN index, at 24% for a scale of about -2.2 between -5 and 5.
 
-This further suggests that the SOC literary landscape is inundated with negative, disgust, fear and sadness-leaning vocabulary.
+This further suggests that the SOC literary landscape mainly has negative, disgust, fear and sadness-leaning vocabulary.
 These trends reflect the incoherent mental spaces, depressive perceptions, and existential angst inherent to SOC literature.
@@
 
-ggplot(sentiment_value, aes(x = value, y = percentage, color = sentiment, group = sentiment)) +
+ggplot(sentiment_value, aes(x = value, y = percentage, fill = sentiment, color = sentiment)) +
   geom_line() +
@@
   labs(
-       x = "Sentiment Value", y = "Percentage") +
+       x = "Sentiment Value", y = "Percentage", fill = "Sentiment") +
   scale_color_manual(values = c("positive" = "lightblue", "negative" = "darkblue")) + # Adjust line colors
@@
 
+### Misconstrued Word Contexts
+
 ## Moving Forward and Next Steps
+
+# Conclusion
 

### 2024-04-23 scripts/00-simulate_data.R
Update 00-simulate_data.R

@@ -26,7 +26,7 @@ simulated_data <-
     novel = sample(novels, 12, replace = TRUE, prob = NULL),  # Randomly select novels from the pre-defined novels dataset, repeatable since work can change
     # Randomly select words from the randomly generated dataset of words that are capped at 12
     word = sample(random_words, 12, replace = FALSE, prob = NULL),  # Randomly select words from the English words dataset, not repeated since this is the count of unique words
-    #use 1 through 602 to represent each count of a word since 602 is the maximum number of a unique word appears out of all nine novels
+    # Use 1 through 602 to represent each count of a word since 602 is the maximum number of a unique word appears out of all nine novels
     n = sample(1:602, 12, replace = TRUE))
 
 # Print the simulated data 

### 2024-04-23 paper/paper.qmd
update data files

@@
 combined_text <- read_parquet(here::here("data/analysis_data/combined_text.parquet"))
-portrait <- read_csv(here::here("data/raw_data/portrait.csv"))
-all_books <- read_csv(here::here("data/raw_data/all_books.csv"))
-combined_books <- read_csv(here::here("data/raw_data/combined_books.csv"))
+portrait <- read_parquet(here::here("data/raw_data/portrait.parquet"))
+all_books <- read_parquet(here::here("data/raw_data/all_books.parquet"))
+combined_books <- read_parquet(here::here("data/raw_data/combined_books.parquet"))
 
@@
 
-# Code sourced from
-# filtered_text <- combined_text[combined_text$n >= 100, ] #only select words with a frequency larger or equal to 100
-
-# wordcloud(words = filtered_text$word, freq = filtered_text$n, scale=c(3,0.5), min.freq = 2, random.order = FALSE, colors= colorRampPalette(c("lightgray", "black"))(9))
-
 ```
@@
 
+combined books / all books
+combined correlations, combined text count
 ## Additional Data Details

### 2024-04-23 scripts/01-download_data.R

@@ -35,7 +35,7 @@ proust_books <- all_books |>
 ts_books <- all_books |>
   filter(str_detect(author, 'Eliot, T. S.'))
 
-combined_books <- rbind(joyce_books, woolf_books, dostoevsky_books, proust_books, ts_books, mansfield_books)
+combined_books <- rbind(joyce_books, woolf_books, proust_books, ts_books, mansfield_books)
 
 # Download the more famous stream-of-consciousness work of each author:
 
@@ -65,13 +65,12 @@ bliss <- gutenberg_download(c('44385', '1429')) |>
   slice(-(1:34))
 
 #### Save data ####
-write_csv(portrait, "data/raw_data/portrait.csv") 
-write_csv(swann_way, "data/raw_data/swann_way.csv") 
-write_csv(mrs_dalloway, "data/raw_data/mrs_dalloway.csv") 
-write_csv(prufrock, "data/raw_data/prufrock.csv") 
-write_csv(bliss, "data/raw_data/bliss.csv") 
-write_csv(all_books, "data/raw_data/all_books.csv") 
-write_csv(combined_books, "data/raw_data/combined_books.csv") 
-
+write_parquet(portrait, "data/raw_data/portrait.parquet") 
+write_parquet(swann_way, "data/raw_data/swann_way.parquet") 
+write_parquet(mrs_dalloway, "data/raw_data/mrs_dalloway.parquet") 
+write_parquet(prufrock, "data/raw_data/prufrock.parquet") 
+write_parquet(bliss, "data/raw_data/bliss.parquet") 
+write_parquet(all_books, "data/raw_data/all_books.parquet") 
+write_parquet(combined_books, "data/raw_data/combined_books.parquet") 
 
 

### 2024-04-23 scripts/02-data_cleaning.R

@@ -4,7 +4,7 @@
 # Date: 30 March 2024 
 # Contact: quangpono@gmail.com
 # License: MIT
-# Pre-requisites: Read four raw datasets containing the novel texts from five authors, namely `portrait.csv`, `swann.csv`. `dalloway.csv`, `bliss.csv` and `prufrock.csv`
+# Pre-requisites: Read four raw datasets containing the novel texts from five authors, namely `portrait.parquet`, `swann.parquet`. `dalloway.parquet`, `bliss.parquet` and `prufrock.parquet`
 
 #### Workspace setup ####
 library(tidyverse)
@@ -19,11 +19,11 @@ library(widyr)
 library(arrow)
 
 #### Clean data ####
-portrait <- read_csv("data/raw_data/portrait.csv")
-swann <- read_csv("data/raw_data/swann_way.csv")
-dalloway <- read_csv("data/raw_data/mrs_dalloway.csv")
-bliss <- read_csv("data/raw_data/bliss.csv")
-prufrock <- read_csv("data/raw_data/prufrock.csv")
+portrait <- read_parquet("data/raw_data/portrait.parquet")
+swann <- read_parquet("data/raw_data/swann_way.parquet")
+dalloway <- read_parquet("data/raw_data/mrs_dalloway.parquet")
+bliss <- read_parquet("data/raw_data/bliss.parquet")
+prufrock <- read_parquet("data/raw_data/prufrock.parquet")
 
 
 # Tokenize the words from the text, breaking the text down into individual words to insert them into individual datasets

### 2024-04-23 paper/paper.qmd
update paper

@@
 This involved using the identification number of the desired literary work and the `gutenberg_download` function to retrieve the HTML texts.
-Subsequently, the data was saved into individual CSV files and later converted into Parquet datasets, as demonstrated in @tbl-joyce example for James Joyce.
+Subsequently, the data was saved into individual Parquet datasets, as demonstrated in @tbl-joyce example for James Joyce.
 
@@
 
+# Create boxplots
 ggplot(frequency_long, aes(x = frequency, y = value, fill = sentiment_binary)) +
@@
 
-##first, n >=250
+## Create word network, first, n >=250
 generate_word_graph <- function(combined_text,
@@
 
-##second, n >=300
+## Create word network, second, n >=300
 generate_word_graph_second <- function(combined_text,
@@
 
-##third, n >=350
+## Create word network, third, n >=350
 generate_word_graph_third <- function(combined_text,
@@
 
-## fourth, n>= 400
+## Create word network, fourth, n>= 400
 generate_word_graph_final <- function(combined_text,
@@
 
+# Generate word networks based on previously created functions
 combined_text|>

### 2024-04-23 scripts/02-data_cleaning.R
Update 02-data_cleaning.R

@@ -99,9 +99,6 @@ prufrock_words <- prufrock_clean |>
   distinct()
 
 # Combine all texts to find the more general trend of word frequency and sentiment value
-
-## Calculate the correlations between all word combinations from all texts
-
 combined_text <- 
   rbind(portrait, swann, dalloway, prufrock, bliss)|>
   na.omit(portrait, swann, dalloway, prufrock, bliss) |>

### 2024-04-23 scripts/03-test_data.R
Update 03-test_data.R

@@ -33,7 +33,7 @@ length(combined_text$word) == 167516
 # check that the number of unique words in Mansfield's texts is 7829
 length(bliss_words$word) == 7829
 
-# check that the number of unique words in Proust's texts is 7829
+# check that the number of unique words in Proust's texts is 12203
 length(swann_words$word) == 12203
 
 # check that the number of unique words in Eliot's texts is 2350

### 2024-04-23 paper/paper.qmd
update paper

@@
 These networks will capture how often each word frequents in these SOC texts and how closely each correlates with others, whether as single words (unigrams) or pairs of words (bigrams).
-In particular, the darker the connecting lines, the more correlative these combinations of words are, and the bluer the words, the more frequent they appear.
-When combined all SOC texts, similarly, temporal nouns like \textit{time, day, life, moment, night} appear the most, at, respectively, 870, 597, 567, 502 and 376 times.
+When combined all SOC texts, temporal nouns like \textit{time, day, life, moment, night} appear the most, at, respectively, 870, 597, 567, 502 and 376 times.
 This further correlates to the non-linear nature of SOC literature and their exploration of cognitive incoherence.

### 2024-04-23 paper/paper.qmd
results updated

@@
 # Results
+
+Upon assessing the generated datasets from Project Gutenberg, we further extrapolate on them here by combining word count with the AFINN, Bing and NRC index to conduct relevant word frequency and sentiment analysis.
+The later sections of Results also compound on these findings to conduct a comparative analysis between gendered and transnational SOC authors, attempting to garner relevant insights in mental health themes in different demographics of the SOC genre.
 

### 2024-04-23 paper/datasheet_project_gutenberg.qmd
datasheet updated and discussion updated

---
title: "Datasheet for Project Gutenberg's 'Stream of Consciousness Literature'"
author: 
  - Quang Mai
thanks: "Code and data are available at: https://github.com/ponolite/stream-consciousness-language.git"
date: today
date-format: long
abstract: "A datasheet for Quang Mai's Stream of Consciousness Literature (Exploring Word Frequency, Sentiment Value and Mental Health Themes in the Works of Joyce, Woolf, Proust, Mansfield and Eliot from Project Gutenberg), orginal textual data available under Project Gutenberg archive [@rGuten] [@gutenberg]. The questions used to form this datasheet were extracted from @gebru2021datasheets. Created using the R statistical programming language [@r]."
format: pdf
number-sections: true
bibliography: references.bib
toc: true
---

# Motivation

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset or HTML texts were created to be archived as literary texts by Project Gutenberg. We were unable to find a publicly available dataset in a structured format that contained the information on SOC literature and their correlation to mental health themes in the West's late 19th to mid-20th century, thus necessitating a need for these datasets.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - Volunteer contributors and proofreaders of Project Gutenberg initially created the dataset or the HTML text files. Later, the author Quang Mai created his own datasets using the HTML of 9 selected SOC novels available on Project Gutenberg using the package `gutenbergr` [@rGuten] [@gutenberg]. These texts were created for the author's own purpose of analyzing their linguistic patterns and contents and their relationship to mental health themes.
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - The creation of these datasets were not funded by anyone.
4. *Any other comments?*
    - N/A

# Composition

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	-  Each row of the raw dataset represents a the Project Gutenberg identification number for that respective SOC text, a line extracted from the HTML text of that SOC text (for instance, "April is the cruellest month, breading" by T.S. Eliot), the book in which the SOC text originates from and the author's name. As each SOC text spans chapters, each dataset spans multiple rows as there are many lines within that text.
2. *How many instances are there in total (of each type, if appropriate)?*
	- There is only 1 type of data, which is the bibliographic data of each SOC text. To support this data type, there are 5 instances (each combining to represent a text's bibliography, like author's name, book title, each line of the text and book ID).
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- The dataset contains all possible instances from the larger set of HTML text data from Project Gutenberg, including all of the abovementioned ones.
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- Each instance consists of a number to represent the book ID (9 IDs), a literary line sourced from the 9 HTML texts from Project Gutenberg (167516 lines in total), the names of the SOC texts (9 text names in total, since all authors beside Proust have 2 texts), and the author's name (9 authors).
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- Each instance is eponymously titled.
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- There is no missing information from individual instance since all texts have been in the public domain for a while, thus all data is up-to-date.
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- Yes, the relationships between individual instances are made explicit through their eponymously titled labels and the book IDs as provided from Project Gutenberg, which helps amalgamate data by each individual SOC text. 
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- The data is split into individual SOC authors. This split ensures a comparative analysis between different demographics of SOC authors later on, for instance, between transnational SOC authors and between gendered SOC authors, enabling a more nuanced analysis of the dataset.
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- No, there are no apparent errors or sources of noise. There are redundancies, although
these are purposeful and provide added meaning to the data (for example, the repeating book ID that helps ensure certain data instances ony belong to certain books).
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- The dataset is strictly self-contained, relying only on metrics provided by the Project Gutenberg. However, it will change depending on changes in legislation and regulation of the archive itself.
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- The data does not provide any confidential information, since all texts are works of the public domain as disclosed on Project Gutenberg's website [@gutenberg].
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- No, the dataset does not include information that would harm a viewer. However, certain words may trigger certain emotional responses since not all literary SOC texts might align with a viewer's point of view.
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- No, it does not categorize texts based on sub-populations like age, gender, or region.
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- It is possible to identify individuals from this dataset because the author names are disclosed, however, this information should be harmless since it's in the public domain.
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- Based on the diverse literary contents, the dataset might contain sensitive information. However, as stressed, since all information is entered into public domain, this information should be accessible to all.
16. *Any other comments?*
	- N/A

# Collection process

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- The data associated with each instance uses a mix of information from two sources:
individual submission of SOC texts by individual volunteers who went through a rigorous process of transcript submission to Project Gutenberg and voluntary proofreaders that ensure each SOC text underwent professional editing before being archived [@gutenberg]. The individual submissions are processed using internal submission tool from Project Gutenberg.
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- Most new Project Gutenberg eBooks are digitized editions of printed books in HTML format, which is validated using the W3C’s online validator. Additionally, the archive prefers plain text versions whenever possible due to their universal accessibility, and long-term usability. PDF, Word, and other word processor formats are avoided as master formats due to the challenges involved in converting them to HTML and updating them [@gutenberg].
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- There was no sampling strategy employed. The data is an inclusive and comprehensive
statistical summary of all SOC texts included.
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- Data collection or book digitization was conducted by both voluntary contributors and proofreaders from Project Gutenberg, which means no monetary compensation is viable.
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- The data is collected continuously and there is no set timeframe. However, since Project Gutenberg doesn't accept copyrighted works but only works in the public domain, the majority of SOC texts were published over 95 years ago.
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- It is unclear whether there was an ethical review of this data, however there is a clearly disclosed submission guideline on Project Gutenberg's website that touched on this aspect [@gutenberg]
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- The data was obtained via Project Gutenberb, a volunteer archive based in the United States.
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- No, as the data is anonymized and does not contain any confidential or sensitive
information the individuals that form this data were not notified.
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- All sourced SOC literary texts are in the public domain, thus no consent is required.
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- United States’s privacy laws may allow individuals to request the deletion of their data
from systems, although it is unclear whether these laws apply to the works in the public domain.
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- No, no analysis was provided.
12. *Any other comments?*
	- N/A

# Preprocessing/cleaning/labeling

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- No preprocessing/cleaning/labeling on the data was done by Project Gutenbeg.
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- N/A
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- N/A
No, it is unclear what software was used to preprocess, clean, and label the data.
4. *Any other comments?*
	- N/A

# Uses

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- There is no feasible way of determining how the data was previously or is currently
used. Project Gutenberg allows any person to utilize the dataset. 
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- No, there is no public log of any papers or systems that currently use the dataset.
3. *What (other) tasks could the dataset be used for?*
	- The data set can be used to conduct topic modeling of the texts involved, especially in regards to SOC literature.
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- No.
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- No. The purpose of Project is to allow the public to use the literary data as they see fit.
6. *Any other comments?*
	- No.

# Distribution

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- The dataset is freely distributed using the `gutenbergr` package accessible using R [@rGuten]. It is also accessible on Project Gutenberg's website found here: https://www.gutenberg.org.
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- The dataset is distributed on Project Gutenberg archive.
3. *When will the dataset be distributed?*
	- The dataset has no set schedule for distribution or updates.
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- The dataset is provided and licenced under Project Gutenberg.
found here: https://www.gutenberg.org.
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- No, the data is non-profit and in the public domain so no third-parties have a legal claim to the data.
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- The dataset is governed by the aforementioned Project Gutenberg Licence provided by the archive based in the United States, and may be further restricted by municipal, provincial, or federal legislation.
7. *Any other comments?*
	- N/A

# Maintenance

1. *Who will be supporting/hosting/maintaining the dataset?*
	- Project Gutenberg solely maintain and update the data, while also supporting and hosting the textual data of SOC literature.
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- The archive may be contacted via this mailing address Project Gutenberg Literary Archive Foundation, PO Box 16327, Salt Lake City, UT 84116.
3. *Is there an erratum? If so, please provide a link or other access point.*
	- No.
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- Since the literary texts have underwent significant proofreading and editing and they are merely digitized versions of physical copies of books, it's unclear if the dataset or the HTML texts will be updated.
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- No.
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- It's unclear if older versions of the dataset continue to be supported, hosted or maintained.
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- With open data, there are always mechanisms for others to extend, augment, or build upon the dataset. However, there is no formal mechanism to bring these changes to the Project Gutenberg platform or the official dataset. Users may take the dataset and augment it on their own.
8. *Any other comments?*
	- N/A

\newpage

# References

### 2024-04-23 paper/paper.qmd
paper updates
datasheet updated and discussion updated

@@
 Thus, the estimand is the correlation between mental health-related themes and words in SOC literature, their frequency and sentiment value as provided by @rTidytext.
-This is considered in terms of nine selected SOC novels only, namely Joyce's \textit{A Portrait of the Artist as a Young Man and Chamber Music}; Woolf's \textit{Mrs Dalloway and Jacob's Room}; Proust's \textit{Swann Way}; Mansfield's \textit{Bliss and The Garden Party}; and Eliot's \textit{The Waste Land and The Love Song of J.
+This is considered in terms of nine selected SOC literary works only, namely Joyce's \textit{A Portrait of the Artist as a Young Man and Chamber Music}; Woolf's \textit{Mrs Dalloway and Jacob's Room}; Proust's \textit{Swann Way}; Mansfield's \textit{Bliss and The Garden Party}; and Eliot's \textit{The Waste Land and The Love Song of J.
 Alfred Prufrock}.
-Through our analysis, we found that (percentage, number and data here, main results)...
-Analysis of the SOC genre shows that higher frequency words are mostly negative (-1.875 median on AFINN index); sentiment-wise, analysis reveals 11% to 15% of words in SOC texts are sadness-inclined (NRC index), and over 60% of words in selected SOC literature are negative, with Eliot's works being the most negative at 68% (Bing index).
-Temporal nouns like "time," "day," "life," "moment," and "night" are most frequent, correlating with SOC's non-linear nature,
+Through our analysis, we found that words that appear with high frequency are mostly negative in select SOC literature (-1.875 median on AFINN index).
+Sentiment-wise, analysis reveals a predominant 27% to 37% of words in SOC texts are negative and sadness-inclined (NRC index), and over 60% of words in these texts are negative, with Eliot's works being the most negative at 68% (Bing index), followed closely by Joyce, Woolf, Proust, and Mansfield.
+Abstract and temporal nouns like "time," "day," "life," "moment," and "night" are most frequent, respectively appearing 870, 597, 567, 502 and 376 times in all texts combined.
+These statistics correlate with SOC's non-linear nature, and the genre's exploration of the unconscious mind which often correlates to mental health themes.
 
-Analysis of SOC literature reveals lower frequency words have mixed emotional connotations, while higher frequency words tend to be negative (-1.875 median on AFINN index).
-Temporal nouns like "time," "day," "life," "moment," and "night" are most frequent.
-Sentiment analysis shows SOC literature is predominantly negative, with Eliot's works being the most negative (37% on AFINN index), followed closely by Joyce, Woolf, Proust, and Mansfield.
-The NRC index indicates a prevalence of negative emotions, particularly sadness-related words.
-
-To further understand the correlation between stream of consciousness novels and mental health themes, in [Introduction], the paper briefly discusses the nature of stream of consciousness literature, relevant authors and the works that I've chosen to analyze.
+To further understand the correlation between stream of consciousness texts and mental health themes, in [Introduction], the paper briefly discusses the nature of stream of consciousness literature, relevant authors and the works that I've chosen to analyze.
 Subsequently, in [Data] and [Results], we talk about the nature of the data obtained and analyze the results garnered from the data with suitable tables and charts.
@@
 Data was generated, extracted and cleaned using the open-source statistical programming language R [@r], leveraging functions from `tidyverse` [@rTidyverse], `tidytext` [@rTidytext], `rmarkdown` [@rMark], `dplyr` [@rDplyr], `ggplot2` [@rGgplot2], `scales` [@rScales], `here` [@rHere], `igraph` [@rIgraph], `widyr` [@rWidy], `ggraph` [@rGg], `textdata` [@rTextdata], `tm` [@rTm], `here` [@here], `kableExtra` [@kableExtra], `arrow` [@arrow], and `knitr` [@rKnitr].
+More information on the dataset or the HTML textual data sourced from Project Gutenberg is available in the documented datasheet available here within this folder, paper/datasheet_project_gutenberg.pdf, located within the paper's GitHub repository.
 
@@
 This involved using the identification number of the desired literary work and the `gutenberg_download` function to retrieve the HTML texts.
-Subsequently, the data was saved into individual Parquet datasets, as demonstrated in @tbl-joyce example for James Joyce.
+Subsequently, the data was saved into individual Parquet datasets, as demonstrated in @tbl-joyce.
 
@@
 #| label: tbl-joyce-token
-#| tbl-cap: Tokenized Novel Text & Word Count after Tokenization (James Joyce)
-#| tbl-subcap: ["Tokenized Text", "Word Count"]
-#| layout-ncol: 2
+#| tbl-cap: Tokenized Stream of Consciousness Text (James Joyce)
 #| warning: false
@@
 
-portrait_clean <- portrait_clean |>
-  select(-book)
-
 portrait_clean |>
   rename("Book ID" = gutenberg_id,
+         "Book" = book,
          "Author" = author,
@@
   kable(
-    col.names = c("Book ID", "Author", "Word"),
+    col.names = c("Book ID", "Book", "Author", "Word"),
     booktabs = TRUE
@@
                 full_width = FALSE)
+
+```
+
+Next, in @tbl-joyce-token, to measure the word count of each literary text, we first tokenized or separated into individual chunks of words using `unnest_tokens()` [@rTidytext].In addition, to account for stop words like, a, of, and, etc., the dataset `stop_words`and the function `anti_join` from @rTidytext were used to exclude all stop words from the self-generated datasets of literary texts, ensuring data validity and meaningful analysis.
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-joyce-count
+#| tbl-cap: Word Count after Tokenization (James Joyce)
+#| warning: false
+#| message: false
 
@@
 
-Next, observable in @tbl-joyce-token, to measure the word count of each literary text, we first tokenized or separated into individual chunks of words using `unnest_tokens()` [@rTidytext].In addition, to account for stop words like, a, of, and, etc., the dataset `stop_words`and the function `anti_join` from @rTidytext were used to exclude all stop words from the self-generated datasets of literary texts, ensuring data validity and meaningful analysis.
-
-Then, to quantify the number of times each word appears in each text, we used `count()` (@tbl-joyce-token). The result manifests into @fig-count.
+Then, to quantify the number of times each word appears in each text, we used `count()` (@tbl-joyce-count).
+The result manifests into @fig-count, which visualizes the top 20 words with the highest word count in each SOC author's texts.
 
@@
 #| label: tbl-boxplot 
-#| tbl-cap: "Overall Boxplot Summary Statistics of all Word Frequency Percentages (AFINN Sentiment)"
+#| tbl-cap: "Overall Boxplot Summary Statistics of all Word Frequency Percentages (AFINN Index)"
 #| warning: false
@@
 ## Data Limitations
-
-### Lack of Thorough Word Cleaning
-
-With the large amount of different literary texts need processing, it's hard to truly conduct thorough word processing.
-Certain texts, while can have their frequencies accounted for, can't have sentiment analysis due to their factual nature.
-These textual elements might include things like chapter titles, Roman numbers and personal names, which can strongly skew and affect the data integrity.
-As observed in @fig-count, "Swann", "Stephen" or "Jacob" are the most repeated words in Proust's, Joyce's and Woolf's, proving this data limitation.
 
@@
 While not entirely direct, this has somewhat portrayed SOC literature as being intricately linked to cognitive dissonance or the mind, and especially mental health issues in which SOC authors have portrayed through their characters [@flow] [@pathology] [@resampling].
-\newpage
+
 ```{r}
@@
 The dominant emotional connotation in SOC literature's vocabulary is negative, standing at 16% to 22%.
-While the second highest percentage range belong to words with positive connotation, ranging from 11% to 17%, most of the other bars are occupied by words with negative-leaning emotions, such as sadness, fear, and anger.
+While the second highest percentage range belong to words with positive connotation, ranging from 11% to 17%.
+However, most of the other bars are occupied by words with negative-leaning emotions, such as sadness, fear, and anger.
 Sadness-related words, in particular, ranges the third highest, from 11% to 15%.
@@
   scale_color_gradient(limits = c(0, 0.001), 
-                       low = "lightblue", high = "darkblue") +
+                       low = "steelblue", high = "darkblue") +
   facet_wrap(~author, ncol = 1) +
@@
   scale_color_gradient(limits = c(0, 0.001), 
-                       low = "lightblue", high = "darkblue") +
+                       low = "steelblue", high = "darkblue") +
   facet_wrap(~author, ncol = 2) +
@@
 ```
-
-### Comparing SOC Literature's Female Authors
-
-### Comparing SOC Literature's Male Authors
 
@@
 
-Discuss vocabulary patterns and word trends
+The findings in this paper reveals a correlative relationship between SOC literature and mental health-inclined vocabulary and linguistic patterns.
+In terms of frequency, a dominant portion-or over 60% of words in the Bing index-of select SOC texts in this analysis have negative and sadness-leaning words ([@fig-sentiment-frequency]) ([@fig-sentiment-category]) ([@fig-sentiment-number]).
+Most repeated words are abstract, temporal nouns like \textit{time}, \textit{mind} or \textit{life}, which are aligned with SOC literature's exploration of mind wandering, non-linearity and unconscious cognition ([@fig-word-networks]).
+
+As scholars Bernini and Fernyhough mentioned, the "mysterious nature of the conscious" is "a simultaneously fragmented and unified mental realm" [@resampling], this analysis has also reflected that sentiment in SOC literature.
+Most of the texts express the desire to replicate the inner workings of the human mind textually, attempting to capture the complexity of thought processes and emotions.
+With these inner-workings revealed to be mostly negative and existential, SOC authors have explored their identity and worldview in a self-aware manner, exploring then-taboo topics such as existential angst, mental health issues and disabilities to their fullest extent.
 
@@
 
-pathological consequences of the constant
-shifts of identity in migrant characters. Most postcolonial studies contend that
-marginalization occasioned by the dominant group result in neurotic conditions in
-members of the minority group.
+Through the analysis, we can see the SOC texts' linguistic patterns reflect the socio-political landscape of the Western hemisphere.
+The SOC genre stems out of modernist literature, whose writers often rebelled against linear and traditional narrative techniques of the 19th century [@resampling].
+According to scholar Nyongesa, "marginalization occasioned by the dominant group result in neurotic conditions in members of the minority group" [@pathology], which rings true to the scenario at hand about SOC literature and its relationship to its birthplace: late 19th to mid-20th century Western society.
 
-The novels' linguistic patterns reflect the socio-political landscape of the Western hemisphere, from  late 19th century to the mid-20th century.
+There are several reasons rooted in the West's modernist era that resulted in the  growth and popularity of the SOC genre.
+This period's rapid industrialization and urbanization  reshaped the traditional institutions of Western society, leading to fragmentation and dislocation, which SOC authors had attempted to capture in their own works [@flow].
+Advancing transportation, photographic media and mass communication affected the West's sense of time and space, which in turn affected how people understood reality.
+This is a central pillar of SOC literature: distorted reality, atemporality and an unconscious perception of the world.
 
-Stream of consciousness literature emerged as a significant literary technique in the early 20th century, particularly associated with the modernist movement in literature.
-Several factors contributed to its rise and prominence:
-
-Psychological Realism: One of the primary motivations for the development of stream of consciousness was the desire to represent the inner workings of the human mind more accurately.
-Writers sought to capture the complexity of thought processes, emotions, and perceptions in a more realistic and nuanced manner.
-Response to Realism and Naturalism: Stream of consciousness literature can be seen as a reaction against the more structured and external focus of realism and naturalism.
-Instead of portraying characters and events objectively, authors wanted to delve deeper into subjective experiences and individual consciousness.
-Technological and Cultural Changes: The early 20th century was a period of rapid technological advancement and social change, including the rise of psychology as a discipline.
-These developments influenced writers to explore new narrative techniques and modes of representation.
-Literary Experimentation: Modernist writers were often interested in pushing the boundaries of traditional narrative forms and experimenting with language, style, and structure.
-Stream of consciousness offered a way to challenge conventional storytelling techniques and engage readers in more innovative ways.
-Exploration of Identity and Perception: Stream of consciousness literature allowed writers to explore themes related to identity, consciousness, and perception.
-By immersing readers in the internal thoughts and sensations of characters, authors could interrogate questions about the nature of reality, self-awareness, and the construction of meaning.
-Overall, stream of consciousness literature became a prominent feature of modernist writing due to its capacity to convey the complexity of human experience and its ability to reflect the uncertainties and ambiguities of the modern world.
-
-The rise of modernism in literature, art, and culture during the late 19th and early 20th centuries was influenced by several interconnected factors:
-
-Industrialization and Urbanization: The rapid industrialization and urbanization of society brought about significant changes in the way people lived, worked, and interacted.
-This upheaval in traditional social structures and lifestyles led to a sense of dislocation and fragmentation, which modernist artists and writers sought to capture in their work.
-Technological Advances: Advances in technology, particularly in transportation, communication, and mass media, reshaped human experience and perception of time and space.
-Innovations such as the telegraph, telephone, and photography altered how people communicated, experienced distance, and understood reality.
-Crisis of Faith and Meaning: The late 19th and early 20th centuries witnessed a crisis of faith and meaning, as traditional religious and philosophical beliefs came into question.
-Developments such as Darwin's theory of evolution, Freudian psychology, and existential philosophy challenged established notions of human existence, morality, and purpose.
-World Wars and Political Turmoil: The turmoil of World War I and the interwar period brought about profound changes in society, politics, and culture.
-The devastation of war, the collapse of empires, and the emergence of new ideologies (such as communism and fascism) shattered old certainties and created an atmosphere of disillusionment and uncertainty.
-Intellectual and Cultural Exchange: The modernist movement was characterized by a spirit of intellectual and cultural exchange, with artists and writers drawing inspiration from diverse sources, including non-Western cultures, folk traditions, and avant-garde movements.
-This openness to new ideas and influences fostered experimentation and innovation in the arts.
-Quest for Individuality and Authenticity: Modernism rejected the conventions and constraints of traditional artistic forms and sought to express individual subjectivity and experience in new and unconventional ways.
-Artists and writers embraced themes of alienation, identity, and the search for authenticity in an increasingly fragmented and alienating world.
+This time period also underwent the West's crisis of faith in traditional religious beliefs.
+With the emergence of Darwin's theory of evolution, Freudian psychology, and existential philosophy came different notions of human existence, morality, and meaning [@resampling].
+The heightened search for individuality of modernism had SOC authors embraced themes of alienation, identity and mental health in an increasing fragmented world.
+The devastation from World War I and the interwar period also further intensified this crisis.
+Post-war collapse of nations created new ideologies like communism and fascism.
+This in turn created an atmosphere of disillusionment and uncertainty that SOC literature attempted to answer with its disassociative and schizophrenic style.
 
@@
 
-This includes words such as chapter titles, Roman numbers and personal names, affecting the integrity of data analysis ("Swann" being the most common word)
+With the large amount of different literary texts need processing, it's hard to truly conduct thorough word processing.
+Certain texts, while can have their frequencies accounted for, can't have sentiment analysis due to their factual nature.
+These textual elements might include things like chapter titles, texts from a language other than English, Roman numbers and personal names, which can strongly skew and affect the data integrity.
+As observed in @fig-count, "Swann", "Stephen" or "Jacob" are the most repeated words in Proust's, Joyce's and Woolf's, proving this data limitation.
 
-### Decontextualized Literature Works and Limiting Publication Editions
+### Decontextualized Words and Limiting Publication Editions
 
-Words are singled out and analyzed without context which could have affected their intended meanings, especially in such a complex genre such as stream of consciousness.
-The limiting novel editions also doesn't make sure that their literary integrity are maintained and the analysis might have missed important texts of other editions.
+Throughout our analysis, words are singled out and analyzed without contexts which could have affected their intended meanings, especially in such a complex genre such as SOC.
+For instance, the commonly repeated words \textit{time}, \textit{mind} or \textit{life} from selected texts when combined with other words or maybe insignificant to the text meaning.
+However,
+
+In addition, the limiting literary text editions from Project Gutenberg also  [@gutenberg] disallow us to ensure that all texst are integral to their 't make sure that their literary integrity are maintained and the analysis might have missed important texts of other editions.
 
@@
 
-### Reliance on Three Different Indices
+### Reliance on A Small Number of Sentiment Indices
 
-Since the novels are chosen based on their worldwide reception and canon, practical constraints like how all novels should be of similar length are ignored.
+With a diverse selection of three indices, namely Bing, NRC and AFINN, to carry out a thorough sentiment analysis of selected SOC texts,Since the novels are chosen based on their worldwide reception and canon, practical constraints like how all novels should be of similar length are ignored.
 This constraint can prevent one novel having more words than others, which can critically affect the integrity of the word frequency analysis, sentiment value and word networks where one novel dominates the others and skews the results.
@@
 
-### Misconstrued Word Contexts
+## Moving Forward and Next Steps
 
-## Moving Forward and Next Steps
+Moving forward, further research with longitudinal studies and diverse data sources including qualitative interviews, alongside a mixed methods approach, is crucial to gain a comprehensive understanding of childlessness, its nuances, and its multifaceted implications.
+This knowledge can inform future policies aimed at fostering a supportive environment for individuals and families as they navigate the complex decisions around childbearing in the 21st century.
 

### 2024-04-24 paper/paper.qmd
appendix updates

@@
 It mirrors the unpredictable and interconnected nature of human thought processes, often revealing the inner workings of the character's mind in an intimate and unfiltered manner [@flow].
-In literature, most scholars agree that stream of consciousness reveals the complexities of mind-scapes, shedding light on the nuances of characters' emotional well-being and psychological struggles [@pathology].
+In literature, most scholars agree that stream of consciousness reveals the complexities of cognition, shedding light on the nuances of characters' emotional well-being and psychological struggles [@pathology].
 As such, this paper has mined the texts of a total of nine novels from the volunteer archive, Project Gutenberg, to examine the mental health themes of famous stream of consciousness authors, namely by Joyce, Woolf, Proust, Mansfield and Eliot, from the modernist era of literature, spanning from late 19th century to mid-20th century [@gutenberg].
@@
 To collect text data from five SOC authors using the `gutenbergr` package, we first downloaded the index of each author's works from the Project Gutenberg database.
-These authors include Joyce, Mansfield, Eliot, Woolf, and Proust [@rGuten] ([Appendix]).
+These authors include Joyce, Mansfield, Eliot, Woolf, and Proust [@rGuten] ([Appendix]'s @tbl-combined and @tbl-all).
 
@@
          "Word" = word) |>
-  tail(3) |>
+  tail(5) |>
   kable(
@@
 These networks will capture how often each word frequents in these SOC texts and how closely each correlates with others, whether as single words (unigrams) or pairs of words (bigrams).
-When combined all SOC texts, temporal nouns like \textit{time, day, life, moment, night} appear the most, at, respectively, 870, 597, 567, 502 and 376 times.
+When combined all SOC texts, temporal nouns like \textit{time, day, life, moment, night} appear the most, at, respectively, 870, 597, 567, 502 and 376 times ([Appendix]'s @tbl-count).
 This further correlates to the non-linear nature of SOC literature and their exploration of cognitive incoherence.
-In terms of correlation, forgoing personal names like \textit{swann, odette or mme}, bigrams with a correlation of 1 include those like \textit{dark, love}; \textit{day, heart}; \textit{moment, life}; \textit{mind, life}; \textit{moment, mind}, among others.
+In terms of correlation, forgoing personal names like \textit{swann, odette or mme}, bigrams with a correlation of 1 include those like \textit{dark, love}; \textit{day, heart}; \textit{moment, life}; \textit{mind, life}; \textit{moment, mind}, among others ([Appendix]'s @tbl-correlation).
 Observable from  @fig-word-networks, with a correlation and a word count of at least 0.8 and 400, \textit{people, moment, life} and \textit{mind} are the most predominant word combinations throughout all 9 SOC texts.
@@
 ```
-
+\newpage
 ## Sentiment Analysis
@@
 Most of the texts express the desire to replicate the inner workings of the human mind textually, attempting to capture the complexity of thought processes and emotions.
-With these inner-workings revealed to be mostly negative and existential, SOC authors have explored their identity and worldview in a self-aware manner, exploring then-taboo topics such as existential angst, mental health issues and disabilities to their fullest extent.
+With these inner-workings revealed to be mostly negative and existential, SOC authors have explored their identity and worldview in an authentic manner, exploring then-taboo topics such as existential angst, mental health issues and cognitive dissonance.
 
@@
 These textual elements might include things like chapter titles, texts from a language other than English, Roman numbers and personal names, which can strongly skew and affect the data integrity.
-As observed in @fig-count, "Swann", "Stephen" or "Jacob" are the most repeated words in Proust's, Joyce's and Woolf's, proving this data limitation.
+As observed in @fig-count, \textit{Swann}, \textit{Stephen} or \textit{Jacob} are the most repeated words in Proust's, Joyce's and Woolf's, proving this data limitation.
 
-### Decontextualized Words and Limiting Publication Editions
+### Decontextualized Words and Limiting Literature Selection
 
 Throughout our analysis, words are singled out and analyzed without contexts which could have affected their intended meanings, especially in such a complex genre such as SOC.
-For instance, the commonly repeated words \textit{time}, \textit{mind} or \textit{life} from selected texts when combined with other words or maybe insignificant to the text meaning.
-However,
+For instance, the commonly repeated words \textit{time}, \textit{mind} or \textit{life} in our selected SOC texts might mean something entirely different when combined with other words.
+Paradoxically, these words may be insignificant to the SOC texts in question themselves as they're only small units of large literary texts that often have convoluted and complex themes.
+Therefore word frequency analysis might have excluded nuances when analyzing the contexts behind mental health themes and their relationship to SOC literature.
 
-In addition, the limiting literary text editions from Project Gutenberg also  [@gutenberg] disallow us to ensure that all texst are integral to their 't make sure that their literary integrity are maintained and the analysis might have missed important texts of other editions.
-
-While these datasets aim to illuminate on the general time trends of childlessness with each variable measured, they are often de-contextualized for the same purpose.
-For instance, in Table 6, there is no clarifying note on whether people who declared themselves as ‘divorced’ are divorced once or more.
-This can be extremely narrowing, and exclude nuances when analyzing the contexts behind childlessness.
+In addition, the limiting SOC text editions or SOC literature from Project Gutenberg also disallow us from obtaining large sample of SOC literature to analyze their relationship to mental health themes [@gutenberg].
+Previously mentioned, since Project Gutenberg only accepts works in the public domain, their archive of literature includes solely old, Western and canonical works that might have missed a diverse range of other SOC texts or other editions that could have garnered better insights into our analysis.
+In addition, not all SOC texts in the archive are available to be transcribed into datasets due to licensing issues, further limiting our analysis.
 
@@
 
-With a diverse selection of three indices, namely Bing, NRC and AFINN, to carry out a thorough sentiment analysis of selected SOC texts,Since the novels are chosen based on their worldwide reception and canon, practical constraints like how all novels should be of similar length are ignored.
-This constraint can prevent one novel having more words than others, which can critically affect the integrity of the word frequency analysis, sentiment value and word networks where one novel dominates the others and skews the results.
-
-In addition, these authors work are expansive and so choosing a select few to bind them to the SOC genre can be limiting as the SOC genre in itself is already an amalgamation of different literary trends.
-Thus, this can affect the integrity of the datasets.
+To ensure a thorough sentiment analysis of SOC texts, this paper has leveraged three widely used sentiment indices, namely Bing, NRC and AFINN.
+However, a sole dependence on the emotional metrics of these three indices might not be nuanced enough.
+This is because to categorize words into a specific emotion is an arduous and complex tasks as emotions are spectral and words are mainly context-based.
+In addition, some of the indices used might not ensure valid data analysis.
+For instance, @tbl-frequency shows that, using the NRC index, the word \textit{accomplished} can carry two different emotions: joy and positive.
+This flexible categorization can very much skew the results of the sentiment analysis, bringing an emotion more to the forefront than the others.
 
@@
 
-Moving forward, further research with longitudinal studies and diverse data sources including qualitative interviews, alongside a mixed methods approach, is crucial to gain a comprehensive understanding of childlessness, its nuances, and its multifaceted implications.
-This knowledge can inform future policies aimed at fostering a supportive environment for individuals and families as they navigate the complex decisions around childbearing in the 21st century.
+Moving forward, to gain a deeper understanding of the relationship between SOC literature and mental health themes, more diverse and better natural language processing (NLP) indices, methods and processes must be used to conduct a more accurate analysis.
+
+As mentioned, relying on three sentiment indices are not enough to accurately capture the emotional sentiments of the words used in SOC literature due to how context-based words are, thus more relevant indices should be incorporated into the analysis.
+
+In addition, due to the abstract nature of SOC texts, using advanced NLP models is more effective for analysis rather than focusing solely on small linguistic units such as words or word frequency.
+Topic modeling, for example, is a technique in NLP used to uncover abstract topics within texts like SOC literature.
+Leveraging algorithms like Latent Dirichlet Allocation (LDA) and Non-negative Matrix Factorization (NMF), topic modeling helps identify patterns of co-occurring words within SOC literature, grouping them into clusters that represent coherent themes or topics that can be relevant to our mental health analysis of the SOC genre.
+
+Lastly, due to the qualitative nature of SOC literature, using a more mixed methods approach, combining both quantitative and qualitative data analysis, will offer a more comprehensive understanding of both the broader mental health trends and the literary genre itself.
 
 # Conclusion
+
+This study analyzes stream of consciousness (SOC) literature's linguistic patterns, revealing its  portrayal of mental health themes.
+Through our analysis of works by Joyce, Woolf, Proust, Mansfield, and Eliot, we uncover that SOC literature is largely negative or sadness-inclined with a philosophical and temporal undertones due to its exploration of unconscious cognition.
+While our quantitative analysis provides insights into the genre's relationship with mental health themes, it's important to recognize its limitations in capturing the richness of SOC language.
+This study therefore asks for further exploration into the relationship between SOC literature and mental health discourse by using more advanced NLP methods.
+We hypothesize that this can provide deeper insights into the genre, and the reasons for its growth in the West's late 19th to mid-20th century society.
 
@@
 
-combined books / all books
-combined correlations, combined text count
-## Additional Data Details
+## All Books from Project Gutenberg Archive
 
 ```{r}
-##| eval: true
-##| echo: false
-##| message: false
-##| warning: false
-#combined_books 
-```
+#| echo: false
+#| eval: true
+#| label: tbl-all
+#| tbl-cap: "A Sample of Project Gutenberg Archive's Book Database"
+#| warning: false
 
-### Data Gathering
+all_books <- all_books[5:10, ]
 
-```{r fig.pos="H"}
-##| echo: false
-##| message: false
-##| label: tbl-reasons-strip-search
-##| tbl-cap:
+all_books |>
+  select(-gutenberg_author_id, -rights, -language, -gutenberg_bookshelf) |>
+  rename("Book ID" = gutenberg_id,
+         "Title" = title,
+         "Author" = author,
+         "Text Included" = has_text) |>
+  tail(5) |>
+  kable(
+    col.names = c("Book ID", "Title", "Author", "Text Included"),
+    booktabs = TRUE
+  ) |>
+  kable_styling(latex_options = c("striped", "hold_position", "scale_down"), 
+                bootstrap_options = "striped")
 
@@
 
-### Data Cleaning
+## Combined Bibliographies of Nine Selected Stream of Consciousness Texts
 
-```{r fig.pos="H"}
-##| echo: false
-##| message: false
-##| label: tbl-items-strip-search
-##| tbl-cap:
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-combined
+#| tbl-cap: "Sample Combined Bibliographies of 9 Selected Stream of Consciousness Texts"
+#| warning: false
+
+combined_books |>
+   select(-gutenberg_author_id, -rights, -language, -gutenberg_bookshelf) |>
+  rename("Book ID" = gutenberg_id,
+         "Title" = title,
+         "Author" = author,
+         "Text Included" = has_text) |>
+  head(5) |>
+  kable(
+    col.names = c("Book ID", "Title", "Author", "Text Included"),
+    booktabs = TRUE
+  ) |>
+  kable_styling(latex_options = c("striped", "hold_position"), 
+                bootstrap_options = "striped", 
+                full_width = FALSE)
+
+```
+
+## Combined Statistcs
+
+### Combined Count
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-count
+#| tbl-cap: "Sample Combined Count of Words from All Nine Stream of Consciousness Texts"
+#| warning: false
+
+combined_text_count <- combined_text |>
+  count(word, sort = TRUE) |>
+  filter(n >= 400)
+
+combined_text_count |>
+  rename("Word" = word,
+         "Count" = n) |>
+  head(5) |>
+  kable(
+    col.names = c("Word", "Count"),
+    booktabs = TRUE
+  ) |>
+  kable_styling(latex_options = c("striped", "hold_position", "scale_down"), 
+                bootstrap_options = "striped", 
+                full_width = FALSE)
+
+```
+
+### Combined Correlation
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-correlation
+#| tbl-cap: "Sample Combined Bibliographies of 9 Selected Stream of Consciousness Texts"
+#| warning: false
+
+combined_correlations <- combined_text |>
+  semi_join(combined_text_count, by = "word") |>
+  pairwise_cor(item = word, feature = book) |>
+  filter(correlation >= 0.8)
+
+combined_correlations |>
+  rename("Word 1" = item1,
+         "Word 2" = item2,
+         "Correlation" = correlation) |>
+  head(5) |>
+  kable(
+    col.names = c("Word 1", "Word 2", "Correlation"),
+    booktabs = TRUE
+  ) |>
+  kable_styling(latex_options = c("striped", "hold_position"), 
+                bootstrap_options = "striped", 
+                full_width = FALSE)
 

### 2024-04-24 paper/paper.qmd
final paper updates

@@
 By conducting word frequency analysis and sentiment analysis on these authors' nine novels, we aim to uncover shared linguistic patterns and gain insights into the authors' mental states.
-Our analysis of the SOC genre shows that higher frequency words are mostly negative and carry sadness or fear-inclined connotations.
+Our analysis of the SOC genre shows that higher frequency words are mostly negative and carry sadness or fear-inclined connotations, with clear linguistic differences between female and male authors.
 With these trends uncovered, we attempt to offer insights into themes of self-identity, anxiety, disassociation and existential contemplation within Western society and its literary circle from late 19th to mid-20th century."
@@
 Stream of consciousness (SOC) is a narrative technique that aims to capture the continuous flow of thoughts, feelings, and sensations experienced by a character without conventional organization or punctuation [@resampling].
-It mirrors the unpredictable and interconnected nature of human thought processes, often revealing the inner workings of the character's mind in an intimate and unfiltered manner [@flow].
-In literature, most scholars agree that stream of consciousness reveals the complexities of cognition, shedding light on the nuances of characters' emotional well-being and psychological struggles [@pathology].
+It mirrors the unpredictable and interconnected nature of human thought processes, often revealing the emotional well-being and psychological struggles of the character's mind in an intimate and unfiltered manner [@flow] [@pathology].
 As such, this paper has mined the texts of a total of nine novels from the volunteer archive, Project Gutenberg, to examine the mental health themes of famous stream of consciousness authors, namely by Joyce, Woolf, Proust, Mansfield and Eliot, from the modernist era of literature, spanning from late 19th century to mid-20th century [@gutenberg].
@@
 By analyzing these textual datasets through word frequency and sentiment analysis, this paper seeks to pose and answer crucial questions: \textit{What are some important factors contributing to this relationship between mental health, disassociation and stream of consciousness?
-Moreover, how does this relationship vary differently across differnet demographics of authors, for instance, authors with different geographical locations and genders?} Understanding these dynamics is crucial in having an informed understanding of the West's late 19th to mid-20th century literature and even socio-political landscape, especially in regards to how authors and creative writers navigate and deal with then-taboo topics such as existential angst, mental health issues and disabilities.
-
-Thus, the estimand is the correlation between mental health-related themes and words in SOC literature, their frequency and sentiment value as provided by @rTidytext.
-This is considered in terms of nine selected SOC literary works only, namely Joyce's \textit{A Portrait of the Artist as a Young Man and Chamber Music}; Woolf's \textit{Mrs Dalloway and Jacob's Room}; Proust's \textit{Swann Way}; Mansfield's \textit{Bliss and The Garden Party}; and Eliot's \textit{The Waste Land and The Love Song of J.
+Moreover, how does this relationship vary differently across differnet demographics of authors, for instance, authors with different genders?} Understanding these dynamics is crucial in having an informed understanding of the West's late 19th to mid-20th century literature and even socio-political landscape, especially in regards to how authors and creative writers navigate and deal with then-taboo topics such as existential angst, mental health issues and anxiety
+
+Thus, the estimand is the correlation between mental health-related themes in SOC literature, their appearance frequency and sentiment value.
+This is considered in terms of nine selected SOC texts only, namely Joyce's \textit{A Portrait of the Artist as a Young Man and Chamber Music}; Woolf's \textit{Mrs Dalloway and Jacob's Room}; Proust's \textit{Swann Way}; Mansfield's \textit{Bliss and The Garden Party}; and Eliot's \textit{The Waste Land and The Love Song of J.
 Alfred Prufrock}.
 Through our analysis, we found that words that appear with high frequency are mostly negative in select SOC literature (-1.875 median on AFINN index).
-Sentiment-wise, analysis reveals a predominant 27% to 37% of words in SOC texts are negative and sadness-inclined (NRC index), and over 60% of words in these texts are negative, with Eliot's works being the most negative at 68% (Bing index), followed closely by Joyce, Woolf, Proust, and Mansfield.
+Sentiment-wise, analysis reveals a predominant 27% to 37% of words in SOC texts being negative and sadness-inclined (NRC index), and over 60% of words in these texts are negative (Bing index).
 Abstract and temporal nouns like "time," "day," "life," "moment," and "night" are most frequent, respectively appearing 870, 597, 567, 502 and 376 times in all texts combined.
-These statistics correlate with SOC's non-linear nature, and the genre's exploration of the unconscious mind which often correlates to mental health themes.
-
-To further understand the correlation between stream of consciousness texts and mental health themes, in [Introduction], the paper briefly discusses the nature of stream of consciousness literature, relevant authors and the works that I've chosen to analyze.
+Gender-wise, male SOC authors are more likely to have negative words compared to female authors.
+These statistics correlate with SOC's non-linear nature, the genre's exploration of the unconscious mind (which often correlates to mental health themes), and its reflection of Western society in late 19th to mid-20th century.
+
+To further understand the correlation between stream of consciousness texts and mental health themes, in [Introduction], the paper briefly discusses the nature of stream of consciousness literature, relevant authors and the works that we have chosen to analyze.
 Subsequently, in [Data] and [Results], we talk about the nature of the data obtained and analyze the results garnered from the data with suitable tables and charts.
@@
 To do this, we downloaded all nine SOC texts from @gutenberg, using functions such as `unnest_tokens()` from @rTidytext to mine the texts, or separating them into individual words.
-Finally, I used `count()` to quantify the word count (`n`), and later divide the each word's count over the total word count of each author to garner the word frequency percentages.
+Finally, we used `count()` to quantify the word count (`n`), and later divide the each word's count over the total word count of each author to garner the word frequency percentages (`n/sum(n)`).
 
@@
 After the survey results are garnered, researchers average the answers to sort each surveyed word into pre-defined categories, specifically '-5 to 5' for 'AFINN', 'postive' or 'negative' for 'bing', and 'anger' or 'fear' for 'nrc'.
-With the continuous work of compiling these lexicons spanning years and decades [@sentiment] comes these functions: `get_sentiments("bing")` and `get_sentiments("nrc")` and `get_sentiments("afinn")` in which I can use `inner_join` to categorize my existing datasets of novel texts into their sentiment value.
+With the continuous work of compiling these lexicons spanning years and decades [@sentiment] comes these functions: `get_sentiments("bing")` and `get_sentiments("nrc")` and `get_sentiments("afinn")`.
+We can then use `inner_join` with these lexicons to categorize our existing datasets of SOC texts into pre-defined sentiment value.
 Systematic and data-driven, these measurement methods ensure that all lexicons faithfully reflects each word's emotionality.
@@
 
-Next, in @tbl-joyce-token, to measure the word count of each literary text, we first tokenized or separated into individual chunks of words using `unnest_tokens()` [@rTidytext].In addition, to account for stop words like, a, of, and, etc., the dataset `stop_words`and the function `anti_join` from @rTidytext were used to exclude all stop words from the self-generated datasets of literary texts, ensuring data validity and meaningful analysis.
+Next, in @tbl-joyce-token, to measure the word count of each literary text, we first tokenized or separated each text into individual chunks of words using `unnest_tokens()` [@rTidytext].In addition, to account for stop words like, a, of, and, etc., the dataset `stop_words`and the function `anti_join` from @rTidytext were used to exclude all stop words from the self-generated datasets of SOC texts, ensuring data validity and meaningful analysis.
 
@@
 
-To accurately compare word count among SOC authors, from @tbl-frequency, qw converted each word count from each author into a percentage.
-This is done by using `mutate` to calculate the word frequency of each word relative to the total word count of each author.
+To accurately compare word count among SOC authors, from @tbl-frequency, we converted each word count from each author into a percentage.
+This is done by using `mutate` to calculate the word frequency of each word relative to the total word count of each author (`n/sum(n)`).
 Due to the different number of words present in each author's literary works, this method accounts for the inconsistencies of between each author's word count.
-As explained, the variables within @tbl-frequency are:
+The variables within @tbl-frequency are:
 
@@
 
-@tbl-boxplot captures the text data garnered and their emotional connotations.
-Overseeing the entire dataset, @tbl-boxplot in particular demonstrates that, for all distinct word frequency, the median sentiment value is -2 on the scale between -5 and 5.
-More specifically, words that appear most frequently, or words with the frequency  of 0.0008511%, have the lowest mean of sentiment value-or the most negative connotation-of -2.33 on the scale between -5 and 5 in the AFINN sentiment index.
+@tbl-boxplot captures the text data garnered and their emotional connotations using the AFINN lexicon.
+Overseeing the entire dataset, @tbl-boxplot demonstrates that, for all distinct word frequency, the median sentiment value is -2 on the scale between -5 and 5.
+More specifically, words that appear most frequently, or words with the frequency of 0.0008511%, have the most negative connotation of -2.33 on the scale between -5 and 5 in the AFINN sentiment index.
 This summary statistic further shows how SOC literature mostly associates with words that carry more negative or even extreme connotations often associated with mental health themes, and topics such as depression, disassociation or schizophrenia [@pathology], a central argument that we will further discuss in [Results].
@@
 
-The actual raw text data sourced from Project Gutenberg may be susceptible to measurement errors due to various factors, such as selection bias towards older, Western and canonical texts, social desirability bias that edits out certain political elements, memory lapses, or misinterpretations of literary texts by voluntary proofreaders.
+The actual raw text data sourced from Project Gutenberg may be susceptible to measurement errors due to various factors, such as selection bias towards older, Western and canonical texts, social desirability bias (which edits out certain political elements), memory lapses, misinterpretations of literary texts by voluntary proofreaders.
 While the archive has strict guidelines to minimize these errors, they cannot be entirely eliminated.
@@
 By exclusively representing the genre through a select group of canonical Western authors, we risk constraining the genre's diversity.
-By oversimplifying the complexity of the the dynamic literary genre, we also affect the integrity of the datasets and its correlation to mental health themes.
+By oversimplifying the complexity of the literary genre, we also affect the integrity of the datasets and its correlation to mental health themes.
 
@@
 
-Upon assessing the generated datasets from Project Gutenberg, we further extrapolate on them here by combining word count with the AFINN, Bing and NRC index to conduct relevant word frequency and sentiment analysis.
-The later sections of Results also compound on these findings to conduct a comparative analysis between gendered and transnational SOC authors, attempting to garner relevant insights in mental health themes in different demographics of the SOC genre.
+Upon assessing our generated datasets from Project Gutenberg, we further extrapolate on them here by combining word count with the AFINN, Bing and NRC index to conduct relevant word frequency and sentiment analysis.
+The later sections of Results also compound on these findings to conduct a comparative analysis between gendered SOC authors, attempting to garner relevant insights in mental health themes in different demographics of the SOC genre.
 
@@
 However, words with high frequency-ranging from 0,0004255% to 0.0008511%-mostly contain negative connotations like fear and negative.
-They also have a -1.875 median on a scale between -5 to 5 in the AFINN index.
+They also have a negative -1.875 median on a scale between -5 to 5 in the AFINN index.
 Similarly, for the Bing index, words with higher frequency-those ranging from 0,0004255% to 0.0008511%-largely carry negative connotations.
@@
 
-To further explore the contextual relationship between each word, @fig-word-networks creates word networks of all nine of these SOC texts.
-These networks will capture how often each word frequents in these SOC texts and how closely each correlates with others, whether as single words (unigrams) or pairs of words (bigrams).
+To further explore the contextual relationship between each word, @fig-word-networks creates word networks of all nine SOC texts.
+These networks capture how often each word frequents in these SOC texts and how closely each correlates with one another, whether as single words (unigrams) or pairs of words (bigrams).
 When combined all SOC texts, temporal nouns like \textit{time, day, life, moment, night} appear the most, at, respectively, 870, 597, 567, 502 and 376 times ([Appendix]'s @tbl-count).
 This further correlates to the non-linear nature of SOC literature and their exploration of cognitive incoherence.
-In terms of correlation, forgoing personal names like \textit{swann, odette or mme}, bigrams with a correlation of 1 include those like \textit{dark, love}; \textit{day, heart}; \textit{moment, life}; \textit{mind, life}; \textit{moment, mind}, among others ([Appendix]'s @tbl-correlation).
+In terms of correlation, ignoring personal names like \textit{swann, odette or mme}, bigrams with a correlation of 1 include those like \textit{dark, love}; \textit{day, heart}; \textit{moment, life}; \textit{mind, life}; \textit{moment, mind}, among others ([Appendix]'s @tbl-correlation).
 Observable from  @fig-word-networks, with a correlation and a word count of at least 0.8 and 400, \textit{people, moment, life} and \textit{mind} are the most predominant word combinations throughout all 9 SOC texts.
@@
 
-For Bing index, @fig-sentiment-category shows that, SOC authors' literary works are predominantly negative.
+For the Bing index, @fig-sentiment-category shows that, SOC authors' literary works are predominantly negative.
 All authors, except for Mansfield, demonstrates this.
@@
 
-The NRC index in @fig-sentiment-category further details the emotional nuances of SOC literature.
+For the NRC index, @fig-sentiment-category further details the emotional nuances of SOC literature.
 The dominant emotional connotation in SOC literature's vocabulary is negative, standing at 16% to 22%.
 While the second highest percentage range belong to words with positive connotation, ranging from 11% to 17%.
-However, most of the other bars are occupied by words with negative-leaning emotions, such as sadness, fear, and anger.
+However, most of the other bars in the graph are occupied by words with negative-leaning emotions, such as sadness, fear, and anger.
 Sadness-related words, in particular, ranges the third highest, from 11% to 15%.
@@
 Eliot's works are the most negative with 37% of words being -2 on the emotional scale between -5 and 5.
-Respectively, Woolf, Proust and Joyce tail beind at 33%, 32% and 31%, averaging out to be ~-2.2 on the emotional scale between -5 and 5.
-While a large percentage Mansfield's works are positive, her highest percentage is negative on the AFINN index, at 24% for a scale of about -2.2 between -5 and 5.
+Respectively, Woolf, Proust and Joyce tail beind at 33%, 32% and 31%, averaging out to be -2.2 on the emotional scale between -5 and 5.
+While a large percentage Mansfield's works are positive, her highest percentage is negative on the AFINN index, at 24% or a -2.2 negative score between -5 and 5.
 
@@
 
-## Gendered Mental Landscape of Stream of Consciousness Novels
-
-\newpage
+## Gendered Mental Landscape of Stream of Consciousness Authorw
+
+To further expand on the analysis, we have also used our word frequency data from @tbl-frequency to conduct a gendered analysis between different SOC authors.
+@fig-word-gender attempts to assess the word frequency between female and male SOC authors.
+Dots closer together around the diagonal line tend to appear with similar frequency for all authors involved.
+Words further up and to the left are more typical for those on the y-axis, and closer to the lower right corner: those on the x-axis.
+
+@fig-word-gender shows that for female SOC authors, while having smaller word frequency due to the lighter blue, still have their words perfectly revolved the regression line, showing that both Woolf and Mansfield are comparatively close in terms of linguistic patterns.
+Woolf and Mansfield mostly have a word frequency of 0.0001421262% and 0.0001277302% for each word present in their texts, with some of their shared vocabulary being \textit{anxious}, \textit{angry}, \textit{anguish}, \textit{dreadful}, among others.
+
+As for male authors, @fig-word-gender shows that, while their words have a higher appearance frequency, they have less correlation as the regression line isn't centered, showing the dissimilar vocabulary used between male authors.
+Mostly, Eliot, Proust and Joyce have a word frequency of 0.0004255319%, 0.0000819% and 0.000112% for each word present in their texts, with some of their shared vocabulary being \textit{agony}, \textit{cry}, \textit{die}, \textit{chaos}, among others.
+Proust and Joyce, two novelists, seem to have more similarities compared to Eliot, who is a poet.
+
+While sharing the same existential quest in their SOC literature, through this gendered analysis and @fig-sentiment-category, we can see discrepancies between male and female authors, especially in terms of shared linguistic patterns and mental health themes.
+@gender poses that in terms of genders, writers-or here SOC writers-tend to write about affects or mental health themes differently.
+While females tend to have reservations over their emotions or are similarly positive, males tend to be more open and individualistic due to less social constraints at the time, which will be further discussed in [Discussion] below.
+
 ```{r}
@@
 #| fig-subcap: ["Female Authors", "Male Authors"]
+
+bing_lexicon <- get_sentiments("bing") %>%
+  rename(sentiment_binary = sentiment)
 
@@
   pivot_longer(`Virginia Woolf`,
-               names_to = "author", values_to = "proportion")
-
-ggplot(frequency_female, aes(x = proportion, y = `Katherine Mansfield`, color = abs(`Katherine Mansfield` - proportion))) +
+               names_to = "author", values_to = "proportion") |>
+    inner_join(bing_lexicon) |>
+  inner_join(get_sentiments("afinn")) |>
+  inner_join(get_sentiments("nrc"))
+
+ggplot(frequency_female, aes(x = proportion, y = `Katherine Mansfield`, fill = value, color = abs(`Katherine Mansfield` - proportion))) +
   geom_abline(color = "gray50", lty = 2) +
   geom_jitter(alpha = 0.1, size = 1, width = 0.5, height = 0.5) +  
-  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
-  scale_x_log10(labels = scales::percent_format()) +
+geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +  scale_x_log10(labels = scales::percent_format()) +
   scale_y_log10(labels = scales::percent_format()) +
@@
   pivot_longer(`Marcel Proust`:`James Joyce`,
-               names_to = "author", values_to = "proportion")
+               names_to = "author", values_to = "proportion") |>
+    inner_join(bing_lexicon) |>
+  inner_join(get_sentiments("afinn")) |>
+  inner_join(get_sentiments("nrc"))
 
@@
                        low = "steelblue", high = "darkblue") +
-  facet_wrap(~author, ncol = 2) +
+  facet_wrap(~author, ncol = 1) +
   theme(legend.position="none") +
@@
 
-## Transnational SOC Novels and Mental Health Themes
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-frequency-int
-#| fig-cap: "Comparative Analysis of Word Frequency in Transnational Stream of Consciousness Authors"
-
-frequency_international <- bind_rows(
-                       mutate(swann_words, author = "Marcel Proust"),
-                       mutate(bliss_words, author = "Katherine Mansfield"),
-                       mutate(portrait_words, author = "James Joyce")) |>
-  mutate(word = str_extract(word, "[a-z']+")) |>
-  count(author, word) |>
-  group_by(author) |>
-  mutate(proportion = n / sum(n)) |>
-  select(-n) |>
-  pivot_wider(names_from = author, values_from = proportion) |>
-  pivot_longer(`Katherine Mansfield`:`James Joyce`,
-               names_to = "author", values_to = "proportion")
-
-ggplot(frequency_international, aes(x = proportion, y = `Marcel Proust`, color = abs(`Marcel Proust` - proportion))) +
-  geom_abline(color = "gray50", lty = 2) +
-  geom_jitter(alpha = 0.05, size = 1, width = 0.8, height = 0.8) +  
-  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
-  scale_x_log10(labels = percent_format()) +
-  scale_y_log10(labels = percent_format()) +
-  scale_color_gradient(limits = c(0, 0.001), 
-                       low = "steelblue", high = "darkblue") +
-  facet_wrap(~author, ncol = 1) +
-  theme(legend.position="none") +
-  labs(y = "Marcel Proust", x = NULL)
-
-```
-
 # Discussion
@@
 The findings in this paper reveals a correlative relationship between SOC literature and mental health-inclined vocabulary and linguistic patterns.
-In terms of frequency, a dominant portion-or over 60% of words in the Bing index-of select SOC texts in this analysis have negative and sadness-leaning words ([@fig-sentiment-frequency]) ([@fig-sentiment-category]) ([@fig-sentiment-number]).
+In terms of frequency, a dominant portion-or over 60% of words in the NRC index-of select SOC texts in this analysis have negative and sadness-leaning words ([@fig-sentiment-frequency]) ([@fig-sentiment-category]) ([@fig-sentiment-number]).
 Most repeated words are abstract, temporal nouns like \textit{time}, \textit{mind} or \textit{life}, which are aligned with SOC literature's exploration of mind wandering, non-linearity and unconscious cognition ([@fig-word-networks]).
@@
 Most of the texts express the desire to replicate the inner workings of the human mind textually, attempting to capture the complexity of thought processes and emotions.
-With these inner-workings revealed to be mostly negative and existential, SOC authors have explored their identity and worldview in an authentic manner, exploring then-taboo topics such as existential angst, mental health issues and cognitive dissonance.
+With these inner-workings revealed to be mostly negative and existential, SOC authors have explored their identity and worldview in an authentic manner, deeply portraying then-taboo topics such as existential angst, anxiety and cognitive dissonance.
 
@@
 
-Through the analysis, we can see the SOC texts' linguistic patterns reflect the socio-political landscape of the Western hemisphere.
-The SOC genre stems out of modernist literature, whose writers often rebelled against linear and traditional narrative techniques of the 19th century [@resampling].
+Through the analysis, we can see the SOC texts' linguistic patterns reflect the socio-political landscape of Western society.
+The SOC genre originated from modernist literature, whose writers often rebelled against linear and traditional narrative techniques of the 19th century [@resampling].
 According to scholar Nyongesa, "marginalization occasioned by the dominant group result in neurotic conditions in members of the minority group" [@pathology], which rings true to the scenario at hand about SOC literature and its relationship to its birthplace: late 19th to mid-20th century Western society.
@@
 
-This time period also underwent the West's crisis of faith in traditional religious beliefs.
+This time period also saw the West's crisis of faith in traditional religious beliefs.
 With the emergence of Darwin's theory of evolution, Freudian psychology, and existential philosophy came different notions of human existence, morality, and meaning [@resampling].
-The heightened search for individuality of modernism had SOC authors embraced themes of alienation, identity and mental health in an increasing fragmented world.
+The heightened search for individuality had SOC authors embraced themes such as alienation, identity and mental health in an increasing fragmented world.
 The devastation from World War I and the interwar period also further intensified this crisis.
 Post-war collapse of nations created new ideologies like communism and fascism.
-This in turn created an atmosphere of disillusionment and uncertainty that SOC literature attempted to answer with its disassociative and schizophrenic style.
-
-## Schizophrenic and Disassociative Tendencies in Female Stream of Consciousness
+This in turn created an atmosphere of disillusionment and uncertainty that SOC literature attempted to answer with its anxious and disassociative writing style.
+
+## Gender Differences within Stream of Consciousness Genre and Mental Health
+
+Further, our gendered analysis of SOC literature in @fig-word-gender reveals differences in linguistic patterns between male and female authors during this time period, particularly in their portrayal of affects and mental health themes.
+Female writers tend to exhibit more reserved emotions or positivity, while male writers display greater emotional range and individualism, influenced by restrictions of female expressions at their time.
+
+Research findings indicate that, on average, female writers use positive words more frequently than male writers, reflecting societal norms attributing positive emotions to females [@gender].
+The similarity in vocabulary among SOC female writers like Woolf and Mansfield may stem from shared-or restricted-literary influences, higher cultural and social expectations, common personal experiences, and considerations of reader norms during the West's late 19th to mid-20th century.
+These factors contribute to a nuanced understanding of gendered linguistic patterns, which has been shown to be more forgiving towards male authors than female authors in the SOC genre.
 
@@
 For instance, the commonly repeated words \textit{time}, \textit{mind} or \textit{life} in our selected SOC texts might mean something entirely different when combined with other words.
-Paradoxically, these words may be insignificant to the SOC texts in question themselves as they're only small units of large literary texts that often have convoluted and complex themes.
-Therefore word frequency analysis might have excluded nuances when analyzing the contexts behind mental health themes and their relationship to SOC literature.
-
-In addition, the limiting SOC text editions or SOC literature from Project Gutenberg also disallow us from obtaining large sample of SOC literature to analyze their relationship to mental health themes [@gutenberg].
-Previously mentioned, since Project Gutenberg only accepts works in the public domain, their archive of literature includes solely old, Western and canonical works that might have missed a diverse range of other SOC texts or other editions that could have garnered better insights into our analysis.
+Paradoxically, these words may be insignificant to the SOC texts in question themselves as they're only small units of already expansive literary works that often have complex themes.
+Therefore, word frequency analysis might have not been nuanced when analyzing the contexts behind mental health themes and their relationship to SOC literature.
 In addition, not all SOC texts in the archive are available to be transcribed into datasets due to licensing issues, further limiting our analysis.
@@
 However, a sole dependence on the emotional metrics of these three indices might not be nuanced enough.
-This is because to categorize words into a specific emotion is an arduous and complex tasks as emotions are spectral and words are mainly context-based.
+This is because to categorize words into a specific emotion is an arduous and complex tasks, as emotions aren't easily quantifiable and words are mainly context-based.
 In addition, some of the indices used might not ensure valid data analysis.
-For instance, @tbl-frequency shows that, using the NRC index, the word \textit{accomplished} can carry two different emotions: joy and positive.
-This flexible categorization can very much skew the results of the sentiment analysis, bringing an emotion more to the forefront than the others.
+For instance, @tbl-frequency shows that, using the NRC index, the word \textit{accomplished} can mean two different emotions: joy and positive.
+This flexible categorization can very much skew the results of the sentiment analysis, bringing one emotion more to the forefront than others.
 
@@
 
-As mentioned, relying on three sentiment indices are not enough to accurately capture the emotional sentiments of the words used in SOC literature due to how context-based words are, thus more relevant indices should be incorporated into the analysis.
+As mentioned, relying on three sentiment indices are not enough to accurately capture the emotional sentiments of the words used in SOC literature due to how context-based words are.
+More relevant indices then should be incorporated into the analysis.
 
@@
 
-Lastly, due to the qualitative nature of SOC literature, using a more mixed methods approach, combining both quantitative and qualitative data analysis, will offer a more comprehensive understanding of both the broader mental health trends and the literary genre itself.
+Lastly, due to the qualitative nature of SOC literature, using a more mixed methods approach-combining both quantitative and qualitative data analysis-will offer a more comprehensive understanding of the broader mental health trends and the literary genre itself.
 
@@
 Through our analysis of works by Joyce, Woolf, Proust, Mansfield, and Eliot, we uncover that SOC literature is largely negative or sadness-inclined with a philosophical and temporal undertones due to its exploration of unconscious cognition.
+In addition, expression and sentiment-wise, female authors' writing is more restricted compared to male authors, revealing gendered differences in SOC literature.
 While our quantitative analysis provides insights into the genre's relationship with mental health themes, it's important to recognize its limitations in capturing the richness of SOC language.
-This study therefore asks for further exploration into the relationship between SOC literature and mental health discourse by using more advanced NLP methods.
-We hypothesize that this can provide deeper insights into the genre, and the reasons for its growth in the West's late 19th to mid-20th century society.
+Therefore, moving forward, using more advanced NLP methods will better benefit our exploration of the relationship between SOC literature and mental health themes.
+We hypothesize that this can provide deeper insights into the genre, and the reasons for its growth in Western society during late 19th to mid-20th century.
 

### 2024-04-24 paper/paper.qmd
final paper updates

@@
 This further correlates to the non-linear nature of SOC literature and their exploration of cognitive incoherence.
-In terms of correlation, ignoring personal names like \textit{swann, odette or mme}, bigrams with a correlation of 1 include those like \textit{dark, love}; \textit{day, heart}; \textit{moment, life}; \textit{mind, life}; \textit{moment, mind}, among others ([Appendix]'s @tbl-correlation).
+In terms of correlation, ignoring personal names like \textit{swann, odette or mme}, bigrams with a correlation of 1 include word pairs like \textit{dark, love}; \textit{day, heart}; \textit{moment, life}; \textit{mind, life}; \textit{moment, mind}, among others ([Appendix]'s @tbl-correlation).
 Observable from  @fig-word-networks, with a correlation and a word count of at least 0.8 and 400, \textit{people, moment, life} and \textit{mind} are the most predominant word combinations throughout all 9 SOC texts.
@@
 
-The AFINN index in @fig-sentiment-number portrays a similar trend with its more quantitative sentiment analysis.
+The AFINN index in @fig-sentiment-number portrays a similar trend.
 On a scale of -5 to 5, SOC literature from these five authors mostly carry a negative sentiment, observable in the significant difference in height between the negative trendlines and the positive trendlines.
 Eliot's works are the most negative with 37% of words being -2 on the emotional scale between -5 and 5.
-Respectively, Woolf, Proust and Joyce tail beind at 33%, 32% and 31%, averaging out to be -2.2 on the emotional scale between -5 and 5.
-While a large percentage Mansfield's works are positive, her highest percentage is negative on the AFINN index, at 24% or a -2.2 negative score between -5 and 5.
+Respectively, Woolf's, Proust's and Joyce's tail beind at 33%, 32% and 31%, averaging out to be -2.2 on the emotional scale between -5 and 5.
+While a large percentage Mansfield's works are positive, 24% of her words-the largest percentage of hers-are negative, earning a -2.2 negative score between -5 and 5 in the AFINN index.
 
@@
 
-While sharing the same existential quest in their SOC literature, through this gendered analysis and @fig-sentiment-category, we can see discrepancies between male and female authors, especially in terms of shared linguistic patterns and mental health themes.
+Through this gendered analysis and @fig-sentiment-category, we can see discrepancies between male and female SOC authors, especially in terms of shared linguistic patterns and mental health themes.
 @gender poses that in terms of genders, writers-or here SOC writers-tend to write about affects or mental health themes differently.
@@
 As scholars Bernini and Fernyhough mentioned, the "mysterious nature of the conscious" is "a simultaneously fragmented and unified mental realm" [@resampling], this analysis has also reflected that sentiment in SOC literature.
-Most of the texts express the desire to replicate the inner workings of the human mind textually, attempting to capture the complexity of thought processes and emotions.
+Most of the texts express the desire to textually replicate the inner workings of the human mind, attempting to capture the complexity of thought processes and emotions.
 With these inner-workings revealed to be mostly negative and existential, SOC authors have explored their identity and worldview in an authentic manner, deeply portraying then-taboo topics such as existential angst, anxiety and cognitive dissonance.

### 2024-04-24 paper/paper.qmd
final paper updates

@@
 The dominant emotional connotation in SOC literature's vocabulary is negative, standing at 16% to 22%.
-While the second highest percentage range belong to words with positive connotation, ranging from 11% to 17%.
+While the second highest percentage range belongs to words with positive connotation, ranging from 11% to 17%.
 However, most of the other bars in the graph are occupied by words with negative-leaning emotions, such as sadness, fear, and anger.
-Sadness-related words, in particular, ranges the third highest, from 11% to 15%.
+Sadness-related words, in particular, range the third highest, from 11% to 15%.
 

### 2024-04-24 paper/paper.qmd
final paper updates

@@
 Respectively, Woolf's, Proust's and Joyce's tail beind at 33%, 32% and 31%, averaging out to be -2.2 on the emotional scale between -5 and 5.
-While a large percentage Mansfield's works are positive, 24% of her words-the largest percentage of hers-are negative, earning a -2.2 negative score between -5 and 5 in the AFINN index.
+While a great portion of Mansfield's works are positive, 24% of her words-her largest percentage-are negative, earning a -2.2 negative score between -5 and 5 in the AFINN index.
 

### 2024-04-24 paper/datasheet_project_gutenberg.qmd
final datasheet updates

@@
 5. *Is there a label or target associated with each instance? If so, please provide a description.*
-	- Each instance is eponymously titled.
+	- Each instance is eponymously labelled, for instance 'book ID' is labelled 'book_id'.
 6.
@@
 This does not include intentionally removed information, but might include, for example, redacted text.*
-	- There is no missing information from individual instance since all texts have been in the public domain for a while, thus all data is up-to-date.
+	- There is no missing information from individual instances since all texts have been in the public domain for a while, thus all data is up-to-date.
 7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*

### 2024-04-24 paper/datasheet_project_gutenberg.qmd
final paper and datasheet updates

@@
 7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
-	- Yes, the relationships between individual instances are made explicit through their eponymously titled labels and the book IDs as provided from Project Gutenberg, which helps amalgamate data by each individual SOC text.
+	- Yes, the relationships between individual instances are made explicit through their eponymously titled labels and the book IDs as provided from Project Gutenberg, which helps clarify which data belongs to which individual SOC text.
 8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*

### 2024-04-24 paper/paper.qmd
final paper and datasheet updates

@@
 Next, custom datasets containing each author's raw texts were created.
-This involved using the identification number of the desired literary work and the `gutenberg_download` function to retrieve the HTML texts.
+This involved using the identification number of the nine desired literary works and the `gutenberg_download` function to retrieve the HTML texts.
 Subsequently, the data was saved into individual Parquet datasets, as demonstrated in @tbl-joyce.
@@
 
-Next, in @tbl-joyce-token, to measure the word count of each literary text, we first tokenized or separated each text into individual chunks of words using `unnest_tokens()` [@rTidytext].In addition, to account for stop words like, a, of, and, etc., the dataset `stop_words`and the function `anti_join` from @rTidytext were used to exclude all stop words from the self-generated datasets of SOC texts, ensuring data validity and meaningful analysis.
+Next, in @tbl-joyce-token, to measure the word count of each literary text, we first tokenized or separated each text into individual chunks of words using `unnest_tokens()` [@rTidytext].
+In addition, to account for stop words like, a, of, and, etc., the dataset `stop_words`and the function `anti_join` from @rTidytext were used to exclude all stop words from the self-generated datasets of SOC texts, ensuring data validity and meaningful analysis.
 

### 2024-04-24 paper/paper.qmd
final paper updates

@@
 @tbl-boxplot captures the text data garnered and their emotional connotations using the AFINN lexicon.
-Overseeing the entire dataset, @tbl-boxplot demonstrates that, for all distinct word frequency, the median sentiment value is -2 on the scale between -5 and 5.
-More specifically, words that appear most frequently, or words with the frequency of 0.0008511%, have the most negative connotation of -2.33 on the scale between -5 and 5 in the AFINN sentiment index.
-This summary statistic further shows how SOC literature mostly associates with words that carry more negative or even extreme connotations often associated with mental health themes, and topics such as depression, disassociation or schizophrenia [@pathology], a central argument that we will further discuss in [Results].
+Overseeing the entire dataset, @tbl-boxplot demonstrates that, for all distinct word frequency, the median sentiment value is a negative -2 on the scale between -5 and 5.
+More specifically, words that appear most frequently, or words with the frequency of 0.0008511%, have the most negative connotation of a -2.33 mean on the scale between -5 and 5 in the AFINN sentiment index.
+This summary statistic further shows how SOC literature mostly associates with words that carry  negative or even extreme connotations often associated with mental health themes, and topics such as depression, disassociation or schizophrenia [@pathology], a central argument that we will further discuss in [Results].
 

### 2024-04-24 paper/paper.qmd
final paper updates

@@
 #| label: tbl-joyce
-#| tbl-cap: "An Examplary Table Containing Unprocessed Novel Text (James Joyce)"
+#| tbl-cap: "Sample Table Containing Unprocessed Novel Text (James Joyce)"
 #| warning: false
@@
 #| label: tbl-joyce-token
-#| tbl-cap: Tokenized Stream of Consciousness Text (James Joyce)
+#| tbl-cap: Sample Table of Tokenized Stream of Consciousness Text (James Joyce)
 #| warning: false
@@
 #| label: tbl-joyce-count
-#| tbl-cap: Word Count after Tokenization (James Joyce)
+#| tbl-cap: Sample Table of Word Count after Tokenization (James Joyce)
 #| warning: false
@@
 #| layout-ncol: 2
-#| fig-cap: "Comparative Analysis of Top 20 Word Frequencies from Famous SOC Authors"
+#| fig-cap: "Comparative Analysis of Top 20 Word Count from Famous SOC Authors"
 #| fig-subcap: ["James Joyce", "Virginia Woolf", "Marcel Proust", "T.S. Eliot", "Katherine Mansfield"]
@@
 #| label: tbl-frequency
-#| tbl-cap: "Word Frequency of Stream of Consciousness Novels, A Comparison Between Five Authors "
+#| tbl-cap: "Sample Table of Word Frequency in Stream of Consciousness Texts"
 #| warning: false

### 2024-04-24 paper/paper.qmd
final paper updates

@@
 
-Thus, the estimand is the correlation between mental health-related themes in SOC literature, their appearance frequency and sentiment value.
-This is considered in terms of nine selected SOC texts only, namely Joyce's \textit{A Portrait of the Artist as a Young Man and Chamber Music}; Woolf's \textit{Mrs Dalloway and Jacob's Room}; Proust's \textit{Swann Way}; Mansfield's \textit{Bliss and The Garden Party}; and Eliot's \textit{The Waste Land and The Love Song of J.
+Thus, the estimand is the correlation between mental health-related words in SOC literature, their appearance frequency and sentiment value.
+This is considered in terms of nine SOC texts only, namely Joyce's \textit{A Portrait of the Artist as a Young Man and Chamber Music}; Woolf's \textit{Mrs Dalloway and Jacob's Room}; Proust's \textit{Swann Way}; Mansfield's \textit{Bliss and The Garden Party}; and Eliot's \textit{The Waste Land and The Love Song of J.
 Alfred Prufrock}.
-Through our analysis, we found that words that appear with high frequency are mostly negative in select SOC literature (-1.875 median on AFINN index).
-Sentiment-wise, analysis reveals a predominant 27% to 37% of words in SOC texts being negative and sadness-inclined (NRC index), and over 60% of words in these texts are negative (Bing index).
-Abstract and temporal nouns like "time," "day," "life," "moment," and "night" are most frequent, respectively appearing 870, 597, 567, 502 and 376 times in all texts combined.
+Through our analysis, we found that words that appear with high frequency are mostly negative in select SOC literature (-1.875 median out of 5 on the AFINN index).
+Sentiment-wise, analysis reveals that a predominant 27% to 37% of words in SOC texts are negative and sadness-inclined (NRC index), and over 60% of words in these texts are negative (Bing index).
+Abstract and temporal nouns like "time," "day," "life," and "moment" are the most frequent, respectively appearing 870, 597, 567, and 502 times in all texts combined.
 Gender-wise, male SOC authors are more likely to have negative words compared to female authors.
@@
 The SOC texts used for analysis were sourced from Project Gutenberg under the library `gutenbergr` [@rGuten] [@gutenberg].
-Data was generated, extracted and cleaned using the open-source statistical programming language R [@r], leveraging functions from `tidyverse` [@rTidyverse], `tidytext` [@rTidytext], `rmarkdown` [@rMark], `dplyr` [@rDplyr], `ggplot2` [@rGgplot2], `scales` [@rScales], `here` [@rHere], `igraph` [@rIgraph], `widyr` [@rWidy], `ggraph` [@rGg], `textdata` [@rTextdata], `tm` [@rTm], `here` [@here], `kableExtra` [@kableExtra], `arrow` [@arrow], and `knitr` [@rKnitr].
-More information on the dataset or the HTML textual data sourced from Project Gutenberg is available in the documented datasheet available here within this folder, paper/datasheet_project_gutenberg.pdf, located within the paper's GitHub repository.
+Data was generated, extracted and cleaned using the open-source statistical programming language R [@r], leveraging functions from `tidyverse` [@rTidyverse], `tidytext` [@rTidytext], `rmarkdown` [@rMark], `dplyr` [@rDplyr], `ggplot2` [@rGgplot2], `scales` [@rScales], `here` [@rHere], `igraph` [@rIgraph], `widyr` [@rWidy], `ggraph` [@rGg], `textdata` [@rTextdata], `tm` [@rTm], `kableExtra` [@kableExtra], `arrow` [@arrow], and `knitr` [@rKnitr].
+More information on the dataset or the HTML textual data sourced from Project Gutenberg is available in the documented datasheet available within this folder, paper/datasheet_project_gutenberg.pdf, located within the paper's GitHub repository.
 
@@
 
-* **Word Frequency:** This variable captures the repetition of a single word (unigram) or a pair of words combination (bigram) throughout a SOC novel text, providing us with a thematic understanding of SOC literature.
+* **Word Frequency:** This variable captures the repetition percentage of a single word (unigram) or a pair of words combination (bigram) throughout a SOC novel text, providing us with a thematic understanding of SOC literature.
 * **Sentiment Value:** This variable enables us to analyze how every word is usually perceived emotionally, whether it be a qualitative feeling or if it is conveyed through a numerical value.

### 2024-04-25 paper/paper.qmd
final paper updates

@@
 However, this paper does recognize how reductive this quantification of language can be.
-When it comes to understand such social and human artifacts as language or literary texts, much is dependent on their context.
+When it comes to understand such social and human artifacts as language, much depends on their context.
 As such, this paper will further discuss these weaknesses of the datasets under [Discussion].
@@
 
-## Combined Bibliographies of Nine Selected Stream of Consciousness Texts
+## Combined Bibliographies from Five Authors
 
@@
 #| label: tbl-combined
-#| tbl-cap: "Sample Combined Bibliographies of 9 Selected Stream of Consciousness Texts"
+#| tbl-cap: "Sample Combined Bibliographies of Stream of Consciousness Texts from Five Authors"
 #| warning: false

### 2024-04-25 paper/paper.qmd
final paper updates

@@
 #| layout-ncol: 2
-#| fig-cap: "Overall Word Frequency and Sentiment Analysis"
+#| fig-cap: "Boxplots of Sentiment Scores by Author (Bing & NRC)"
 #| fig-subcap: ["Bing lexicon", "NRC lexicon"]
@@
   geom_boxplot() +
-  labs(title = "Boxplot of Sentiment Scores by Author (Bing)",
+  labs(
        x = "Word Frequency Percentage",
@@
   geom_boxplot() +
-  labs(title = "Boxplot of Sentiment Scores by Author (NRC)",
+  labs(
        x = "Word Frequency Percentage",

### 2024-04-25 paper/paper.qmd
final paper updates

@@
 #| layout-ncol: 2
-#| fig-cap: "Boxplots of Sentiment Scores by Author (Bing & NRC)"
+#| fig-cap: "Boxplots of Sentiment Scores by Word Frequency (Bing & NRC)"
 #| fig-subcap: ["Bing lexicon", "NRC lexicon"]

### 2024-04-25 paper/paper.qmd
final paper updates

@@
 The findings in this paper reveals a correlative relationship between SOC literature and mental health-inclined vocabulary and linguistic patterns.
-In terms of frequency, a dominant portion-or over 60% of words in the NRC index-of select SOC texts in this analysis have negative and sadness-leaning words ([@fig-sentiment-frequency]) ([@fig-sentiment-category]) ([@fig-sentiment-number]).
+In terms of frequency, a dominant portion-over 60% of words in the NRC index-from select SOC texts have negative and sadness-leaning words ([@fig-sentiment-frequency]) ([@fig-sentiment-category]) ([@fig-sentiment-number]).
 Most repeated words are abstract, temporal nouns like \textit{time}, \textit{mind} or \textit{life}, which are aligned with SOC literature's exploration of mind wandering, non-linearity and unconscious cognition ([@fig-word-networks]).

### 2024-04-25 paper/paper.qmd
final paper updates

@@
   geom_text(aes(label = n), hjust = -0.3, vjust = 0.5, color = "grey40", size = 3) +
-  labs(x = "Frequency", y = "Word")
+  labs(x = "Count", y = "Word")
 

### 2024-04-25 paper/datasheet_project_gutenberg.qmd
final datasheet updates

@@
 If not, you may skip the remaining questions in this section.*
-	- No preprocessing/cleaning/labeling on the data was done by Project Gutenbeg.
+	- No preprocessing/cleaning/labeling on the data was done by Project Gutenbeg as all archived texts are digitized versions of printed books. They are not new editions.
 2.

### 2024-04-25 paper/paper.qmd
final paper updates

@@
 #| layout-ncol: 2
-#| fig-cap: "Boxplots of Sentiment Scores by Word Frequency (Bing & NRC)"
+#| fig-cap: "Boxplots of SOC Literature's Sentiment Scores by Word Frequency (Bing & NRC)"
 #| fig-subcap: ["Bing lexicon", "NRC lexicon"]
@@
 #| layout-ncol: 2
-#| fig-cap: "Word Networks Measured by Frequency and Correlation when Combining All Stream of Consciousness Novels"
+#| fig-cap: "Word Networks Measured by Frequency and Correlation when Combining All Stream of Consciousness Texts"
 #| fig-subcap: ["250 Frequency, 0.2 Correlation", "300 Frequency, 0.4 Correlation", "350 Frequency, 0.6 Correlation", "400 Frequency, 0.8 Correlation"]
@@
 #| layout-ncol: 2
-#| fig-cap: "Categorical Sentiment Analysis of All SOC Novel Texts by Authors"
+#| fig-cap: "Categorical Sentiment Analysis of All SOC Texts by Authors"
 #| fig-subcap: ["Bing lexicon", "NRC lexicon"]
@@
 #| label: fig-sentiment-number
-#| fig-cap: "Numerical Sentiment Analysis of All SOC Novel Texts by Authors (AFINN lexicon)"
+#| fig-cap: "Numerical Sentiment Analysis of All SOC Texts by Authors (AFINN lexicon)"
 
@@
 #| layout-ncol: 2
-#| fig-cap: "Comparative Analysis of Word Frequency in Female and Male Stream of Consciousness Authors"
+#| fig-cap: "Comparative Analysis of Word Frequency in Female and Male SOC Authors"
 #| fig-subcap: ["Female Authors", "Male Authors"]

### 2024-04-25 paper/paper.qmd
final paper updates

@@
 #| layout-ncol: 2
-#| fig-cap: "Word Networks Measured by Frequency and Correlation when Combining All Stream of Consciousness Texts"
+#| fig-cap: "Word Networks Measured by Frequency and Correlation when Combining All SOC Texts"
 #| fig-subcap: ["250 Frequency, 0.2 Correlation", "300 Frequency, 0.4 Correlation", "350 Frequency, 0.6 Correlation", "400 Frequency, 0.8 Correlation"]
